### 2023-10-26 README.Rmd
update README

---
title: ""
output: github_document
editor_options: 
  markdown: 
    wrap: sentence
always_allow_html: true
---

```{r setup, include=F, echo=F}
library(tidyverse)
devtools::load_all("D:/repositories/projectlog")
#source('scripts/dependencies.R')
knitr::opts_chunk$set(echo = F, message = F, warning = F)

```
<br>
This repository contains data, code, and output for a project entitled "Working memory performance in adverse environments: Enhanced, impaired, or intact?". A Stage 1 snapshot of the project has been submitted to [Peer Community in Registered Reports (PCI-RR)](https://rr.peercommunityin.org/)

# Project History

Below is an overview of all the major milestones of this project, including a description and links.
The links take you to the historic version of the project that corresponds to the milestone.

```{r}
proj_history <-
  gert::git_tag_list() |>
  dplyr::rename(
    milestone = name,
    milestone_hash = commit) |>
  mutate(
    commit_hash = (\(x) purrr::pmap_chr(list(milestone, ref, milestone_hash), .f = function(milestone, ref, milestone_hash) {
      gert::git_commit_info(milestone_hash)$id
    }))()
  ) |>
  left_join(
    gert::git_log() |>
      rename(commit_hash = commit)
  ) |>
  mutate(
    milestone_link = paste0(get_git_url(), '/releases/tag/', milestone),
    link = commit_hash,
    commit_link = paste0(get_git_url(), '/tree/', commit_hash)
  )

reactable::reactable(
  proj_history |> select(milestone, timestamp = time, author, message, link),
  columns = list(
    milestone = reactable::colDef(html = TRUE, cell = function(value, index) {
      sprintf("<a href='%s' target='_blank'>%s</a>", proj_history$milestone_link[index], value)
    }),
    link = reactable::colDef(html = TRUE, cell = function(value, index) {
      sprintf("<a href='%s' target=' _blank'>%s</a>", proj_history$commit_link[index], value)
    })
  )
)
```
### Contact

For questions or comments, feel free to contact me at p.c.s.vermeent@gmail.com.

### 2023-10-26 README.Rmd
update README to html
README with flextable

@@
 ---
-title: ""
-output: github_document
+title: "Working memory performance in adverse environments: Enhanced, impaired, or intact?"
+output: 
+  github_document:
 editor_options: 
@@
     wrap: sentence
-always_allow_html: true
 ---
@@
     commit_link = paste0(get_git_url(), '/tree/', commit_hash)
+  ) |> 
+  separate(message, into = c("message", "code"), sep = "code") |> 
+  mutate(
+    code = str_replace_all(code, "\\|\\>", "\\|\\>\\\n"),
+    message = str_remove(message, "\\\nobject_hash.*$")
   )
 
-reactable::reactable(
-  proj_history |> select(milestone, timestamp = time, author, message, link),
-  columns = list(
-    milestone = reactable::colDef(html = TRUE, cell = function(value, index) {
-      sprintf("<a href='%s' target='_blank'>%s</a>", proj_history$milestone_link[index], value)
-    }),
-    link = reactable::colDef(html = TRUE, cell = function(value, index) {
-      sprintf("<a href='%s' target=' _blank'>%s</a>", proj_history$commit_link[index], value)
-    })
+flextable::flextable(proj_history) |> 
+  flextable::autofit() |> 
+  flextable::compose(
+    j = "milestone",
+    value = flextable::as_paragraph(
+      flextable::hyperlink_text(x = milestone, url = milestone_link)
     )
+  ) |> 
+  flextable::compose(
+    j = "link",
+    value = flextable::as_paragraph(
+      flextable::hyperlink_text(x = "Click here", url = commit_link)
     )
+  ) |> 
+  flextable::delete_columns(c("ref", "milestone_hash", "commit_hash", "files", "merge", "milestone_link", "commit_link"))
 ```
+
 ### Contact

### 2023-10-26 README.Rmd
add README files

@@
     code = str_replace_all(code, "\\|\\>", "\\|\\>\\\n"),
-    message = str_remove(message, "\\\nobject_hash.*$")
+    message = str_remove(message, "\\\nobject_hash.*$"),
+    milestone = str_c("[", milestone, "]", "(www.google.com)")
   )

### 2023-10-26 README.qmd
final version of readme

--- 2023-10-26 README.Rmd
@@
 title: "Working memory performance in adverse environments: Enhanced, impaired, or intact?"
-output: 
-  github_document:
+output: github_document
 editor_options: 
@@
 #source('scripts/dependencies.R')
-knitr::opts_chunk$set(echo = F, message = F, warning = F)
+
+knitr::opts_chunk$set("ft.shadow" = FALSE, echo = F, message = F, warning = F)
 
@@
 This repository contains data, code, and output for a project entitled "Working memory performance in adverse environments: Enhanced, impaired, or intact?".
-A Stage 1 snapshot of the project has been submitted to [Peer Community in Registered Reports (PCI-RR)](https://rr.peercommunityin.org/)
+A Stage 1 snapshot of the project has been submitted to [Peer Community in Registered Reports (PCI-RR)](https://rr.peercommunityin.org/).
 
@@
 
-flextable::flextable(proj_history) |> 
+knitr::knit_print(flextable::flextable(proj_history) |> 
   flextable::autofit() |> 
@@
   ) |> 
-  flextable::delete_columns(c("ref", "milestone_hash", "commit_hash", "files", "merge", "milestone_link", "commit_link"))
+  flextable::delete_columns(c("ref", "milestone_hash", "commit_hash", "files", "merge", "milestone_link", "commit_link")))
 ```

### 2023-10-26 README.Rmd
revert to simple README for now

--- 2023-10-26 README.Rmd
@@
 ---
-title: ""
+title: "Working memory performance in adverse environments: Enhanced, impaired, or intact?"
 output: github_document
@@
 This repository contains data, code, and output for a project entitled "Working memory performance in adverse environments: Enhanced, impaired, or intact?".
-A Stage 1 snapshot of the project has been submitted to [Peer Community in Registered Reports (PCI-RR)](https://rr.peercommunityin.org/)
-
-# Project History
-
-Below is an overview of all the major milestones of this project, including a description and links.
-The links take you to the historic version of the project that corresponds to the milestone.
-
-```{r}
-proj_history <-
-  gert::git_tag_list() |>
-  dplyr::rename(
-    milestone = name,
-    milestone_hash = commit) |>
-  mutate(
-    commit_hash = (\(x) purrr::pmap_chr(list(milestone, ref, milestone_hash), .f = function(milestone, ref, milestone_hash) {
-      gert::git_commit_info(milestone_hash)$id
-    }))()
-  ) |>
-  left_join(
-    gert::git_log() |>
-      rename(commit_hash = commit)
-  ) |>
-  mutate(
-    milestone_link = paste0(get_git_url(), '/releases/tag/', milestone),
-    link = commit_hash,
-    commit_link = paste0(get_git_url(), '/tree/', commit_hash)
-  )
-
-reactable::reactable(
-  proj_history |> select(milestone, timestamp = time, author, message, link),
-  columns = list(
-    milestone = reactable::colDef(html = TRUE, cell = function(value, index) {
-      sprintf("<a href='%s' target='_blank'>%s</a>", proj_history$milestone_link[index], value)
-    }),
-    link = reactable::colDef(html = TRUE, cell = function(value, index) {
-      sprintf("<a href='%s' target=' _blank'>%s</a>", proj_history$commit_link[index], value)
-    })
-  )
-)
-```
-### Contact
-
-For questions or comments, feel free to contact me at p.c.s.vermeent@gmail.com.
+A Stage 1 snapshot of the project has been submitted to [Peer Community in Registered Reports (PCI-RR)](https://rr.peercommunityin.org/).

### 2023-12-13 scripts/0_pilot/01_create_scales.R
initiation of simulated data

@@ -1,0 +1,413 @@
+# Packages ----------------------------------------------------------------
+library(psych)
+library(tidyverse)
+library(qualtRics)
+library(jsonlite)
+library(here)
+library(sjlabelled)
+
+
+# Data --------------------------------------------------------------------
+pilot_data <-
+  fetch_survey(
+    surveyID = "SV_1B4wQ0QzRnlNEZU",
+    verbose  = T,
+    force_request = T,
+    label = F,
+    convert = F,
+    add_var_labels = F
+  ) %>%
+  rename_with(tolower) %>%
+  mutate(id = 1:n()) %>%
+  sjlabelled::var_labels(
+    id = "Blinded participant ID"
+  ) %>%
+  filter(finished==1, status == 0, `duration (in seconds)` > 0) %>%
+  select(-session_id)
+
+pilot2_data <-
+  fetch_survey(
+    surveyID = "SV_6JVjAWrPoHnjah8",
+    verbose  = T,
+    force_request = T,
+    label = F,
+    convert = F,
+    add_var_labels = F
+  ) %>%
+  rename_with(tolower) %>%
+  mutate(id = 1:n()) %>%
+  filter(finished==1) %>%
+  select(-session_id)
+
+
+# Self-report -------------------------------------------------------------
+
+## meta data ----
+
+vars01_meta <-
+  pilot_data %>%
+  rename(
+    meta_duration       = `duration (in seconds)`,
+    meta_start          = startdate,
+    meta_end            = enddate,
+    meta_recorded       = recordeddate,
+    meta_finished       = finished,
+    meta_feedback       = feedback_tasks
+  ) %>%
+  separate(meta_resolution, into = c("meta_resolution_width", "meta_resolution_height"), sep = "x") %>%
+  mutate(
+    meta_resolution_height = as.numeric(meta_resolution_height),
+    meta_resolution_width  = as.numeric(meta_resolution_width),
+    meta_resolution_ratio  = meta_resolution_width / meta_resolution_height) %>%
+  mutate(
+    meta_task_duration        = timestamp_tasks - timestamp_consent,
+    meta_ace_duration         = timestamp_ace - timestamp_tasks,
+    meta_dems_duration        = timestamp_dems - timestamp_ace,
+
[…]

### 2023-12-13 scripts/0_pilot/02_assess_data.R

@@ -1,0 +1,132 @@
+
+ospan_data03 <- ospan_data |>
+  filter(variable == 'recall') |>
+  mutate(acc_prop = accuracy / set_size) |>
+  group_by(id, set_size) |>
+  summarise(wm_cap = mean(acc_prop)) |>
+  mutate(set_size = str_c("wm_cap_", set_size))
+
+ospan_data_math03 <-
+  ospan_data |>
+  filter(variable == "math") |>
+  select(id, rt, block, counterbalance, correct) |>
+  group_by(id) |>
+  summarise(acc_ospan_math = sum(correct) / n())
+
+ospan_data03 |>
+  ggplot(aes(wm_cap)) +
+  geom_histogram() +
+  facet_grid(set_size ~ .)
+
+ospan_data_math03 |>
+  ggplot(aes(acc_ospan_math)) +
+  geom_density()
+
+
+bind_upd_number_data03 <- bind_upd_number_data |>
+  mutate(type = ifelse(nUpd > 0, "updating", "binding")) |>
+  mutate(steps = ifelse(type == "binding", nBind, nUpd)) |>
+  mutate(comb = str_c(nBind, nUpd, sep = "_")) |>
+  group_by(id, comb) |>
+  summarise(num_acc = sum(accuracy)/n()) |>
+  ungroup() |>
+  pivot_wider(names_from = 'comb', values_from = 'num_acc') |>
+  select(-id) |>
+  cor() |>
+  corrplot::corrplot(method = "number")
+
+bind_upd_color_data03 <- bind_upd_color_data |>
+  mutate(type = ifelse(nUpd > 0, "updating", "binding")) |>
+  mutate(steps = ifelse(type == "binding", nBind, nUpd)) |>
+  mutate(type_steps = str_c(type, steps, sep = "_")) |>
+  group_by(id, type_steps) |>
+  summarise(num_acc = sum(accuracy)/n()) |>
+  ungroup() |>
+  pivot_wider(names_from = 'type_steps', values_from = 'num_acc') |>
+  left_join(ospan_data02) |>
+  select(-id) |>
+  cor() |>
+  corrplot::corrplot(method = "number")
+
+
+
+# Timing ------------------------------------------------------------------
+
+timing_no_outliers <-
+  bind_rows(
+    study_data |>
+      select(id, data_bind_upd_number01, data_bind_upd_number02, data_bind_upd_number03) |>
[…]

### 2023-12-13 scripts/1_sim_pipeline/0_ivs.R

@@ -1,0 +1,238 @@
+# 1. Load libraries ----------------------------------------------------------
+
+library(tidyverse)
+
+# 2. Functions ---------------------------------------------------------------
+
+# INCOME-TO-NEEDS RATIO FOR DUTCH HOUSEHOLDS
+# the information up until 2021 was derived from van den Brakel et al. 2023 (See main text).
+# The information for 2022 was derived from https://www.cbs.nl/nl-nl/nieuws/2023/45/laagste-armoederisico-in-45-jaar-door-energietoeslag#:~:text=De%20lage%2Dinkomensgrens%20voor%20een,de%20grens%201%20830%20euro.
+# At the moment of writing, there was no final data for 2023, so we used the same values as 2022.
+inr_conv_table <- tibble(
+  year            = rep(str_c(20, c("07", "08", "09", 10:23)), each = 8),
+  adult           = rep(c(1, 1, 1, 1, 2, 2, 2, 2), 17),
+  aantalki        = rep(c(0, 1, 2, 3, 0, 1, 2, 3), 17),
+  eq_fct_pre2018  = rep(c(1, 1.33, 1.51, 1.76, 1.37, 1.67, 1.88, 2.06), 17),
+  eq_fct_post2018 = rep(c(1, 1.32, 1.52, 1.73, 1.40, 1.69, 1.91, 2.09), 17)
+) |>
+  mutate(
+    # `base` is the low-income threshold for a single-person household without children.
+    # All other threshold are derived from the base by multiplying it with the equivalence factor.
+    base = case_when(
+      year %in% c("2007", "2008", "2009")                 ~ 870,
+      year %in% c("2010", "2011", "2012", "2013", "2014") ~ 940,
+      year %in% c("2015", "2016", "2017")                 ~ 1030,
+      year == "2018"                                      ~ 1060,
+      year == "2019"                                      ~ 1090,
+      year == "2020"                                      ~ 1100,
+      year == "2021"                                      ~ 1130,
+      year == "2022"                                      ~ 1200,
+      year == "2023"                                      ~ 1200
+    ),
+    threshold = ifelse(str_detect(year, "(18|19|20|21|22|23)$"), base*eq_fct_post2018, base*eq_fct_pre2018) |> plyr::round_any(x = _, 10, f = ceiling))
+
+
+# 3. Stage 1: Simulate data for computational reproducibility ----------------
+
+set.seed(4652757)
+
[…]

### 2023-12-13 scripts/1_sim_pipeline/3_sem.R

@@ -1,0 +1,66 @@
+
+# 1. Libraries and Data ------------------------------------------------------
+
+library(lavaan)
+library(tidyverse)
+
+# Simulated dataset with standardized variables
+data_clean <- tibble(
+  id             = 1:800,
+  family_id  = 1:800,
+  ospan          = rnorm(800, 0, 1), # Operation Span
+  rspan          = rnorm(800, 0, 1), # Rotation Span
+  upd            = rnorm(800, 0, 1), # Updating
+  bind           = rnorm(800, 0, 1), # Binding
+  dep            = rnorm(800, 0, 1), # Material deprivation composite
+  threat         = rnorm(800, 0, 1), # Threat composite
+  age            = rnorm(800, 0, 1), # age
+  noise          = rnorm(800, 0, 1), # Environmental noise
+  interrupt      = sample(c(0,1), size = 800, prob = c(0.95, 0.05), replace = TRUE), # Was interrupted
+  leave_computer = sample(c(0,1), size = 800, prob = c(0.95, 0.05), replace = TRUE)  # Left computer during tasks
+)
+
+# 2. Specify Clustering Design -----------------------------------------------
+
+# Some children are clustered within families.
+# We account for this using the lavaan.survey package
+
+cluster_design <- survey::svydesign(ids=~family_id, prob=~1, data = data_clean)
+
+
+# 3. Model specification(s) --------------------------------------------------
+
+model1 <-
+  '
+  # Measurement Model
+
+  ## Latent factors
+  wmc_l      =~ 1*ospan + rspan + bind + upd
+  upd_l      =~ 1*upd
+
+  ## Covariances
+  upd   ~~ 0*upd
+
+  dep   ~~ threat
+
+  # Structural Model
+  wmc_l ~ dep + threat + age + interrupt + leave_computer
+  upd_l ~ dep + threat + age + interrupt + leave_computer
+'
+
+
+sem_fit1 <- lavaan::sem(
+  model = model1,
+  data = data_clean,
+  orthogonal = TRUE,
+  auto.cov.lv.x = FALSE,
+  auto.cov.y = FALSE,
+  std.lv = FALSE,
+  fixed.x = FALSE
+)
+
+sem_fit1_cluster <- lavaan.survey::lavaan.survey(lavaan.fit = sem_fit1, survey.design = cluster_design, estimator = "ML")
+
+summary(sem_fit1_cluster, fit.measures = TRUE, standardized = TRUE)
+
+

### 2023-12-13 scripts/1_simulations/unpredictability_measures.R

@@ -1,0 +1,57 @@
+
+# Libraries ---------------------------------------------------------------
+
+library(tidyverse)
+
+
+# Functions ---------------------------------------------------------------
+
+# Compute rowwise SDs
+RowSD <- function(x) {
+  sqrt(rowSums((x - rowMeans(x))^2)/(dim(x)[2] - 1))
+}
+
+# Create waves of data in line with within-subject specified Means and SDs
+create_wave <- function(data, names) {
+  data <- map(names, function(x) {
+    data <- data |>
+      rowwise() |>
+      mutate("{x}" := rnorm(1, mean, sd)) |>
+      ungroup()
+  }) |>
+    reduce(left_join)
+}
+
+
+# Simulate data -----------------------------------------------------------
+
+set.seed(3554)
+
+data <-
+  faux::rnorm_multi(n = 1000, vars = 2, mu = 0, sd = 1, r = -0.7, varnames = c("mean", "sd")) |>
+  as_tibble() |>
+  mutate(
+    id = 1:1000,
+    sd = (sd+abs(min(sd)))/2,
+    mean = (mean+abs(min(mean)))
+    ) %>% # make sure SDs are positive
+  create_wave(data = ., names = str_c("wave", 1:100)) |> # Simulate 100 waves
+  mutate(mat_dep = across(starts_with('wave')) %>% rowMeans) |> # Compute within-person mean level material deprivation
+  mutate(unpred  = across(starts_with("wave")) %>% RowSD) |>  # compute within-person SD of material deprivation over time
+  pivot_longer(starts_with('wave'), names_to = "wave", values_to = "value") |>
+  group_by(id) |>
+  mutate(prop_change = (value - lag(value, n = 1)) / lag(value, n = 1)) |>
+  summarise(
+    mat_dep = mean(value),
+    unpred = sd(value),
+    prop_change = mean(abs(prop_change), na.rm = T) # Unpredictability option 1: Average proportion of change
+  ) |>
+  mutate(sd_corrected = unpred / mat_dep) # Unpredictability option 2: within-subject SD divided by within-subject mean
+
+
+ggplot(data, aes(prop_change, sd_corrected)) +
+  geom_point()
+
+ggplot(data, aes(prop_change, unpred)) +
+  geom_point()
+

### 2023-12-13 manuscript/manuscript.qmd
update of in-progress Stage 1 manuscript

---
bibliography: references.bib
csl: apa.csl
format: 
  docx:
    reference-doc: reference-doc.docx
output:
  officedown::rdocx_document:
    page_margins:
      bottom: 1
      footer: 0
      gutter: 0
      header: 0.5
      left: 1
      right: 1
      top: 1
    plots:
      align: center
      caption:
        pre: 'Figure '
        sep: '. '
        style: Image Caption
    tables:
      caption:
        pre: 'Table '
        sep: '. '
        style: Table Caption
  pdf_document: default
  word_document: default
editor: 
  markdown: 
    wrap: sentence
---

```{r include = FALSE}
library(flextable)
library(stringr)
library(dplyr)
library(tidyr)
library(tibble)
library(lavaan)

knitr::opts_chunk$set(
  echo = F,
  fig.align = "center",
  fig.pos = "!t", 
  out.extra = "",
  fig.show = "asis",
  message = FALSE,
  tab.topcaption = T,
  warning = FALSE
)

# set up flextable for tables
set_flextable_defaults(
  font.family = "Times", 
  font.size = 10,
  font.color = "black",
  line_spacing = 1,
  padding.bottom = 1, 
  padding.top = 1,
  padding.left = 1,
  padding.right = 1
)

```

#### **Working memory performance in adverse environments: Enhanced, impaired, or intact?**

<br>

#### Stefan Vermeent^1,2^, Anna-Lena Schubert^3^, Meriah L. DeJoseph^4^, Jaap Denissen^1^, Jean-Louis van Gelder^2,5^, & Willem E. Frankenhuis^1,2,6^

#### ^1^ Department of Psychology, Utrecht University, Utrecht, The Netherlands

#### ^2^ Max Planck Institute for the Study of Crime, Security, and Law, Freiburg, Germany

#### ^3^ Department of Psychology, University of Mainz, Mainz, Germany

#### ^4^ Graduate School of Education, Stanford University, USA

#### ^5^ Department of Education and Child Studies, Leiden University, Leiden, The Netherlands

#### ^6^ Evolutionary and Population Biology, Institute for Biodiversity and Ecosystem Dynamics, University of Amsterdam, Amsterdam, the Netherlands

<br>

# Data Availability

All scripts and materials needed to reproduce the findings are available on the article's Github repository ([https://github.com/stefanvermeent/liss-wm-profiles-2023/README.md](https://github.com/stefanvermeent/liss-wm-profiles-2023/README.md)).
We also include instructions on how to reproduce each step of our analyses. In this paper, we make use of data from the LISS panel (Longitudinal Internet studies for the Social Sciences) managed by the non-profit research institute Centerdata (Tilburg University, the Netherlands). All datasets are available in the LISS data archive. Researchers who want to access the data are required to sign a statement confirming that information about individual persons, households, etc., will not be released to others (Go to [https://statements.centerdata.nl](https://statements.centerdata.nl) for more information).

# Funding statement
WEF’s contributions have been supported by the Dutch Research Council (V1.Vidi.195.130) and the James S. McDonnell Foundation (https://doi.org/10.37717/220020502). MLD was supported by the NICHD National Research Service Award (#1F32HD112065-01).

# Disclosures
We declare no conflicts of interest.

# Ethics Approval Statement

This study was approved by the Ethics Review Board of the Faculty of Social & Behavioural Sciences of Utrecht University (FETC20-490) and the Ethics committee for research in the Sciences and Life Sciences of the University of Amsterdam (FNWI-41_2023).

\pagebreak

# Proposal abstract

XXX

\pagebreak

# Working memory performance in adverse environments: Enhanced, impaired, or intact?

Living in adverse conditions, with prolonged exposure to intense stress, tends to have a substantial impact on cognitive functioning.
A domain that seems to be particularly affected by adversity is working memory (WM), a system for mentally building, maintaining, and updating immediately relevant information [@oberauer_2018].
WM is associated with a host of social and cognitive abilities, such as mathematics ability [@peng_2016], reading ability [@chiappe_2000], learning [@cowan_2014], general intelligence [@conway_2003], and mentalizing [@mutter_2006].
Not surprisingly, then, deficits in WM have negative consequences for educational and professional outcomes [@ahmed_2018; @alloway_2010; @guo_2020; @spiegel_2021].
Although decades of research show that adversity is generally negatively associated with WM [@goodman_2019], a small, emerging literature suggests that specific aspects of WM might be intact or even enhanced through developmental adaptations to adversity.
So far, these literatures have tended to focus on related, but different aspects of WM, limiting a fuller integration.
Here, we take a psychometric modeling approach to simultaneously examine potential decreases and enhancements in WM performance.

# Deficit-based and adaptation-based models

A large literature has shown negative associations between exposures to adversity and WM, potentially attributable to the enduring influence of stress on several key brain regions that support WM [@duval_2017; @hanson_2012].
Although not always mentioned explicitly, much of this work has focused on WM capacity, or the ability to keep multiple pieces of information simultaneously available for further processing.
For early-life adversity, this negative association is already present during childhood, and persists into adulthood [@bos_2009; @evans_2009; @hackman_2010; @farah_2006; @goodman_2019; @noble_2007; but see @nweze_2021].
Other studies have found a similar link between recent adverse experiences and lowered WM capacity in college student samples [@klein_2001; @shields_2017; @shields_2019].
The most often used tasks have been simple span tasks (repeating a string of stimuli of increasing length), complex span tasks (remembering a string of stimuli while being engaged by a secondary task), and *n*-back tasks (judging whether the current stimulus in a string is identical to the stimulus *n* steps ago)  [@goodman_2019].
Performance on these tasks is assessed through the number of items that can be retained in WM, that is, its overall capacity [with the exception of *n*-back\; for concerns about the construct validity of this task, see @frost_2021; @kane_2007].
Other research has shown that recent life stress is negatively associated with WM capacity, but not the precision of representations in WM [@shields_2019].

Although both early-life and recent adversity appear to lower WM capacity, a small set of studies suggest that exposure to adversity may leave intact, or even enhance, WM updating (i.e., the ability to rapidly replace old information in WM with new information) [@young_2018; @young_2022].
These findings exemplify an emerging, theoretical framework grounded in adaptation reasoning that is complementary to deficit frameworks [@frankenhuis_2013; @ellis_2017; @ellis_2022; @frankenhuis_2020].
This adaptation framework assumes that developmental processes tailor an individual's cognitive abilities to the unique challenges and opportunities posed by their environment.
The link between adversity and cognitive abilities is further assumed to be specific; as different types of adversity (e.g., threat vs. deprivation) pose different challenges, they should (at least in part) shape different cognitive abilities.
WM updating has been argued to be especially adaptive in environments that are unpredictable, which can be defined as random variation in harshness over space or time [@ellis_2009].
In such environments, WM updating allows people to maintain an up-to-date overview of the current state of the environment [@young_2018].
Improved WM updating performance has also been documented for threat exposure [@young_2022], and unpredictability might particularly shape cognition in circumstances of high deprivation, such as lower-resource contexts like poverty [@li_2018].

# Associations between WM capacity and updating

With deficit frameworks mainly focusing on WM capacity and adaptation frameworks on WM updating, we may wonder how these WM components are related to each other.
Research to date has demonstrated substantial overlap between capacity and updating tasks using structural equation modeling (SEM).
For example, correlations between latent capacity and updating factors can be high, to the extent that both complex span and updating paradigms are seen by some as valid measures of WM capacity [@wilhelm_2013].
In fact, some research suggests that the amount of updating-specific variance on typical updating tasks might be as low as 15% [@frischkorn_2022].
This overlap might stem from the fact that both WM updating and capacity tasks require maintaining information in WM for a short time [@ecker_2010; @frischkorn_2022; @oberauer_2000; @schmiedek_2009; @wilhelm_2013].

Nevertheless, updating tasks might involve processes other than maintenance that set them apart from WM capacity tasks.
Updating tasks require different combinations of retrieval (making information available for immediate processing), transformation (changing a prior value into a new one, e.g., by addition or subtraction), and substitution (replacing a prior value for a new value) [@ecker_2010].
@ecker_2010 measured three indicators of WM capacity as well as eight versions of a WM updating task that required different combinations of retrieval, transformation, and substitution.
They found positive correlations of around .50 between WM capacity with retrieval and transformation accuracy, whereas substitution accuracy was not correlated with WM capacity.
Thus, the ability to accurately substitute old with new information---a key aspect of WM updating---seems to be independent from WM capacity.

These findings underscore the importance of accounting for WM capacity when assessing a person's WM updating ability.
This is especially important in the context of adversity research, as previous studies suggest that certain types of adverse conditions might have opposing effects on WM capacity and updating [e.g., @goodman_2019; @young_2018; @young_2022].
Yet, to our knowledge, no previous research has analyzed both abilities within a single statistical model.
This could lead us to underestimate---or altogether miss---ways in which adversity enhances WM components, with implications for basic and applied science.
For basic science, it could bias inferences about WM abilities, especially when the negative association between adversity and WM capacity is stronger than the positive association with WM updating.
For applied science, it could hide from view potential pathways to leverage people's existing strengths in school or work contexts.

# Current study

In this study, we will estimate associations between three types of adversity exposure and latent estimates of WM capacity and updating.
Specifically, we will measure exposure to two forms of harshness: *neighborhood threat* (perceived neighborhood safety, personal crime victimization), and *material deprivation* (perceived scarcity, income-to-needs).
We further include a measure of unpredictability *unpredictability*, indexed as random variation in deprivation-induced harshness over time.
Together, these adversity types capture key dimensions in contemporary models of adversity [threat vs. deprivation vs. unpredictability\; @ellis_2009; @ellis_2022; @mclaughlin_2016; @mclaughlin_2019].
Although it would have also been theoretically interesting to look at unpredictability in neighborhood threat, our dataset does not include enough timepoints to compute such estimates.
We will address three research questions.
First, what is the association of adversity with WM capacity?
Second, what is the association of adversity with WM updating *after* accounting for WM capacity?
Third, are these two associations similar or different for neighborhood threat, material deprivation, and unpredictability?

Theoretically, we will evaluate evidence for deficit and adaptation frameworks.
Figure 1 summarizes which combinations of lowered, enhanced, and intact WM capacity and updating performance would provide support for deficit and adaptation frameworks.
Crucially, we assume that deficit and adaptation processes can operate simultaneously [@frankenhuis_2020], meaning that we could find (lack of) support for both frameworks in the same model.
A hypothesis generated within a deficit framework would be consistent with lowered performance on WM capacity or updating, but not with enhanced or intact performance.
This follows from previous literature hypothesizing broad WM deficits [@farah_2006; @sheridan_2020].
A hypothesis generated within an adaptation framework would be consistent with two different patterns.
First, if performance on WM updating is enhanced.
Second, if performance on WM updating is intact, but only in combination with lowered WM capacity performance.
If both updating and capacity are intact, this would instead suggest that the WM system is unaffected by adversity (i.e., neither impairment nor adaptation processes have shaped WM).

Figure 1B depicts the theoretical possibilities in relation to deficit or adaptation frameworks (as formalized in Figure 1A) and how we hypothesize them to be linked to threat, material deprivation, and unpredictability.
First, we expect that all three, and in particular material deprivation, will be associated with deficits in WM capacity, consistent with literature showing that deprivation is most strongly associated with cognitive deficits [@salhi_2021; @sheridan_2020].
In this literature, most studies have focused on cognitive deprivation, which is not identical to material deprivation (as measured here).
However, both are related to a person's access to stimulating resources (e.g., formal education, books\; @heppt_2022).
Most research has focused on children, so it is unclear whether and how cognitive and material deprivation affects cognitive outcomes in adults.

In contrast, we expect adaptations in WM updating to be most strongly associated with unpredictability [@young_2018; @young_2022], as well as with threat [in line with existing empirical findings\; @young_2022].
In addition, we expect an interaction between unpredictability and mean levels of deprivation, such that WM updating shows the highest increase under conditions of high material deprivation *and* high unpredictability.
Unpredictability is challenging to define, and has been measured in many different ways, ranging from number of residential changes, to perceived inconsistencies in household routines, to random variation in income [@ugarte_2023; @walasek_2023; @young_2020].

One issue is that timescales of unpredictability measures may not always align with those of the cognitive abilities that are assumed to be shaped by them.
For example, while residential changes occur on timescales of months or years, WM updating constitutes a process unfolding on the timescale of seconds to minutes.
However, we expect that our measure of unpredictability (random variation in perceived resource scarcity and income-to-needs based on yearly measures) is a proxy for everyday situations that require rapid updating of information in WM.
For example, yearly fluctuations in material deprivation, especially for someone living close to the poverty line, could reflect many everyday instances in which the person needs to update information about their current financial state in order to meet basic needs (e.g., because of an unpredictable income, or unforeseen costs or debts, that constrain everyday decisions like how much to spend on groceries).

<br>

```{r}
#| label: Figure1
#| fig-width: 6.5
#| fig-height: 7
#| dpi: 600
#| out-width: 6in
#| fig-cap: | 
#|   **Figure 1.** Overview of predictions derived from deficit and adaptation frameworks. Panel A depicts the most likely data patterns based on previous literature, and whether we would consider them consistent with deficit and adaptation frameworks (see the main text for more details). Panel B depicts an overview of the Structural Equation Model, including predictions for associations between different dimensions of adversity and WM components. All four manifest WM measures load on a latent WM capacity factor, reflecting the fact that people have to hold information active in WM on all tasks. WM updating is modeled as a latent factor capturing the residual variance in the updating task after accounting for variance related to WM capacity. 
knitr::include_graphics("fig/fig1.png")
```

<br>

# Methods

## Participants

We collected data in the Longitudinal Internet studies for the Social Sciences (LISS) panel [@scherpenzeel_2011].
The LISS panel is a representative probability sample of roughly 5,000 Dutch households (~7,500 individuals) drawn from the population register by Statistics Netherlands on an invite-only basis.
Households without a computer or internet connection are provided with these facilities by LISS.
Participants complete a yearly core battery of questionnaires about various aspects of their lives.
In addition, participants can complete online questionnaires every month, which are more variable.
The current study includes archived data as well as new data collected as part of a monthly data collection that took place between October 2nd and December 22nd, 2023.
We signed a contract with LISS stipulating that we will receive access to the data only after Stage 1 acceptance of the Registered Report.

We based our power analysis on simulations reported by @kretzschmar_2019, determining the required sample size to detect a small effect size ($\beta$ = 0.1) with at least 90% power at $\alpha$ = 0.05.
Assuming a reliability of at least 0.7 [which is typical for WM tasks with a number of trials similar to ours\; e.g., @wilhelm_2013], we would require a sample size of N = 730.
Anticipating some exclusions, we therefore decided to include 800 participants.
Participants were eligible for inclusion if they 1) were currently between 18 and 55 years old, 2) had completed at least one wave of the 'Conventional and computer crime victimization' survey ([https://doi.org/10.17026/dans-zch-j8xt](https://doi.org/10.17026/dans-zch-j8xt)) which ran between 2008 and 2018 at two-year intervals, and 3) had given permission to link their LISS data to government microdata (not relevant here). 
To ensure sufficient representation of people from lower socioeconomic backgrounds, half the total sample was sampled from participants who reported one or more of the following at least once in the three years: (1) a monthly income < €1,500, (2) HAVO or VWO as highest completed education (which are the two highest levels in Dutch secondary education), or (3) a score of 4 or lower on the 'ladder of life' ("If you imagine a ‘ladder of life’, where the first step represents the worst possible life, and the tenth (top) step the best possible life, on what step would you place yourself?").
Participants will be excluded if they (1) switched to and interacted with other browser tabls *during* one or more of the cognitive tasks, except when this happened in between tasks or during a scheduled break, (2)

## Measures

### Neighborhood threat

**Neighborhood crime.** We included four items from existing LISS data collected across six waves ([https://doi.org/10.17026/dans-zch-j8xt](https://doi.org/10.17026/dans-zch-j8xt)), in which participants answered how often it happens that they 1) "avoid certain areas in your place of residence because you perceive them as
unsafe", 2) "do not respond to a call at the door because you feel that it is unsafe", 3) "leave valuable items at home to avoid theft or robbery in the street?", 4) "make a detour, by car or on foot, to avoid unsafe areas?" on a scale of 1 ("(Almost) never"), 2 ("Sometimes"), or 3 ("Often").
We will recode these items so that 0 indicates "(Almost) never". 
We then sum the responses within each wave for which participants have data, and calculate an average across the waves.

In addition, we will use the Neighborhood Violence Scale [NVS\; @frankenhuis_2018; @frankenhuis_2020] which was included in the current study.
The NVS includes seven items measuring perceived exposure to neighborhood violence (e.g., "Crime is common in the neighborhood where I live"; "Where I live, it is important to be able to defend yourself against physical harm").
Participants answered these questions on a scale of 1 ("Completely disagree") to 7 ("Completely agree").

**Crime victimization.** We used existing LISS data collected across six waves ([https://doi.org/10.17026/dans-zch-j8xt](https://doi.org/10.17026/dans-zch-j8xt)), in which participants indicated whether they fell victim to eight types of crime over the two years prior to a particular wave.
We included seven items concerning physical crime: 1) burglary or attempted burglary; 2) theft from their car; 3) theft of their wallet or purse, handbag, or other personal possession; 4) wreckage of their car or other private property; 5) intimidation by any other means; 6) maltreatment of such serious nature that it required medical attention; 7) maltreatment that did not require medical attention.
For each type of crime that participants indicated having been exposed to, they also indicated the frequency.
We will average the frequencies for each type of crime separately across waves for which participants have data.

**Overall composite.** To compute an overall measure of neighborhood threat exposure, we will submit the averages of each measure (the two measures of neighborhood crime and average frequencies of each crime exposure) to a Principal Component Analysis (PCA).
We will extract a single factor score that explains the largest part of the variance.
Higher scores will indicate higher levels of neighborhood threat.

### Material deprivation

**Perceived scarcity.** We will use a few items from LISS data that was collected on a yearly basis between 2007 and 2023 ([https://doi.org/10.57990/1gr4-bf42](https://doi.org/10.57990/1gr4-bf42)) to index perceived scarcity.
First, participants indicated how hard or easy it currently is to live off the income of their household, on a scale of 0 (very hard) to 10 (very easy).
Second, participants were asked to choose which of the following best applied to their current situation: 1) "we are accumulating debt"; 2) "we are somewhat eating into savings"; 3) "we are just managing to make ends meet"; 4) "we have a little bit of money to spare"; 5) "we have a lot of money to spare".
Their response was reverse-coded, so that a higher score indicated a worse financial situation.
Third, participants answered which of the following issues they were confronted with at present (0 = no, 1 = yes): 1) "having trouble making ends meet"; 2) unable to quickly replace things that break"; 3) "having to lend money for necessary expenditures"; 4) "running behind in paying rent/mortgage or general utilities"; 5) "debt collector/bailiff at the door in the last month"; 6) "received financial support from family or friends in the last month".
Fourth, participants answered whether their household expenditure was more than, equal to, or less than their household income over the last 12 months, not counting large investments (e.g., buying a house). 
We will dummy-code this variable, using expenditure lower than income as the reference category.

**Income-to-needs.** We will calculate an income-to-needs ratio for each year using monthly self-reported net household income ([https://doi.org/10.57990/qn3k-as78](https://doi.org/10.57990/qn3k-as78)).
First, we average monthly reported income across years for each participant.
Then, we divide the average income per year by the *poverty threshold*, as determined by Statistics Netherlands (CBS) [@brakel_2023].
The poverty threshold changes across time to account for purchasing power.
In addition, different thresholds are determined each year for different household compositions (single adult vs (married) couple and number of children (up to three)) by multiplying the threshold of a single-person household without children with a particular equivalence factor.
As thresholds are only provided for households with up to three children, we will use the equivalence factor of a household with three children for households with more than three children.
Likewise, we will use the equivalence factor of a household with a (married) couple for households with more than two adults that are not children.

**Overall composite.** Similar to neighborhood threat, we will use PCA to extract a single estimate of material deprivation measured by perceived scarcity and income-to-needs.
In contrast to neighborhood threat, we will apply the PCA to the most recent data (2023) and use the resulting weights to compute material deprivation scores for previous years.
This approach will allow us to compute both the mean level of material deprivation, as well as random variation in material deprivation over time.
If a participant has partial data for a particular year, we will use multiple imputation to impute the value using the *mice* package [@buuren_2011].
We will not impute data for years where a participant has missings on all indicators.
We will compute within-subject averages of material deprivation across all years for which participants data available.
Higher scores will indicate more material deprivation.

### Unpredictability

**Overall composite.** Unpredictability will be based on the same PCA as used for material deprivation, involving the same indicators.
we operationalize unpredictability as within-person variability in material deprivation over time [@ugarte_2023; @young_2020].
Specifically, we will compute unpredictability by dividing the within-subject standard deviation of material deprivation by the within-subject mean-level of material deprivation. 
Dividing by the mean accounts for the fact that the mean and standard deviation in income have been found to be strongly negatively correlated, indicating that people with lower incomes tend to experience less variability in income [@li_2018; @young_2023].
For that reason, the standard deviation alone has been called into question as a measure of adversity, as the same fluctuation in income can have a greater relative impact for people close to the poverty line than for people with high incomes.

### WM tasks

All materials and scripts for the cognitive tasks can be found at [https://github.com/StefanVermeent/liss_wm_profiles_2023/tree/master/materials](https://github.com/StefanVermeent/liss_wm_profiles_2023/tree/master/materials).
Prior to collecting LISS data, we conducted a pilot study through Prolific in a Dutch sample of N = 100.
The main goals of this pilot study were to collect participant feedback (e.g., difficulty of instructions, whether we included sufficient breaks) and to analyze performance and correlations between tasks.
The results of this pilot study are described in more detail in the Supplemental Materials.

**Operation Span Task.** The Operation Span Task (Figure 2A) is a common measure of WM capacity [@conway_2005; @wilhelm_2013].
In this task, participants alternated between a primary memorization task and a secondary processing task.
On each trial, participant's task was to memorize a sequence of letters in the correct order (from a set of 12 letters).
Each letter was presented for 1,000 ms in the center of the screen.
Next, participants saw a simple mathematical equation including the outcome.
Their task was to indicate whether the outcome was correct or incorrect by pressing either the 'a' or 'l' key on their keyboard.
The equations always contained one addition or subtraction, with numbers ranging between one and 10.
Outcomes were always positive integers.
On each trial, participants had to memorize between four and six letters, with each set size repeated three times.
At the end of each sequence, all letters were presented in a 3X4 grid, and participants clicked the letters in the correct order.

Participants first practiced the letter task (three times), then the math task (eight times), and then the full task (three times).
If they performed at or below chance, they  had the opportunity to either repeat a part or advance to the next part.
After practicing, participants completed 12 test trials.

**Rotation Span Task.** The Rotation Span Task (Figure 2B) was similar to the Operation Span Task and was adopted from @wilhelm_2013.
On each trial, participants' task was to memorize the orientation of a sequence of arrows in the correct order.
Arrows could take on eight different orientations, with steps of 45$\circ$.
Each arrow was presented for 1,000 ms in the center of the screen. 
Next, participants saw a capital 'G' or 'F' that was rotated at one of eight different orientations, with steps of 45$\circ$.
Their task was to indicate whether the letter was mirrored or not.
On each trial, participants had to memorize between two and five arrows, with each set size repeated three times.
At the end of each sequence, all arrows were presented simultaneously, and participants clicked the arrows in the correct order.

Participants first practiced the arrow task (three times), then the letter task (eight times), and then the full task (three times).
If they performed at or below chance, they  had the opportunity to either repeat a part or advance to the next part.
After practicing, participants completed 12 test trials.

**Binding-Updating Task.** The Binding-Updating task (Figure 2C) was adopted from [XXX].
On each trial, participants saw a 3x3 grid, with a fixation cross in the central cell.
After 1,000 ms, they were presented with a sequence of numbers (0-9) in random locations of the grid.
Each new number was presented for 1,500 ms, after which it disappeared for 500 ms before the next number was presented.
The participants' task was to remember the last number they saw in each location.
Memory set sizes (i.e., the number of unique locations in the grid) ranged between three and five.
On half of the trials, only one number was presented in each location.
These constituted the binding-only trials.
On the other half of the trials, some letters were presented in the same location as previous numbers, requiring mentally replacing the old number with the new number.
These constituted the updating trials.
We used two, three, and four updating steps, each repeated in combination with the different set sizes.
At the end of the trials, participants indicated which letter they saw last in each location in random order.

Participants first completed four practice trials.
If they performed at or below chance, they had the opportunity to either repeat the practice trials or advance to the full task.
After practicing, they completed 18 test trials.
We will compute a binding score by calculating the overall accuracy (%) across trials with zero updating steps.
We will compute an updating score by calculating the overall accuracy (%) across trials containing updating steps.

```{r}
#| label: Figure2
#| fig-width: 6.5
#| fig-height: 7
#| dpi: 600
#| out-width: 6in
#| fig-cap: | 
#|   **Figure 2.** Overview of the working memory tasks. Panel A: Operation Span Task. Participants had to memorize letters in the correct order, while engaging in a secondary math task. Panel B: Rotation Span Task. Participants had to memorize the orientation of arrows, while judging whether letters were mirrored or normal in a secondary task. Panel C: Participants had to memorize numbers in the correct location in a 3x3 grid. On half of the trials, all numbers were presented in unique locations, only requiring binding the numbers to the correct position. On the other half, some numbers were presented in the same location as a previously presented number, requiring updating. Note that stimuli are not to scale. 
knitr::include_graphics("fig/fig2.png")
```

## Procedure

We received ethical approval from the first institutions ethical board.
Upon starting the study, participants were informed that the study could only be completed on a laptop or desktop PC.
Participants started with the WM tasks, which on average took between 20 and 25 minutes.
The WM tasks were completed in fullscreen mode. 
If participants left fullscreen mode at any moment during the tasks, they saw instructions at the top of their screen that allowed them to return to fullscreen mode.
the order of the WM tasks was counterbalanced, and participants had the opportunity to take breaks at regular intervals.

After the cognitive tasks, participants answered three questions about the environment in which they completed the WM tasks: 1) "How much noise was there in your environment during the memory tasks?"; 2) "Were you at any moment interrupted during the memory tasks?"; 3) "Did you at any moment during the memory tasks leave the computer?". 
Next, they completed questionnaires about their future orientation (not considered here), personality (not considered here), past adversity exposure, and recent adversity exposure.
Finally, they completed a standard set of evaluation questions asking about their experiences with the study, with the possibility to provide open-ended feedback.
This part on average took 5 minutes.
Participants received €7.50 for their participation through LISS.
If participants experienced difficulties of any sort, they could contact the LISS helpdesk.

## Proposed analysis plan

### Data access

The working memory data and one of the neighborhood threat indices were collected through October-December 2023, prior to submitting the Stage 1 protocol.
These data will only be made available to the first author after Stage 1 acceptance, as stipulated in a signed contract with LISS.
During planning of the study, the first author received access to the LISS data archive and inspected three waves of the LISS data containing the items about neighborhood safety and crime exposure, as well as the three most recent monthly data collections containing basic demographic info.
The reason was to discover how many people who completed the waves a few years back were still active in the panel (i.e., to see if we could reasonably create a link between these previous waves and newly collected data). 
All data access events were automatically detected and logged on the GitHub repository using the 'projectlog' R package [@vermeent_2023].
We took the following measures to prevent bias: 1) we randomly shuffled the participant IDs in each data set using 'projectlog', so that we were unable to link participant data between (waves of) studies; 2) we did not inspect any of the measures that will be part of our adversity composites; 3) we did not know which participants would be selected for the new LISS study; 4) the primary analyses will be based on composite measures that combine measures from these previous studies with data that were collected at a later timepoint.

### Primary analyses

See Figure 1B for an overview of the model specification.
We will fit a single model containing all adversity measures using the *lavaan* package [@rosseel_2012].
Missing data will be handled using full information maximum likelihood (FIML).
If participants are from the same household, this clustering within families will be accounted for using the *lavaan.survey* package [@oberski_2014].

WM capacity will be estimated as a latent factor loading on all outcome measures.
In addition, we will estimate WM updating as a latent factor capturing residual variance in the updating measure.
Thus, this factor accounts for updating-specific variance after accounting for WM capacity.
We will estimate the effect of each adversity type (dashed lines in Figure 1B) through regression analyses.
Each association is controlled for: (1) age in years ; (2) the quadratic effect of age; (2) environmental noise ("How noisy was your environment during the memory tasks", rated on a scale of 1 (very little noise) to 5 (a lot of noise)); (3) two items measuring interruptions ("Where you at any moment interrupted during the memory tasks?" and "Did you at any moment during the memory tasks leave your computer?", rated as yes or no). 
Goodness of fit will be assessed using the comparative fit index (CFI) and the root mean square error of approximation (RMSEA).
CFI values \> .90 and RMSEA values \< .08 will be interpreted as acceptable model fit, and CFI values \> .95 and RMSEA values \≤ .06 as good model fit.

We anticipate that we may have to optimize the model further in case of bad model fit, and will therefore estimate the model in two steps to prevent bias.
First, we will construct the measurement model only, without including the adversity measures.
This step will be carried out prior to accessing any of the adversity measures.
Once we obtain at least acceptable model fit, we will access and add the adversity measures to the model.
This procedure will be tracked and timestamped on the GitHub repository using the procedure outlined above.
We will control for multiple testing using the false discovery rate [@benjamini_1995; @cribbie_2007].
We will use two one-sided tests (TOST) equivalence testing to test whether small effects---which we define as standardized effects between -.10 and .10---are practically equivalent, which we will interpret as evidence for intact performance.

\pagebreak

# References

::: {#refs}
:::

\pagebreak

### 2023-12-13 materials/test/ospan/0_ospan_instructions.js
update WM materials

@@ -1,215 +1,0 @@
-//-------------------- Welcome
-var ospan_welcome = {
-  type: jsPsychInstructions,
-  pages: [
-    "U gaat nu beginnen aan het <b>Letters en Nummers</b> spel!"
-  ],
-  show_clickable_nav: true,
-  allow_backward: true,
-  key_forward: -1,
-  key_backward: -1,
-  button_label_next: "verder",
-  data: {
-    task: "ospan_welcome",
-    variable: 'welcome'}
-};
-
-
-//-------------------- Instructions
-
-// If participants start with the standard Ospan version, they will be introduced with the letter task in the standard visual appearance
-var ospan_instructions_letters = {
-  timeline: [
-    {
-      type: jsPsychInstructions,
-      pages: [
-        //Page 1
-        "<p style = 'text-align: center;'>"+
-          "In dit spel zult u steeds twee dingen moeten doen:<br><br>" +
-          "1. <strong>Letters</strong> onthouden.<br>" +
-          "2. Simpele <strong>rekensommen</strong> oplossen.",
-
-        //page 2
-        "<p style = 'text-align: center;'>"+
-          "<strong>Taak 1: Letters onthouden</strong><br><br>" +
-          "U krijgt &#233;&#233;n voor &#233;&#233;n een paar letters te zien.<br>" +
-          "Uw taak is om alle letters <strong>in de juiste volgorde</strong> te onthouden.<br>" +
-          "Een klein voorbeeld: U krijgt eerst een '<strong>S</strong>' te zien, dan een '<strong>Q</strong>', en tot slot een '<strong>F</strong>'.<br><br><br>" +
-          "<img width = 250 src = img/instr01.png></img><br><br>",
-
-        //page 5
-
-          "Na afloop klikt u op de letters <strong>in de juiste volgorde</strong>.<br><br>" +
-          "Gebruik de '<strong>?</strong>' knop als u een letter bent vergeten.<br>" +
-          "Gebruik de '<strong>corrigeer</strong>' knop om uw antwoord aan te passen.<br><br><br>" +
-          '<img width = 250 src = img/instr02.png></img><br><br>',
-
-          "Het is normaal als u niet alle letters kunt onthouden.<br><br>" +
-          "Gebruik <strong>GEEN</strong> hulpmiddelen, zoals pen en papier, uw mobiel, of uw vingers.<br>" +
[…]

### 2023-12-13 materials/test/ospan/1_ospan_procedure.js

@@ -1,350 +1,0 @@
-var preload_ospan = {
-  type: jsPsychPreload,
-  images: [
-     'img/instr01.png',
-     'img/instr02.png',
-     'img/instr03.png',
-  ]
-}
-
-
-var step_number = 0
-var possibleLetters = ["F","H","J","K","L","N","P","Q","R","S","T","V"];
-var nCorrectOspanRecall = 0
-var mathCorrect = false
-var nMathCorrect = 0
-var ospan_block = 1
-var OspanLettersPracticeCorrect = 0
-var OspanMathPracticeCorrect = 0
-
-// Empty grid with fixation cross
-var ospan_fixation = {
-
-  type: jsPsychHtmlKeyboardResponse,
-  stimulus: '<div style="font-size:60px;">+</div>',
-  choices: "NO_KEYS",
-  trial_duration: 1500,
-  data: {
-    task: "ospan_fixation",
-    variable: 'fixation',
-    step_number: step_number
-  }
-};
-
-var ospan_fixation_short = {
-
-  type: jsPsychHtmlKeyboardResponse,
-  stimulus: '<div style="font-size:60px;">+</div>',
-  choices: "NO_KEYS",
-  trial_duration: 1000,
-  data: {
-    task: "ospan_fixation",
-    variable: 'fixation',
-    step_number: step_number
-  }
-};
-
-var ospan_start_new_trial = {
-  type: jsPsychHtmlKeyboardResponse,
-  stimulus: '<div style="font-size:18px;">' + "Plaats uw vingers op de '<strong>A</strong>' en '<strong>L</strong>' toetsen op uw toetsenbord en druk op een willekeurige toets om te beginnen." + '</div>',
-  choices: "ALL_KEYS",
-  response_ends_trial: true,
-  data: {
-    task: "ospan_pretrial",
-    variable: 'pretrial',
-    step_number: step_number
-  }
-};
-
-var ospan_letter = {
-  type: jsPsychHtmlKeyboardResponse,
-  stimulus: function(){
-    return '<div style="font-size:60px;">' + jsPsych.timelineVariable("stim_l")[step_number] + '</div>'
-  },
-  choices: "No_KEYS",
-  trial_duration: 1000,
-  data: {
-    task: function(){return(jsPsych.timelineVariable("task"))},
-    variable: "letter",
-    block: ospan_block
-  }
-}
-
-var ospan_letter_feedback = {
[…]

### 2023-12-13 materials/test/ospan/functions/generate_math.js

@@ -1,33 +1,0 @@
-var cogloadf = function(correct){  // generates math questions
-
-
-
-  var possibleOperations = [" + ", " - "];
-  operation = jsPsych.randomization.sampleWithReplacement(possibleOperations, 1)[0];
-  if (operation==" + "){
-    num1 = Math.floor(jStat.uniform.sample(1, 10));
-    num2 =  Math.floor(jStat.uniform.sample(1, 10));
-    ans = num1 + num2;
-  } else if (operation==" - "){
-    num1 = Math.floor(jStat.uniform.sample(1, 10));
-    num2 = Math.floor(jStat.uniform.sample(1, (num1-1)));
-    ans = num1 - num2;
-  }
-
-  if (!correct){   // generates incorrect answers
-    ansDiff = jsPsych.randomization.sampleWithReplacement([1,2],1)[0];
-    coinFlip = jsPsych.randomization.sampleWithReplacement([true, false],1)[0];
-    if (coinFlip){
-      ans += ansDiff;
-    } else {
-      ans -= ansDiff;
-    }
-    if (ans<0){
-      ans += 2*ansDiff; //ensuring no negative incorrect answers
-    }
-  }
-
-  var equation = num1+operation+num2+' = '+ans;
-
-  return equation;
-};

### 2023-12-13 materials/test/ospan/functions/generate_timelines.js

@@ -1,103 +1,0 @@
-function shuffle_identical(arr1, arr2) {
-  var index = arr1.length;
-  var rnd, tmp1, tmp2;
-
-  while (index) {
-    rnd = Math.floor(Math.random() * index);
-    index -= 1;
-    tmp1 = arr1[index];
-    tmp2 = arr2[index];
-    arr1[index] = arr1[rnd];
-    arr2[index] = arr2[rnd];
-    arr1[rnd] = tmp1;
-    arr2[rnd] = tmp2;
-  }
-}
-
-
-var generate_bind_upd_timeline = function(nBind, nUpd, stimset, task){
-
-  var bindStims = shuffleWithoutReplacement(arr = stimset, size = nBind)
-
-  var uniqueRecallStims = []
-
-
-  // Make sure that array of recall stimuli does not contain any duplicate stimuli
-  while(uniqueRecallStims.length < bindStims.length) {
-
-    var updStims = shuffleWithReplacement(arr = stimset.filter(x => !bindStims.includes(x)), size = nUpd, immediate_reps = false)
-    var stims = bindStims.concat(updStims)
-
-    //var stims = shuffleWithoutReplacement(arr = stimset, size = nBind+nUpd)
-    var bindPos = shuffleWithoutReplacement(arr = [0,1,2,3, 5,6,7,8], size = nBind)
-    var updPos = shuffleWithReplacement(arr = bindPos, size = nUpd, immediate_reps = false)
-    var allPos = bindPos.concat(updPos)
-
-
-    var recallStims = []
-    var recallPos = []
-
-    var recall_i = allPos.length -1
-
-    while (recallPos.length < nBind) {
-
-      if (recallPos.includes(allPos[recall_i])) {
-        recall_i -= 1
-      } else {
-        recallPos.push(allPos[recall_i])
-        recallStims.push(stims[recall_i])
-        recall_i -= 1
-      }
-    }
-
-    uniqueRecallStims = recallStims.filter((item, i, ar) => ar.indexOf(item) === i);
-  }
-
-  // Randomize recall order and make sure stimuli and their positions are still matched across arrays
-  shuffle_identical(recallStims, recallPos)
-
-  return({stim: stims, pos: allPos, nBind: nBind, nUpd: nUpd, recall_pos: recallPos, recall_stim: recallStims, task: task})
-}
-
-
-generate_timeline_variables_ospan = function(setSize, task){
[…]

### 2023-12-13 materials/test/rspan/0_rspan_instructions.js

@@ -1,223 +1,0 @@
-//-------------------- Welcome
-var rspan_welcome = {
-  type: jsPsychInstructions,
-  pages: [
-    "U gaat nu beginnen aan het <b>Pijlen en Letters</b> spel!"
-  ],
-  show_clickable_nav: true,
-  allow_backward: true,
-  key_forward: -1,
-  key_backward: -1,
-  button_label_next: "verder",
-  data: {
-    task: "rspan_welcome",
-    variable: 'welcome'}
-};
-
-
-//-------------------- Instructions
-
-var rspan_instructions_arrows = {
-  timeline: [
-    {
-      type: jsPsychInstructions,
-      pages: [
-        //Page 1
-        "<p style = 'text-align: center;'>"+
-          "In dit spel zult u steeds twee dingen moeten doen:<br><br>" +
-          "1. De richting van <strong>pijlen</strong> onthouden.<br>" +
-          "2. Bepalen of letters <strong>gespiegeld</strong> zijn.",
-
-        //page 2
-        "<p style = 'text-align: center;'>"+
-          "<strong>Taak 1: Pijlen onthouden</strong><br><br>" +
-          "U krijgt &#233;&#233;n voor &#233;&#233;n een paar pijlen te zien.<br>" +
-          "Uw taak is om alle pijlen <strong>in de juiste volgorde</strong> te onthouden.<br>" +
-          "Een klein voorbeeld: U krijgt eerst een '<strong>pijl omhoog</strong>' te zien, dan een '<strong>pijl naar links</strong>', en tot slot een '<strong>pijl rechts naar beneden</strong>'.<br><br><br>" +
-          "<img width = 250 src = img/instr01.png></img><br><br>",
-
-        //page 5
-
-          "Na afloop klikt u op de pijlen <strong>in de juiste volgorde</strong>.<br><br>" +
-          "Gebruik de '<strong>?</strong>' knop als u een pijl bent vergeten.<br>" +
-          "Gebruik de '<strong>corrigeer</strong>' knop om uw antwoord aan te passen.<br><br><br>" +
-          '<img width = 250 src = img/instr02.png></img><br><br>',
-
-          "Het is normaal als u niet alle pijlen kunt onthouden.<br><br>" +
-          "Gebruik <strong>GEEN</strong> hulpmiddelen, zoals pen en papier, uw mobiel, of uw vingers.<br>" +
-          "U mag de pijlen wel hardop of in uw hoofd herhalen.<br><br>"
-      ],
[…]

### 2023-12-13 materials/test/rspan/1_rspan_procedure.js

@@ -1,516 +1,0 @@
-var preload_rspan_instr = {
-  type: jsPsychPreload,
-  images: [
-    "img/instr01.png",
-    "img/instr02.png",
-    "img/instr03.png"
-  ]
-}
-
-var preload_rspan_arr = {
-  type: jsPsychPreload,
-  images: [
-     'img/a1_short.png',
-     'img/a2_short.png',
-     'img/a3_short.png',
-     'img/a4_short.png',
-     'img/a5_short.png',
-     'img/a6_short.png',
-     'img/a7_short.png',
-     'img/a8_short.png',
-  ]
-}
-
-var preload_g_normal = {
-  type: jsPsychPreload,
-  images: [
-     'img/g_0_normal.png',
-     'img/g_45_normal.png',
-     'img/g_90_normal.png',
-     'img/g_135_normal.png',
-     'img/g_180_normal.png',
-     'img/g_225_normal.png',
-     'img/g_270_normal.png',
-     'img/g_315_normal.png'
-  ]
-}
-
-var preload_g_mirror = {
-  type: jsPsychPreload,
-  images: [
-     'img/g_0_mirror.png',
-     'img/g_45_mirror.png',
-     'img/g_90_mirror.png',
-     'img/g_135_mirror.png',
-     'img/g_180_mirror.png',
-     'img/g_225_mirror.png',
-     'img/g_270_mirror.png',
-     'img/g_315_mirror.png',
-  ]
-}
-
-
-var preload_r_normal = {
-  type: jsPsychPreload,
-  images: [
-     'img/r_0_normal.png',
-     'img/r_45_normal.png',
-     'img/r_90_normal.png',
-     'img/r_135_normal.png',
-     'img/r_180_normal.png',
-     'img/r_225_normal.png',
-     'img/r_270_normal.png',
-     'img/r_315_normal.png'
-  ]
-}
-
-var preload_r_mirror = {
-  type: jsPsychPreload,
-  images: [
-     'img/r_0_mirror.png',
-     'img/r_45_mirror.png',
-     'img/r_90_mirror.png',
-     'img/r_135_mirror.png',
-     'img/r_180_mirror.png',
-     'img/r_225_mirror.png',
-     'img/r_270_mirror.png',
-     'img/r_315_mirror.png',
-  ]
-}
-
-var preload_f_normal = {
-  type: jsPsychPreload,
-  images: [
-     'img/f_0_normal.png',
[…]

### 2023-12-13 materials/test/rspan/functions/recall_display.js

@@ -1,51 +1,0 @@
-var recall_display = function(stimulus = "?", position, correct){
-
-      // SET LETTERS FOR THE DISPLAY
-      var stim1 = (position == 0)? stimulus : "";
-      var stim2 = (position == 1)? stimulus : "";
-      var stim3 = (position == 2)? stimulus : "";
-      var stim4 = (position == 3)? stimulus : "";
-      var stim5 = (position == 4)? stimulus : "";
-      var stim6 = (position == 5)? stimulus : "";
-      var stim7 = (position == 6)? stimulus : "";
-      var stim8 = (position == 7)? stimulus : "";
-      var stim9 = (position == 8)? stimulus : "";
-
-      // Construct and display 3 X 3 grid
-      var html = "<style>" +
-        ".grid-container {" +
-          "display: inline-grid;" +
-          "grid-template-columns: auto auto auto;" +
-          "grid-template-rows: auto auto auto;" +
-          "grid-gap: 0px;" +
-        "}" +
-
-        ".grid-container > div {" +
-          "display: flex;" +
-          "justify-content: center;" +
-          "align-items: center;" +
-          "background-color: solid black;" +
-          "border: 1px solid black;" +
-          "font-size: 70px;" +
-          "font-family: sans-serif;" +
-          "height: 90px;" +
-          "width: 90px;" +
-        "}" +
-        "</style>" +
-
-      "<div class='grid-container'>" +
-      "<div>" + stim1 + "</div>" +
-      "<div>" + stim2 + "</div>" +
-      "<div>" + stim3 + "</div>" +
-      "<div>" + stim4 + "</div>" +
-      "<div>" + stim5 + "</div>" +
-      "<div>" + stim6 + "</div>" +
-      "<div>" + stim7 + "</div>" +
-      "<div>" + stim8 + "</div>" +
-      "<div>" + stim9 + "</div>" +
-      "</div>" +
-      "<br><br>" +
-      "<div style = 'margin: auto'><p><strong>Antwoord:</strong></p><span><span><input name='antwoord' type='text' size='4'></span></div><br><br>";
-
-  return html;
-};

### 2023-12-13 materials/test/rspan/global_variables.js

@@ -1,21 +1,0 @@
-// General variables
-var fullscreenmode = {
-  type: jsPsychFullscreen,
-  fullscreen_mode: true,
-  message: "<p>Het experiment zal verder gaan in <i>volledig scherm modus</i>.<br>We verzoeken je om in <i>volledig scherm modus</i> te blijven tijdens de gehele duur van het experiment.</p>",
-  button_label: "Ga verder"
-};
-
-var cursor_off = {
-    type: jsPsychCallFunction,
-    func: function() {
-        document.body.style.cursor= "none";
-    }
-};
-
-var cursor_on = {
-    type: jsPsychCallFunction,
-    func: function() {
-        document.body.style.cursor= "auto";
-    }
-};

### 2023-12-13 materials/test/rspan/plugins/rspan-recall.js

@@ -1,242 +1,0 @@
-var jsPsychRspanRecall = (function (jspsych) {
-  "use strict";
-
-  const info = {
-    name: "rspan-recall",
-    parameters: {
-      trial_duration: {
-        type: jspsych.ParameterType.INT,
-        pretty_name: 'Trial duration',
-        default: null,
-        description: 'How long to show the trial.'
-      },
-      size_cells: {
-        type: jspsych.ParameterType.INT,
-        pretty_name: 'size of cells',
-        default: 70,
-      },
-      correct_order: {
-        type:jspsych.ParameterType.INT,
-        default: undefined,
-        description: 'Record the correct array'
-      }
-    }
-  };
-
-  /**
-    * **PLUGIN-NAME**
-    *
-    * SHORT PLUGIN DESCRIPTION
-  *
-    * @author YOUR NAME
-  * @see {@link https://DOCUMENTATION_URL DOCUMENTATION LINK TEXT}
-  */
-    class RspanRecallPlugin {
-      constructor(jsPsych) {
-        this.jsPsych = jsPsych;
-      }
-      trial(display_element, trial) {
-        console.log(trial.correct_order)
-        // making matrix:
-        var grid = 3;
-        var recalledGrid = [];
-        //var correctLetters = trial.correct_order
-        var display = " "
-        var lengthOfChosenArrows = []
-
-
-        var leftOver = 12-setSize
-
-        var numbertobutton = {
-          "0": "&nwarr;",
-          "1": "&uarr;",
-          "2": "&nearr;",
-          "3": "&larr;",
-          "4": "?",
-          "5": "&rarr;",
-          "6": "&swarr;",
-          "7": "&darr;",
-          "8": "&searr;",
-        }
-
-        var imgtoarrow = {
-          "<img src='img/a1_short.png' height=300>": "&nwarr;",
-          "<img src='img/a2_short.png' height=300>": "&nearr;",
-          "<img src='img/a3_short.png' height=300>": "&searr;",
-          "<img src='img/a4_short.png' height=300>": "&swarr;",
-          "<img src='img/a5_short.png' height=300>": "&uarr;",
-          "<img src='img/a6_short.png' height=300>": "&rarr;",
-          "<img src='img/a7_short.png' height=300>": "&darr;",
-          "<img src='img/a8_short.png' height=300>": "&larr;"
-        }
-
-        var correctArrows = []
-        for (var i=0; i<trial.correct_order.length; i++){
-
[…]

### 2023-12-13 materials/test/rspan/plugins/updating-binding-number.js

@@ -1,157 +1,0 @@
-var jsPsychBindingUpdatingNumberGrid = (function (jspsych) {
-  "use strict";
-
-  const info = {
-    name: 'binding-updating-number-grid',
-    parameters: {
-      stimulus: {
-        type: jspsych.ParameterType.HTML_STRING,
-        default: null,
-        description: 'The HTML string to be displayed'
-      },
-      pos: {
-        type: jspsych.ParameterType.INT,
-        default: undefined,
-        description: "Position in the grid where the stimulus will be presented",
-      },
-      nBind: {
-        type: jspsych.ParameterType.INT,
-        default: undefined,
-        description: "Number of bindings steps",
-      },
-      nUpd: {
-        type: jspsych.ParameterType.INT,
-        default: undefined,
-        description: "Number of updating steps",
-      },
-      step_number: {
-        type: jspsych.ParameterType.INT,
-        default: undefined,
-        description: "Keeps track of the current step in the trial"
-      },
-      choices: {
-        type: jspsych.ParameterType.KEYCODE,
-        array: true,
-        default: "NO_KEYS",
-        description: 'The keys the subject is allowed to press to respond to the stimulus.'
-      },
-      trial_duration: {
-        type: jspsych.ParameterType.INT,
-        default: 1000,
-        description: 'How long to show the trial.'
-      },
-      response_ends_trial: {
-        type: jspsych.ParameterType.BOOL,
-        default: false,
-        description: 'If true, trial will end when subject makes a response.'
-      },
-    }
-  };
-
-  /**
-   * **UpdatingBindingGrid**
-   *
-   * Implements the updating and binding working memory task used in Wilhelm et al., 2013.
-   *
-   * @author Stefan Vermeent
-   * @see {@link https://github.com/stefanvermeent/liss_wm_profiles_2023 Github link}
-   */
-
-    class BindingUpdatingNumberGridPlugin {
-      constructor(jsPsych) {
-        this.jsPsych = jsPsych;
-      }
-      trial(display_element, trial) {
-
-        // ** CURRENT STIMULUS TO PRESENT
-      var current_letter = trial.stimulus
-
-      // SET LETTERS FOR THE DISPLAY
-      var letter1 = (trial.pos == 0)? current_letter : "";
-      var letter2 = (trial.pos == 1)? current_letter : "";
[…]

### 2023-12-13 materials/test/updating_binding/0_upd_bind_instructions.js

@@ -1,104 +1,0 @@
-//-------------------- Welcome
-var bind_upd_welcome = {
-  type: jsPsychInstructions,
-  pages: [
-    "U gaat nu beginnen aan het <b>Geheugen</b> spel!"
-  ],
-  show_clickable_nav: true,
-  allow_backward: true,
-  key_forward: -1,
-  key_backward: -1,
-  button_label_next: "verder",
-  data: {variable: 'welcome', task_version: "introduction"}
-};
-
-
-//-------------------- Instructions
-
-// If participants start with the number version, they will be introduced with the task using numbers instead of colors.
-var bind_upd_number_instructions = {
-  timeline: [
-    {
-      type: jsPsychInstructions,
-      pages: [
-        //Page 1
-        "<p style = 'text-align: center;'>" +
-          "In dit spel ziet u een vierkant met negen vakjes, zoals hieronder.<br><br>" +
-          "In de vakjes worden &#233;&#233;n voor &#233;&#233;n nummers getoond.<br>" +
-          "Uw taak is om te onthouden welke nummer <strong>als laatste</strong> wordt getoond in een bepaald vakje.<br><br><br>" +
-          "<img width = 250 src = img/instr01.png></img><br><br>",
-
-        //page 2
-        "<p style = 'text-align: center;'>"+
-          "Soms wordt in elk vakje maar &#233;&#233;n nummer getoond.<br><br>" +
-          "In het voorbeeld hieronder ziet u eerst een '<strong>6</strong>' midden-boven, dan een '<strong>2</strong>' linksonder, en tot slot een '<strong>9</strong>' linksmidden.<br>" +
-          "De nummers die u moet onthouden zijn dus '<strong>6</strong>', '<strong>2</strong>', en '<strong>9</strong>'.<br><br><br>" +
-          "<img width = 500 src = img/instr02.png></img><br><br>",
-
-        //page 3
-          "Soms worden meerdere nummers getoond in hetzelfde vakje.<br><br>" +
-          "In het voorbeeld hieronder ziet u eerst een '<strong>0</strong>' rechtsmidden, dan een '<strong>3</strong>' linksboven, en tot slot een '<strong>5</strong>' rechtsmidden.<br>" +
-          "In dit geval is de '<strong>0</strong>' vervangen door een '<strong>5</strong>'.<br>" +
[…]

### 2023-12-13 materials/test/updating_binding/1_upd_bind_number_procedure.js

@@ -1,251 +1,0 @@
-// Preload Instruction Images
-var preload_bind_upd = {
-  type: jsPsychPreload,
-  images: [
-     'img/instr01.png',
-     'img/instr02.png',
-     'img/instr03.png',
-     'img/instr04.png',
-  ]
-}
-
-
-var bind_upd_number_step = -1 // Keep track of where we are in the current trial flow
-var bind_upd_number_recall_step = 0 // Keep track of where we are in the recall flow
-var number_set = ['0', '1', '2', '3', '4', '5', '6', '7', '8', '9'] // Set of all stimuli
-var nCorrectBindUpdNumberRecall = 0
-var BindUpdPracticeCorrect = 0
-
-
-// Empty grid with fixation cross
-var fixation = {
-   on_start: function(){
-    bind_upd_number_step += 1
-  },
-  stimulus: "+",
-  type: jsPsychBindingUpdatingFixation,
-  trial_duration: 1000,
-  pos: 4,
-  nBind: -1,
-  nUpd: -1,
-  step_number: bind_upd_number_step
-};
-
-// Grid with memory item
-var bind_upd_number = {
-  type: jsPsychBindingUpdatingNumberGrid,
-  stimulus: function(){return jsPsych.timelineVariable("stim")[bind_upd_number_step]},
-  pos: function(){return jsPsych.timelineVariable("pos")[bind_upd_number_step]},
-  nBind: function(){return jsPsych.timelineVariable("nBind")},
-  nUpd: function(){return jsPsych.timelineVariable("nUpd")},
-  step_number: bind_upd_number_step,
-  trial_duration: function(){return 1500},
-  data: {
-    nBind: function(){return jsPsych.timelineVariable("nBind")},
-    nUpd: function(){return jsPsych.timelineVariable("nUpd")},
-    task: function(){return jsPsych.timelineVariable("task")}
-  }
-}
-
-// Empty grid in between memory items
-var bind_upd_number_empty = {
-  type: jsPsychBindingUpdatingNumberGrid,
-  stimulus: "",
-  pos: 0,
-  nBind: function(){return jsPsych.timelineVariable("nBind")},
-  nUpd: function(){return jsPsych.timelineVariable("nUpd")},
-  step_number: bind_upd_number_step,
-  trial_duration: function(){return 500},
-  data: {
-    stimulus: function(){return jsPsych.timelineVariable("stim")},
[…]

### 2023-12-13 materials/test/updating_binding/functions/utils.js

@@ -1,43 +1,0 @@
-function shuffleWithoutReplacement(arr, size) {
-    var shuffled = arr.slice(0), i = arr.length, temp, index;
-    while (i--) {
-        index = Math.floor((i + 1) * Math.random());
-        temp = shuffled[index];
-        shuffled[index] = shuffled[i];
-        shuffled[i] = temp;
-    }
-    return shuffled.slice(0, size);
-}
-
-function shuffleWithReplacement(arr, size, immediate_reps = false) {
-
-  i = size
-  last_bind_pos = arr.slice(-1)
-
-  out_array = []
-  previous_pos = null
-
-  if(size == 0) {
-    return []
-  }
-
-  while (i--) {
-    pos = arr[Math.floor(Math.random()*arr.length)];
-    if (immediate_reps == false) {
-      while (pos == previous_pos) {
-        pos = arr[Math.floor(Math.random()*arr.length)];
-      }
-
-      if (i == size-1) {
-        while (pos == last_bind_pos[0]) {
-          pos = arr[Math.floor(Math.random()*arr.length)];
-        }
-      }
-
-
-    }
-    previous_pos = pos
-    out_array.push(pos);
-  }
-  return out_array
-}

### 2023-12-13 materials/test/updating_binding/plugins/ospan-recall.js

@@ -1,205 +1,0 @@
-var jsPsychOspanRecall = (function (jspsych) {
-  "use strict";
-
-  const info = {
-    name: "ospan-recall",
-    parameters: {
-      trial_duration: {
-        type: jspsych.ParameterType.INT,
-        pretty_name: 'Trial duration',
-        default: null,
-        description: 'How long to show the trial.'
-      },
-      size_cells: {
-        type: jspsych.ParameterType.INT,
-        pretty_name: 'size of cells',
-        default: 70,
-      },
-      correct_order: {
-        type:jspsych.ParameterType.INT,
-        default: undefined,
-        description: 'Recored the correct array'
-      }
-    }
-  };
-
-  /**
-    * **PLUGIN-NAME**
-    *
-    * SHORT PLUGIN DESCRIPTION
-  *
-    * @author YOUR NAME
-  * @see {@link https://DOCUMENTATION_URL DOCUMENTATION LINK TEXT}
-  */
-    class OspanRecallPlugin {
-      constructor(jsPsych) {
-        this.jsPsych = jsPsych;
-      }
-      trial(display_element, trial) {
-
-        // making matrix:
-        var grid = 3;
-        var recalledGrid = [];
-        var correctLetters = trial.correct_order
-        var display = " "
-
-        var setSize = correctLetters.length
-        var leftOver = 12-setSize
-
-        var numbertobutton = {
-          "0": "F",
-          "1": "H",
-          "2": "J",
-          "3": "K",
-          "4": "L",
-          "5": "N",
-          "6": "P",
-          "7": "Q",
-          "8": "R",
-          "9": "S",
-          "10": "T",
-          "11": "V",
-        }
-
-        var matrix = [];
-        for (var i=0; i<4; i++){
-          var m1 = i;
-          for (var h=0; h<3; h++){
-            var m2 = h;
-            matrix.push([m1,m2])
-          }
-        };
-
-        var paper_size = [(3*(trial.size_cells+30)), ((4*(trial.size_cells+20))+100)];
-
-        var html = '<div id="jspsych-html-button-response-btngroup" style= "position: relative; width:' + paper_size[0] + 'px; height:' + paper_size[1] + 'px"></div>';
-        html += '<div style="width: 270"><div class="jspsych-btn-numpad" style="display: inline-block; margin:'+10+' '+2+'" id="jspsych-spatial-span-grid-button-12">?</div>';
[…]

### 2023-12-13 materials/test/updating_binding/plugins/rspan-recall.js

@@ -1,237 +1,0 @@
-var jsPsychRspanRecall = (function (jspsych) {
-  "use strict";
-
-  const info = {
-    name: "rspan-recall",
-    parameters: {
-      trial_duration: {
-        type: jspsych.ParameterType.INT,
-        pretty_name: 'Trial duration',
-        default: null,
-        description: 'How long to show the trial.'
-      },
-      size_cells: {
-        type: jspsych.ParameterType.INT,
-        pretty_name: 'size of cells',
-        default: 70,
-      },
-      correct_order: {
-        type:jspsych.ParameterType.INT,
-        default: undefined,
-        description: 'Recored the correct array'
-      }
-    }
-  };
-
-  /**
-    * **PLUGIN-NAME**
-    *
-    * SHORT PLUGIN DESCRIPTION
-  *
-    * @author YOUR NAME
-  * @see {@link https://DOCUMENTATION_URL DOCUMENTATION LINK TEXT}
-  */
-    class RspanRecallPlugin {
-      constructor(jsPsych) {
-        this.jsPsych = jsPsych;
-      }
-      trial(display_element, trial) {
-        console.log(trial.correct_order)
-        // making matrix:
-        var grid = 3;
-        var recalledGrid = [];
-        //var correctLetters = trial.correct_order
-        var display = " "
-        var lengthOfChosenArrows = []
-
-
-        var leftOver = 12-setSize
-
-        var numbertobutton = {
-          "0": "&nwarr;",
-          "1": "&uarr;",
-          "2": "&nearr;",
-          "3": "&larr;",
-          "4": "?",
-          "5": "&rarr;",
-          "6": "&swarr;",
-          "7": "&darr;",
-          "8": "&searr;",
-        }
-
-        var imgtoarrow = {
-          "<img src='rspan/img/a1_short.png' height=300>": "&nwarr;",
-          "<img src='rspan/img/a2_short.png' height=300>": "&nearr;",
-          "<img src='rspan/img/a3_short.png' height=300>": "&searr;",
-          "<img src='rspan/img/a4_short.png' height=300>": "&swarr;",
-          "<img src='rspan/img/a5_short.png' height=300>": "&uarr;",
-          "<img src='rspan/img/a6_short.png' height=300>": "&rarr;",
-          "<img src='rspan/img/a7_short.png' height=300>": "&darr;",
-          "<img src='rspan/img/a8_short.png' height=300>": "&larr;"
-        }
-
-        var correctArrows = []
[…]

### 2023-12-13 materials/test/updating_binding/plugins/updating-binding-fixation.js

@@ -1,165 +1,0 @@
-var jsPsychBindingUpdatingFixation = (function (jspsych) {
-  "use strict";
-
-  const info = {
-    name: 'binding-updating-fixation',
-    parameters: {
-      stimulus: {
-        type: jspsych.ParameterType.HTML_STRING,
-        default: null,
-        description: 'The HTML string to be displayed'
-      },
-      pos: {
-        type: jspsych.ParameterType.INT,
-        default: undefined,
-        description: "Position in the grid where the stimulus will be presented",
-      },
-      nBind: {
-        type: jspsych.ParameterType.INT,
-        default: undefined,
-        description: "Number of bindings steps",
-      },
-      nUpd: {
-        type: jspsych.ParameterType.INT,
-        default: undefined,
-        description: "Number of updating steps",
-      },
-      step_number: {
-        type: jspsych.ParameterType.INT,
-        default: undefined,
-        description: "Keeps track of the current step in the trial"
-      },
-      choices: {
-        type: jspsych.ParameterType.KEYCODE,
-        array: true,
-        default: "NO_KEYS",
-        description: 'The keys the subject is allowed to press to respond to the stimulus.'
-      },
-      trial_duration: {
-        type: jspsych.ParameterType.INT,
-        default: 1000,
-        description: 'How long to show the trial.'
-      },
-      response_ends_trial: {
-        type: jspsych.ParameterType.BOOL,
-        default: false,
-        description: 'If true, trial will end when subject makes a response.'
-      },
-    }
-  };
-
-  /**
-   * **UpdatingBindingGrid**
-   *
-   * Implements the updating and binding working memory task used in Wilhelm et al., 2013.
-   *
-   * @author Stefan Vermeent
-   * @see {@link https://github.com/stefanvermeent/liss_wm_profiles_2023 Github link}
-   */
-
-    class BindingUpdatingFixationPlugin {
-      constructor(jsPsych) {
-        this.jsPsych = jsPsych;
-      }
-      trial(display_element, trial) {
-
-        // ** CURRENT STIMULUS TO PRESENT
-      var current_stim = trial.stimulus;
-
-      // SET LETTERS FOR THE DISPLAY
-      var stim1 = (trial.pos == 0)? current_stim : "";
-      var stim2 = (trial.pos == 1)? current_stim : "";
-      var stim3 = (trial.pos == 2)? current_stim : "";
[…]

### 2023-12-13 materials/test/updating_binding/plugins/updating-binding-number-recall.js

@@ -1,251 +1,0 @@
-var jsPsychBindingUpdatingNumberRecall = (function (jspsych) {
-  "use strict";
-
-  const info = {
-    name: 'binding-updating-number-recall',
-    parameters: {
-      stimulus: {
-        type: jspsych.ParameterType.HTML_STRING,
-        default: "?",
-        description: 'The HTML string to be displayed',
-      },
-      recall_position: {
-        type: jspsych.ParameterType.INT,
-        default: undefined,
-        description: "Position in the grid where the stimulus will be presented",
-      },
-      preamble: {
-        type: jspsych.ParameterType.HTML_STRING,
-        default: "Welk nummer werd als laatst getoond op de plaats van het vraagteken?"
-      },
-      correct_number: {
-        type: jspsych.ParameterType.INT,
-        default: undefined
-      },
-      nBind: {
-        type: jspsych.ParameterType.INT,
-        default: undefined,
-        description: "Number of bindings steps",
-      },
-      nUpd: {
-        type: jspsych.ParameterType.INT,
-        default: undefined,
-        description: "Number of updating steps",
-      },
-      trial_duration: {
-        type: jspsych.ParameterType.INT,
-        default: 10000000
-      },
-      response_ends_trial: {
-        type: jspsych.ParameterType.BOOL,
-        default: true,
-        description: 'If true, trial will end when subject makes a response.'
-      },
-    }
-  };
-
-  /**
-   * **UpdatingBindingGrid**
-   *
-   * Implements the updating and binding working memory task used in Wilhelm et al., 2013.
-   *
-   * @author Stefan Vermeent
-   * @see {@link https://github.com/stefanvermeent/liss_wm_profiles_2023 Github link}
-   */
-
-    class BindingUpdatingNumberRecallPlugin {
-      constructor(jsPsych) {
-        this.jsPsych = jsPsych;
-      }
-      trial(display_element, trial) {
-
-        // ** CURRENT STIMULUS TO PRESENT
-      var current_stim = trial.stimulus;
-
-      // SET LETTERS FOR THE DISPLAY
-      var stim1 = (trial.recall_position == 0)? current_stim : "";
-      var stim2 = (trial.recall_position == 1)? current_stim : "";
-      var stim3 = (trial.recall_position == 2)? current_stim : "";
-      var stim4 = (trial.recall_position == 3)? current_stim : "";
-      var stim5 = (trial.recall_position == 4)? current_stim : "";
-      var stim6 = (trial.recall_position == 5)? current_stim : "";
[…]

### 2024-01-22 scripts/1_sim_pipeline/0_ivs.R
updates to simulated analysis pipeline

@@ -2,19 +2,22 @@
 
 library(tidyverse)
 
+source('scripts/0_custom_functions/codebooks.R')
+
 # 2. Functions ---------------------------------------------------------------
 
 # INCOME-TO-NEEDS RATIO FOR DUTCH HOUSEHOLDS
 # the information up until 2021 was derived from van den Brakel et al. 2023 (See main text).
 # The information for 2022 was derived from https://www.cbs.nl/nl-nl/nieuws/2023/45/laagste-armoederisico-in-45-jaar-door-energietoeslag#:~:text=De%20lage%2Dinkomensgrens%20voor%20een,de%20grens%201%20830%20euro.
 # At the moment of writing, there was no final data for 2023, so we used the same values as 2022.
-inr_conv_table <- tibble(
-  year            = rep(str_c(20, c("07", "08", "09", 10:23)), each = 8),
-  adult           = rep(c(1, 1, 1, 1, 2, 2, 2, 2), 17),
-  aantalki        = rep(c(0, 1, 2, 3, 0, 1, 2, 3), 17),
-  eq_fct_pre2018  = rep(c(1, 1.33, 1.51, 1.76, 1.37, 1.67, 1.88, 2.06), 17),
-  eq_fct_post2018 = rep(c(1, 1.32, 1.52, 1.73, 1.40, 1.69, 1.91, 2.09), 17)
-) |>
+inr_conv_table <-
+  tibble(
+    year            = rep(str_c(20, c("07", "08", "09", 10:23)), each = 8),
+    adult           = rep(c(1, 1, 1, 1, 2, 2, 2, 2), 17),
+    aantalki        = rep(c(0, 1, 2, 3, 0, 1, 2, 3), 17),
+    eq_fct_pre2018  = rep(c(1, 1.33, 1.51, 1.76, 1.37, 1.67, 1.88, 2.06), 17),
+    eq_fct_post2018 = rep(c(1, 1.32, 1.52, 1.73, 1.40, 1.69, 1.91, 2.09), 17)
+  ) |>
   mutate(
     # `base` is the low-income threshold for a single-person household without children.
     # All other threshold are derived from the base by multiplying it with the equivalence factor.
@@ -42,52 +45,74 @@ set.seed(4652757)
 ## 3.1 Material deprivation ----
 
 # Note: The variable *numbers* are identical to those in the LISS archive. However, the prefixes are standardized here as they are variable across waves.
-core_income <- tibble(
[…]

### 2024-01-22 scripts/1_sim_pipeline/1_dvs.R

@@ -1,0 +1,38 @@
+# 1. Libraries ----------------------------------------------------------------
+
+library(tidyverse)
+
+
+# 2. Stage 1: Simulate data for computational reproducibility ----------------
+
+dvs_sim <-
+  tibble(
+    nomem_encr = 1:800 |> as.character(),
+    rspan_cap     = rnorm(800, 0.75, 0.15) %>% ifelse(. > 1, 1, .),
+    rspan_sec_acc = rnorm(800, 0.88, 0.11) %>% ifelse(. > 1, 1, .),
+    rspan_sec_rt  = rnorm(800, 0.88, 0.11) %>% ifelse(. < 0, 300, .),
+    ospan_cap     = rnorm(800, 0.75, 0.15) %>% ifelse(. > 1, 1, .),
+    ospan_sec_acc = rnorm(800, 0.88, 0.11) %>% ifelse(. > 1, 1, .),
+    ospan_sec_rt  = rnorm(800, 1500, 476)  %>% ifelse(. < 0, 300, .),
+    bu_bind_acc   = rnorm(800, 0.88, 0.11) %>% ifelse(. > 1, 1, .),
+    bu_upd_acc    = rnorm(800, 0.70, 0.11) %>% ifelse(. > 1, 1, .),
+    brows_int     = sample(c(0, 1), size = 800, replace = T, prob = c(0.95, 0.05))
+  ) |>
+  sjlabelled::var_labels(
+    rspan_cap     = "Rspan Capacity. Average number of arrows recalled in the correct serial position on each trial (Proportion).",
+    rspan_sec_acc = "Rspan Accuracy on secondary rotation task. Average accuracy across trials (Proportion)",
+    rspan_sec_rt  = "Rspan RT on secondary rotation task. Average RT across trials (RT in miliseconds)",
+    ospan_cap     = "Ospan Capacity. Average number of letters recalled in the correct serial position on each trial. (Proportion)",
+    ospan_sec_acc = "Ospan Accuracy on secondary math task. Average accuracy across trials (Proportion)",
+    ospan_sec_rt  = "Ospan RT on secondary math task. Average RT across trials (RT in miliseconds)",
+    bu_bind_acc   = "Binding-Updating, Binding accuracy. Average amount of numbers recalled in the correct location across trials that did not involve an updating component (Proportion)",
[…]

### 2024-01-22 scripts/1_sim_pipeline/2_exclusions.R

@@ -1,0 +1,27 @@
+# 1. Libraries ---------------------------------------------------------------
+
+library(tidyverse)
+
+source('scripts/0_custom_functions/codebooks.R')
+source('scripts/0_custom_functions/binom_test.R')
+
+# 2. Load data ---------------------------------------------------------------
+
+load("data/dvs_sim.RData")
+load("data/ivs_sim.RData")
+
+
+# 3. Exclusions -----------------------------------------------------------
+
+full_data <- full_join(ivs_sim, dvs_sim) |>
+  mutate(
+    ex_brows_int       = ifelse(brows_int == 0, FALSE, TRUE),
+    ex_acc_ospan_sec   = ifelse(ospan_sec_acc*100 < gbinom(45, 0.50), TRUE, FALSE),
+    ex_acc_rspan_sec   = ifelse(rspan_sec_acc*100 < gbinom(45, 0.50), TRUE, FALSE),
+  )
+
+clean_data <- full_data |>
+  filter(ex_brows_int == FALSE & ex_acc_ospan_sec == FALSE & ex_acc_rspan_sec == FALSE)
+
+save(full_data, clean_data, file = "data/full_data.Rdata")
+

### 2024-01-22 scripts/1_sim_pipeline/3_sem.R

@@ -1,66 +1,0 @@
-
-# 1. Libraries and Data ------------------------------------------------------
-
-library(lavaan)
-library(tidyverse)
-
-# Simulated dataset with standardized variables
-data_clean <- tibble(
-  id             = 1:800,
-  family_id  = 1:800,
-  ospan          = rnorm(800, 0, 1), # Operation Span
-  rspan          = rnorm(800, 0, 1), # Rotation Span
-  upd            = rnorm(800, 0, 1), # Updating
-  bind           = rnorm(800, 0, 1), # Binding
-  dep            = rnorm(800, 0, 1), # Material deprivation composite
-  threat         = rnorm(800, 0, 1), # Threat composite
-  age            = rnorm(800, 0, 1), # age
-  noise          = rnorm(800, 0, 1), # Environmental noise
-  interrupt      = sample(c(0,1), size = 800, prob = c(0.95, 0.05), replace = TRUE), # Was interrupted
-  leave_computer = sample(c(0,1), size = 800, prob = c(0.95, 0.05), replace = TRUE)  # Left computer during tasks
-)
-
-# 2. Specify Clustering Design -----------------------------------------------
-
-# Some children are clustered within families.
-# We account for this using the lavaan.survey package
-
-cluster_design <- survey::svydesign(ids=~family_id, prob=~1, data = data_clean)
-
-
-# 3. Model specification(s) --------------------------------------------------
-
-model1 <-
-  '
-  # Measurement Model
-
-  ## Latent factors
-  wmc_l      =~ 1*ospan + rspan + bind + upd
-  upd_l      =~ 1*upd
-
-  ## Covariances
-  upd   ~~ 0*upd
-
-  dep   ~~ threat
-
-  # Structural Model
-  wmc_l ~ dep + threat + age + interrupt + leave_computer
-  upd_l ~ dep + threat + age + interrupt + leave_computer
-'
-
-
-sem_fit1 <- lavaan::sem(
-  model = model1,
-  data = data_clean,
-  orthogonal = TRUE,
-  auto.cov.lv.x = FALSE,
-  auto.cov.y = FALSE,
-  std.lv = FALSE,
-  fixed.x = FALSE
-)
-
-sem_fit1_cluster <- lavaan.survey::lavaan.survey(lavaan.fit = sem_fit1, survey.design = cluster_design, estimator = "ML")
-
-summary(sem_fit1_cluster, fit.measures = TRUE, standardized = TRUE)
-
-

### 2024-01-22 scripts/1_sim_pipeline/4_sem.R

@@ -1,0 +1,70 @@
+
+# 1. Libraries and Data ------------------------------------------------------
+
+library(lavaan)
+library(tidyverse)
+library(lavaan.survey)
+
+# 2. Load data ------------------------------------------------------------
+
+load("data/full_data.RData")
+
+# 2. Specify Clustering Design -----------------------------------------------
+
+# Some children are clustered within families.
+# We account for this using the lavaan.survey package
+
+cluster_design <- survey::svydesign(ids=~nohouse_encr, prob=~1, data = data_clean)
+
+
+# 3. Model specification(s) --------------------------------------------------
+
+## 3.1 Measurement Model ----
+mod_meas <-
+  '
+  # Measurement Model
+
+  ## Latent factors
+  wmc_l      =~ 1*bu_bind_acc + rspan_cap + ospan_cap + bu_upd_acc
+  upd_l      =~ 1*bu_upd_acc
+
+  ## Covariances
+  bu_upd_acc ~~ 0*bu_upd_acc
+'
+
+
+fit_meas1 <- lavaan::sem(
+  model = mod_meas,
+  data = data_clean,
+  estimator = "MLR", # Robust SEs in case of non-normal data
+  missing = "ML",    # FIML for missing data
+  orthogonal = TRUE,
+  auto.cov.lv.x = FALSE,
+  auto.cov.y = FALSE,
+  std.lv = FALSE,
+  fixed.x = FALSE
+)
+
+fit_meas1_cluster <- lavaan.survey::lavaan.survey(lavaan.fit = fit_meas1, survey.design = cluster_design, estimator = "MLR")
+summary(fit_meas1_cluster, fit.measures = TRUE, standardized = TRUE)
+
+## 3.2 Full Structural model
+
+mod_full <-
+  '
+  # Measurement Model
+
+  ## Latent factors
+  wmc_l      =~ 1*bu_bind_acc + rspan_cap + ospan_cap + bu_upd_acc
+  upd_l      =~ 1*bu_upd_acc
+
+  ## Covariances
+  bu_upd_acc ~~ 0*bu_upd_acc
+
+  mat_dep        ~~ threat
+
+  # Structural Model
+  wmc_l ~ mat_dep + unp + threat + age + interrupt + leave
+  upd_l ~ mat_dep + unp + threat + age + interrupt + leave
+'
+

### 2024-01-22 scripts/1_simulations/unpredictability_measures.R
remove old script

@@ -1,57 +1,0 @@
-
-# Libraries ---------------------------------------------------------------
-
-library(tidyverse)
-
-
-# Functions ---------------------------------------------------------------
-
-# Compute rowwise SDs
-RowSD <- function(x) {
-  sqrt(rowSums((x - rowMeans(x))^2)/(dim(x)[2] - 1))
-}
-
-# Create waves of data in line with within-subject specified Means and SDs
-create_wave <- function(data, names) {
-  data <- map(names, function(x) {
-    data <- data |>
-      rowwise() |>
-      mutate("{x}" := rnorm(1, mean, sd)) |>
-      ungroup()
-  }) |>
-    reduce(left_join)
-}
-
-
-# Simulate data -----------------------------------------------------------
-
-set.seed(3554)
-
-data <-
-  faux::rnorm_multi(n = 1000, vars = 2, mu = 0, sd = 1, r = -0.7, varnames = c("mean", "sd")) |>
-  as_tibble() |>
-  mutate(
-    id = 1:1000,
-    sd = (sd+abs(min(sd)))/2,
-    mean = (mean+abs(min(mean)))
-    ) %>% # make sure SDs are positive
-  create_wave(data = ., names = str_c("wave", 1:100)) |> # Simulate 100 waves
-  mutate(mat_dep = across(starts_with('wave')) %>% rowMeans) |> # Compute within-person mean level material deprivation
-  mutate(unpred  = across(starts_with("wave")) %>% RowSD) |>  # compute within-person SD of material deprivation over time
-  pivot_longer(starts_with('wave'), names_to = "wave", values_to = "value") |>
-  group_by(id) |>
-  mutate(prop_change = (value - lag(value, n = 1)) / lag(value, n = 1)) |>
-  summarise(
-    mat_dep = mean(value),
-    unpred = sd(value),
-    prop_change = mean(abs(prop_change), na.rm = T) # Unpredictability option 1: Average proportion of change
-  ) |>
-  mutate(sd_corrected = unpred / mat_dep) # Unpredictability option 2: within-subject SD divided by within-subject mean
-
-
-ggplot(data, aes(prop_change, sd_corrected)) +
-  geom_point()
-
-ggplot(data, aes(prop_change, unpred)) +
-  geom_point()
-

### 2024-01-22 README.Rmd
initiate README

@@
 ---
-title: "Working memory performance in adverse environments: Enhanced, impaired, or intact?"
-output: github_document
+title: ""
+output: 
+  html_document:
+    template: assets/template.html
+    css: assets/style.css
+    keep_md: true
 editor_options: 
@@
 library(tidyverse)
+library(glue)
 devtools::load_all("D:/repositories/projectlog")
@@
 A Stage 1 snapshot of the project has been submitted to [Peer Community in Registered Reports (PCI-RR)](https://rr.peercommunityin.org/).
+
+Do you want to download or clone the materials for this project? Go to [https://github.com/stefanvermeent/liss_wm_profiles_2023](https://github.com/stefanvermeent/liss_wm_profiles_2023).
+
+## Directory Structure {#structure}
+
+The names of each folder are intended to be self-explanatory.
+There are eight top-level folders to organize the inputs and outputs of this project:
+
+1.  [`Manuscript`](https://stefanvermeent.github.io/abcd_ddm/registered_report/README.html): The Registered Report written in Quarto.
+2.  [`Supplement`](https://stefanvermeent.github.io/abcd_ddm/supplement/README.html): a supplemental text (to be submitted with the manuscript) documenting all secondary analyses in detail.
+3.  [`Scripts`](https://stefanvermeent.github.io/abcd_ddm/scripts/README.html): R-scripts that read, analyze, and produce all analysis objects.
+4.
+[`Data`](https://stefanvermeent.github.io/abcd_ddm/data/README.html): Folder in which real LISS data can be placed to make the analyses fully reproducible.
+Note that we cannot openly share the raw data on the open repository.
+The folder contains synthetic data to facilitate computational reproducibility in the absence of LISS access.
+5.  [`Analysis Objects`](https://stefanvermeent.github.io/abcd_ddm/analysis_objects/README.html): Folder containing all analysis objects.
+6.  [`Codebooks`](https://stefanvermeent.github.io/abcd_ddm/codebooks/README.html): lists of variable names, labels, and value labels (where applicable).
+
+Click on each of the folders to get more details.
+
+## Overview of project milestones
+
+Below is an overview of all the project milestones, such as first-time data access, submissions, and revisions.
+Data access events were automatically captured using custom code, which over the course of this project was collected in the R package `projectlog` [https://stefanvermeent.github.io/projectlog/](https://stefanvermeent.github.io/projectlog/).
+For more information about how tracking worked, Go to the [Open science workflow tab](https://stefanvermeent.github.io/abcd_ddm/opensci_workflow/README.html).
+
+```{r results ='asis', echo = FALSE}
+x <- gert::git_log(max = 10000) |> 
+  select(commit, author, time, message) |>
+  mutate(
+    milestone = map_chr(milestone, function(milestone) {
+      milestone |> 
+        str_remove("\\[\\[") |>
+        str_replace("_", " ") |> 
+        str_to_title()
+    }),
+    data_hash = str_remove(data_hash, "^object_hash ")
+  ) |> 
+  select(time, commit, milestone, description, code, data_hash) |> 
+  mutate(
+    code = str_replace_all(code, "\\|\\>", "\\|\\>\\\n"),
+    link_to_git = paste0(
+      "<anonymized repository>"
+     # gert::git_remote_list()$url |> str_remove(string = _, pattern = "\\.git$"), "/tree/", commit)
+    )
+  ) |>
+  arrange(time) |> 
+  glue_data("  - **[{time}]({link_to_git}): {description}**
+        - **Milestone:** {milestone}
+        - **Data MD5 hash**: {data_hash}
+        - [Link to code snippet](https://anonymous.4open.science/r/anon-255D/.gitlog/{commit}.R)
+        \n
+    ")
+x
+
+```

### 2024-01-22 README.Rmd
remove milestones for now

@@
 For more information about how tracking worked, Go to the [Open science workflow tab](https://stefanvermeent.github.io/abcd_ddm/opensci_workflow/README.html).
-
-```{r results ='asis', echo = FALSE}
-x <- gert::git_log(max = 10000) |> 
-  select(commit, author, time, message) |>
-  mutate(
-    milestone = map_chr(milestone, function(milestone) {
-      milestone |> 
-        str_remove("\\[\\[") |>
-        str_replace("_", " ") |> 
-        str_to_title()
-    }),
-    data_hash = str_remove(data_hash, "^object_hash ")
-  ) |> 
-  select(time, commit, milestone, description, code, data_hash) |> 
-  mutate(
-    code = str_replace_all(code, "\\|\\>", "\\|\\>\\\n"),
-    link_to_git = paste0(
-      "<anonymized repository>"
-     # gert::git_remote_list()$url |> str_remove(string = _, pattern = "\\.git$"), "/tree/", commit)
-    )
-  ) |>
-  arrange(time) |> 
-  glue_data("  - **[{time}]({link_to_git}): {description}**
-        - **Milestone:** {milestone}
-        - **Data MD5 hash**: {data_hash}
-        - [Link to code snippet](https://anonymous.4open.science/r/anon-255D/.gitlog/{commit}.R)
-        \n
-    ")
-x
-
-```

### 2024-01-22 README.Rmd
update README files

@@
 For more information about how tracking worked, Go to the [Open science workflow tab](https://stefanvermeent.github.io/abcd_ddm/opensci_workflow/README.html).
+
+```{r results ='asis', echo = FALSE}
+gert::git_tag_list() |> 
+  mutate(
+    tags = pmap(list(name = name, ref = ref, commit = commit), function(name, ref, commit) {
+      gert::git_commit_info(commit) |> 
+        as_tibble() |> 
+        select(commit = id, author, time, message) 
+    })
+  ) |> 
+  select(-commit, -ref) |> 
+  unnest(tags) |> 
+  separate(message, into = c("description", "hash", "code"), sep = "\\n") |> 
+  mutate(
+    name = map_chr(name, function(name) {
+      name |> 
+        str_remove("\\[\\[") |>
+        str_remove("\\d*$") |> 
+        str_replace("_", " ") |> 
+        str_to_title()
+    })
+  ) |> 
+  mutate(
+    code = str_remove_all(code, "^code\\s") |> str_replace_all(string = _, "\\|\\>", "\\|\\>\\\n"),
+    link_to_git = paste0(
+       gert::git_remote_list()$url |> str_remove(string = _, pattern = "\\.git$"), "/tree/", commit)
+    ) |>
+  arrange(desc(time)) |> 
+  glue_data("  - **[{time}]({link_to_git}): {description}**
+        - **Milestone:** {name}
+        - **Data MD5 hash**: {hash}
+        - [Link to code snippet](https://https://github.com/StefanVermeent/abcd_ddm/tree/main/.projectlog/{commit}.R)
+        \n
+    ")
+```

### 2024-01-22 README.Rmd
updated top-level READMEs

--- 2024-01-22 README.Rmd
@@
 For more information about how tracking worked, Go to the [Open science workflow tab](https://stefanvermeent.github.io/abcd_ddm/opensci_workflow/README.html).
+
+```{r results ='asis', echo = FALSE}
+gert::git_tag_list() |> 
+  mutate(
+    tags = pmap(list(name = name, ref = ref, commit = commit), function(name, ref, commit) {
+      gert::git_commit_info(commit) |> 
+        as_tibble() |> 
+        select(commit = id, author, time, message) 
+    })
+  ) |> 
+  select(-commit, -ref) |> 
+  unnest(tags) |> 
+  separate(message, into = c("description", "hash", "code"), sep = "\\n") |> 
+  mutate(
+    name = map_chr(name, function(name) {
+      name |> 
+        str_remove("\\[\\[") |>
+        str_remove("\\d*$") |> 
+        str_replace("_", " ") |> 
+        str_to_title()
+    })
+  ) |> 
+  mutate(
+    code = str_remove_all(code, "^code\\s") |> str_replace_all(string = _, "\\|\\>", "\\|\\>\\\n"),
+    link_to_git = paste0(
+       gert::git_remote_list()$url |> str_remove(string = _, pattern = "\\.git$"), "/tree/", commit)
+    ) |>
+  arrange(desc(time)) |> 
+  glue_data("  - **[{time}]({link_to_git}): {description}**
+        - **Milestone:** {name}
+        - **Data MD5 hash**: {hash}
+        - [Link to code snippet](https://https://github.com/StefanVermeent/liss_wm_profiles_2023/tree/main/.projectlog/{commit}.R)
+        \n
+    ")
+```

### 2024-01-22 analysis_objects/README.Rmd
READMEs for the various folders

---
title: ""
output: 
  html_document:
    template: ../assets/template.html
    css: ../assets/style.css
editor_options: 
  markdown: 
    wrap: sentence
---

# Analysis objects

This folder contains all analysis objects, i.e., the outputs of the various data processing scripts.
Similar to the data, analysis objects containing individual-level data cannot be shared openly on GitHub.
These analysis objects can be created either by using the simulated data files or by requesting access to the LISS data (see [how to reproduce this repository](#reproduce)).

### 2024-01-22 codebooks/README.Rmd
READMEs for the various folders

---
title: ""
output: 
  html_document:
    template: ../assets/template.html
    css: ../assets/style.css
editor_options: 
  markdown: 
    wrap: sentence
---

# Codebooks

Below you can find the codebooks of the data sets used for the primary analyses. To download them, go [here](https://github.com/stefanvermeent/liss_wm_profiles_2023/tree/main/codebooks).

## ddm_data

```{r, include = FALSE, echo = FALSE, warning = FALSE, message=TRUE}
library(readr)
ddm_data <- read_csv("ddm_data.csv")
```

```{r}
ddm_data
```

## iv_data

```{r, include = FALSE, echo = FALSE, warning = FALSE, message=TRUE}
iv_data <- read_csv("iv_data.csv")
```

```{r}
iv_data
```

## test_set

```{r, include = FALSE, echo = FALSE, warning = FALSE, message=TRUE}
test_set <- read_csv("test_set.csv")
```

```{r}
test_set
```

### 2024-01-22 scripts/README.Rmd
READMEs for the various folders

---
title: ""
output: 
  html_document:
    template: ../assets/template.html
    css: ../assets/style.css
editor_options: 
  markdown: 
    wrap: sentence
---

# Data Processing Scripts

There are four types of R-scripts in this repository, each with a separate folder.

-   [`o_custom_functions/`](https://github.com/stefanvermeent/liss_wm_profiles_2023/tree/main/scripts/custom_functions): Custom R-functions written for this project.
-   [`0_pilot/`](https://github.com/stefanvermeent/liss_wm_profiles_2023/tree/main/scripts/0_pilot): Scripts analyzing the Pilot data.
-   [`1_sim_pipeline/`](https://github.com/stefanvermeent/liss_wm_profiles_2023/tree/main/scripts/1_sim_pipeline): A full analysis pipeline, written before accessing/analyzing the data and based on simulated data.

Each script takes an input(s) and produces output(s).
All outputs are stored in [analysis_objects/](https://github.com/stefanvermeent/liss_wm_profiles_2023/tree/main/analysis_objects).
The tables below provides an overview of the inputs and outputs of each script.

```{r include = FALSE}
library(tidyverse)
```

```{r echo = FALSE}

file_list <- list.files(pattern = ".R$", full.names = T, recursive = T) %>% 
  map_df(function(x){
    script_text <- 
      read_lines(x) %>% 
      str_subset("load.*(RData)|read.*(csv)|save.image\\(.*\\)|file = .*RData|filename = .*pdf|write_csv.*csv") %>% 
      str_remove_all("glue|\\{.*\\}| |\"|\\(|\\)|,") %>% 
      str_c(collapse = " ") %>% 
      str_replace_all("load","input: ") %>% 
      str_replace("save|write_csv|save.imagefile = |save.image|filename = |file = ","~output: ") %>% 
      str_replace_all("save|write_csv|save.imagefile = |save.image|filename = |file = ","output: ") %>% 
      str_split("~", simplify = T)

    tibble(
      dir       = x %>% str_remove_all("\\./") %>% str_remove_all("/.*") %>% str_remove("^\\d-"),
      script    = x %>% str_remove_all(".*/"),
      raw       = script_text %>% str_squish() %>% str_c(collapse = ", "),
      n_input   = str_count(script_text[1], "input"),
      input1    = script_text[1],
      input2    = input1 %>% str_remove_all("input: ") %>% str_trim(),
      input     = input2 %>% str_split(" ", simplify = T) %>% str_remove_all(".*/") %>% str_c(collapse = "<br> ") %>% ifelse(. == "<br> ","",.),
      n_output  = str_count(script_text[2], "output") %>% ifelse(is.na(.), 0, .),
      output1   = script_text[2],
      output2   = output1 %>% str_remove_all("output: ") %>% str_trim(),
      output    = output2 %>% str_split(" ", simplify = T) %>% str_remove_all(".*/") %>% str_c(collapse = "<br> ") %>%  ifelse(. == "<br> NA","",.)
    ) |> 
      mutate(
        script = paste0("[", script, "]", "(", "https://github.com/stefanvermeent/liss_wm_profiles_2023/tree/main", "/scripts/", dir, "/", script, ")")
      )
  })

```

### Simulations

```{r, echo = FALSE, results='asis'}
pilot <- 
  file_list %>% 
  select(dir, script, input, output) %>% 
  filter(dir == "0_pilot") %>% 
  select(-dir) 
cat(
  "|script|input|output|",
  "|------|-----|------|",
  glue::glue_data(
    pilot,
    "|{script}|{input}|{output}|"
  ),
  sep = "\n"
)
```

### Data Prep

```{r, echo = FALSE, results='asis'}
sim_pipeline <- 
  file_list %>% 
  select(dir, script, input, output) %>% 
  filter(dir == "1_sim_pipeline") %>% 
  select(-dir) 
cat(
  "|script|input|output|",
  "|------|-----|------|",
  glue::glue_data(
    sim_pipeline,
    "|{script}|{input}|{output}|"
  ),
  sep = "\n"
)
```

### 2024-01-22 scripts/0_custom_functions/binom_test.R

@@ -1,0 +1,16 @@
+gbinom <- function(n, p, low=0, high=n,scale = F) {
+
+  sd <- sqrt(n * p * (1 - p))
+  if(scale && (n > 10)) {
+    low <- max(0, round(n * p - 4 * sd))
+    high <- min(n, round(n * p + 4 * sd))
+  }
+  values <- low:high
+  probs <- dbinom(values, n, p)
+
+  max_chance_performance <- max(values[which(probs > 0.025)])
+
+
+
+  return((max_chance_performance/n)*100)
+}

### 2024-01-22 scripts/0_custom_functions/codebooks.R

@@ -1,0 +1,53 @@
+extract_attrs <- function(data, attr_type){
+  purrr::map_chr(data, function(x){
+
+    # For variable labels
+    if(attr_type == "label"){
+      extracted_attr <- attr(x, which = attr_type)
+      # If it's null, make it an empty string
+      if (is.null(extracted_attr)) {
+        extracted_attr <- ""
+      }
+      extracted_attr
+    }
+
+    # For variable values
+    if(attr_type == "labels"){
+      extracted_attr <- attr(x, which = attr_type)
+      if (is.null(extracted_attr)) {
+        extracted_attr <- ""
+      }
+      else {
+        extracted_attr <- paste(names(extracted_attr),"=", extracted_attr, collapse = ", ")
+      }
+      extracted_attr
+    }
+
+    extracted_attr
+  })
+}
+
+create_codebook <- function (data) {
+  # create a lookup table for a labelled dataset
+  lookup_data <-
+    dplyr::tibble(
+      Variable = names(data),
+      Label    = extract_attrs(data, "label"),
+      Values   = extract_attrs(data, "labels")
+    )
+
+  codebook <- lookup_data
+  no_labels <- codebook %>% dplyr::filter(Label == "")
+  no_values <- codebook %>% dplyr::filter(Values == "")
+  if (nrow(no_labels) > 0) {
+    message("---------The following variables have no labels:---------\n",
+            paste(no_labels %>% dplyr::pull(Variable), collapse = "\n"))
+  }
+  if (nrow(no_values) > 0) {
+    message("---------The following variables have no value labels:---------\n",
+            paste(no_values %>% dplyr::pull(Variable), collapse = "\n"))
+  }
+  codebook
+
+
+}

### 2024-01-22 scripts/0_custom_functions/corr_plot.R

@@ -1,0 +1,79 @@
+corr_table <- function (data, sample_size = T, use = "pairwise", method = "pearson", 
+                        stats = NULL, c.names = names(data), change = F, numbered = F, 
+                        flagged = T) {
+  
+  my_ifelse <- function(...) {
+    suppressWarnings(ifelse(...))
+  }
+  
+  descriptives <- psych::describe(data) %>% as.data.frame() %>% round(2) 
+  corr_data <- psych::corr.test(x = data, use = use, method = method)
+  ns <- corr_data$n
+  rs <- corr_data$r %>% round(2) %>% formatC(digits = 2, width = 3, flag = '0', format = 'f')
+  
+  descriptives <- 
+    descriptives %>% 
+    select(-vars) %>% 
+    rownames_to_column(var = "var") %>% 
+    mutate(across(3:13, ~ formatC(.x, digits = 2, width = 3, flag = '0', format = 'f'))) %>% 
+    pivot_longer(cols = c(-var), names_to = "stat", values_to = "value", values_transform = list(value = as.character)) %>% 
+    pivot_wider(names_from = var, values_from = value)
+  
+  if (!is.null(stats)) {
+    descriptives <- 
+      descriptives %>% 
+      filter(stat %in% stats) %>% 
+      slice(match(stats, stat)) %>% 
+      mutate(stat = str_to_title(stat), stat = ifelse(stat=="Sd","SD",stat))
+  }
+  if (flagged) {
+    ps <- 
+      corr_data$p %>% 
+      as.data.frame() %>% 
+      mutate(across(.cols = everything(), .fns = ~ my_ifelse(as.numeric(.x) <  .01, "", as.character(.x)))) %>% 
+      mutate(across(.cols = everything(), .fns = ~ my_ifelse(as.numeric(.x) <= .05 & !is.na(as.numeric(.x)), "", .x))) %>% 
+      mutate(across(.cols = everything(), .fns = ~ my_ifelse(as.numeric(.x) >  .05 & !is.na(as.numeric(.x)), " ", .x))) %>% 
+      as.matrix()
+    
+    flagged.rs <- 
+      paste(rs, ps, sep = "") %>% 
+      matrix(nrow = nrow(rs), ncol = ncol(rs))
+    
+    rs[lower.tri(rs)] <- flagged.rs[lower.tri(flagged.rs)]
+  }
+  
+  if (sample_size) {
[…]

### 2024-01-23 materials/README.Rmd
README file for materials

---
title: ""
output: 
  html_document:
    template: ../assets/template.html
    css: ../assets/style.css
editor_options: 
  markdown: 
    wrap: sentence
---

## Materials {#materials}

This folder contains all the materials of the cognitive tasks.
All tasks were programmed in [JsPsych 7.3.2](https://www.jspsych.org/7.3/).

1.  [`0_pilot/`](https://github.com/StefanVermeent/liss_wm_profiles_2023/tree/master/materials/0_pilot): Folder containing the pilot materials for the Operation Span Task and the Binding-Updating Task.
2.  [`0_pilot2/`](https://github.com/StefanVermeent/liss_wm_profiles_2023/tree/master/materials/0_pilot2): Folder containing the pilot materials for the Rotation Span Task.
3.  [`1_liss/`](https://github.com/StefanVermeent/liss_wm_profiles_2023/tree/master/materials/1_liss): Folder containing the final materials of all three tasks as used in the main study.

In each of these study folders, you will find several files and folders:

1.  `css/`. Handles the visual appearance of the cognitive tasks.
2.  `functions/`. Several helper functions to create stimuli and/or timelines.
3.  `plugins/`. Custom JsPsych plugins to handle trial behavior that is not included in the default JsPsych library.
4.  `global_variables.js`. Global functions for control across studies, e.g., full-screen mode, turning off the cursor.
5.  `index.html`. Script which combines the code of all three tasks, handles counterbalancing, and initiates the tasks.

Within each study folder, each task also has it's own folder (`ospan`, `rspan`, or `updating_binding`.)
In each folder, you will find two files:
`0_[task]_instructions.js` contains the Javascript code for all task instructions.
`1_[task]_procedure.js` contains Javascript code defining the stimuli and trials for both practice and test blocks.

**Run tasks locally.** To run the tasks locally on your device, navigate to one of the three main folders as listed above, and open the `index.html` file in a browser. 
This will automatically start the cognitive tasks (in counterbalanced order).

**Run tasks online.** The tasks are hosted online at [https://liss-wm-profiles-2023-main.netlify.app/](https://liss-wm-profiles-2023-main.netlify.app/).
Navigating to this url will automatically start the tasks.

### 2024-01-23 scripts/0_pilot/01_create_scales.R
tidy up scripts for website

@@ -245,7 +245,7 @@ ospan_data_math_clean <-
   filter(variable == "math") |>
   select(id, prolific_pid, rt, block, counterbalance, correct) |>
   group_by(prolific_pid) |>
-  summarise(ospan_math_acc = sum(correct) / n())
+  summarise(ospan_sec_acc = sum(correct) / n())
 
 ## Rspan ----
 
@@ -279,13 +279,15 @@ rspan_data_rotation_clean <-
   filter(variable == "rotation") |>
   select(id, prolific_pid, rt, block, correct) |>
   group_by(prolific_pid) |>
-  summarise(rspan_rotation_acc = sum(correct) / n())
+  summarise(rspan_sec_acc = sum(correct) / n())
 
 
 # Combine All Data --------------------------------------------------------
 
-pilot_data_clean <-
+pilot_full_data <-
   left_join(ospan_data_letter_clean, rspan_data_arrow_clean) |>
+  left_join(ospan_data_math_clean) |>
+  left_join(rspan_data_rotation_clean) |>
   left_join(
     bind_upd_color_clean |>
       pivot_wider(names_from = "version", values_from = 'color_acc') |>
@@ -311,103 +313,4 @@ pilot_data_clean <-
   mutate(id = 1:n()) |>
   select(-prolific_pid)
 
-
-# Partial correlations, controlling for age
-pilot_data_clean |>
-  select(-id) |>
-  cor(use = "pairwise.complete.obs") |>
-  partial.r(x = c(1,2,3,4,5,6,7,8,9), y = c(10))
-
-
-
-pilot_data_clean |>
-  pivot_longer(c(ends_with('cap'), starts_with("binding"), starts_with("updating")), names_to = 'task', values_to = 'score') |>
-  pivot_longer(c(pcunp_mean, vio_comp, ses_subj_comp), names_to = "adversity", values_to = "adv_value") |>
-  ggplot(aes(adv_value, score)) +
-  geom_point() +
-  geom_smooth(method = "lm") +
-  facet_grid(task~adversity)
-
-
-# Timestamps --------------------------------------------------------------
-
-overall_duration <-
-  pilot_data |>
-  select(prolific_pid, timestamp_tasks) |>
-  mutate(timestamp_tasks = timestamp_tasks / 60)
-
-
-task_duration <- list(bind_upd_color_data, bind_upd_color_practice, bind_upd_number_data, bind_upd_number_practice, ospan_data, ospan_practice) |>
-  map_dfr(function(x){
-    x |>
[…]

### 2024-01-23 scripts/0_pilot/02_assess_data.R

@@ -1,132 +1,0 @@
-
-ospan_data03 <- ospan_data |>
-  filter(variable == 'recall') |>
-  mutate(acc_prop = accuracy / set_size) |>
-  group_by(id, set_size) |>
-  summarise(wm_cap = mean(acc_prop)) |>
-  mutate(set_size = str_c("wm_cap_", set_size))
-
-ospan_data_math03 <-
-  ospan_data |>
-  filter(variable == "math") |>
-  select(id, rt, block, counterbalance, correct) |>
-  group_by(id) |>
-  summarise(acc_ospan_math = sum(correct) / n())
-
-ospan_data03 |>
-  ggplot(aes(wm_cap)) +
-  geom_histogram() +
-  facet_grid(set_size ~ .)
-
-ospan_data_math03 |>
-  ggplot(aes(acc_ospan_math)) +
-  geom_density()
-
-
-bind_upd_number_data03 <- bind_upd_number_data |>
-  mutate(type = ifelse(nUpd > 0, "updating", "binding")) |>
-  mutate(steps = ifelse(type == "binding", nBind, nUpd)) |>
-  mutate(comb = str_c(nBind, nUpd, sep = "_")) |>
-  group_by(id, comb) |>
-  summarise(num_acc = sum(accuracy)/n()) |>
-  ungroup() |>
-  pivot_wider(names_from = 'comb', values_from = 'num_acc') |>
-  select(-id) |>
-  cor() |>
-  corrplot::corrplot(method = "number")
-
-bind_upd_color_data03 <- bind_upd_color_data |>
-  mutate(type = ifelse(nUpd > 0, "updating", "binding")) |>
-  mutate(steps = ifelse(type == "binding", nBind, nUpd)) |>
-  mutate(type_steps = str_c(type, steps, sep = "_")) |>
-  group_by(id, type_steps) |>
-  summarise(num_acc = sum(accuracy)/n()) |>
-  ungroup() |>
-  pivot_wider(names_from = 'type_steps', values_from = 'num_acc') |>
-  left_join(ospan_data02) |>
-  select(-id) |>
-  cor() |>
-  corrplot::corrplot(method = "number")
-
-
-
-# Timing ------------------------------------------------------------------
-
-timing_no_outliers <-
-  bind_rows(
-    study_data |>
-      select(id, data_bind_upd_number01, data_bind_upd_number02, data_bind_upd_number03) |>
[…]

### 2024-01-23 scripts/1_sim_pipeline/2_exclusions.R

@@ -13,15 +13,16 @@ load("data/ivs_sim.RData")
 
 # 3. Exclusions -----------------------------------------------------------
 
-full_data <- full_join(ivs_sim, dvs_sim) |>
+full_data_sim <- full_join(ivs_sim, dvs_sim) |>
   mutate(
     ex_brows_int       = ifelse(brows_int == 0, FALSE, TRUE),
     ex_acc_ospan_sec   = ifelse(ospan_sec_acc*100 < gbinom(45, 0.50), TRUE, FALSE),
     ex_acc_rspan_sec   = ifelse(rspan_sec_acc*100 < gbinom(45, 0.50), TRUE, FALSE),
   )
 
-clean_data <- full_data |>
+clean_data_sim <- full_data_sim |>
   filter(ex_brows_int == FALSE & ex_acc_ospan_sec == FALSE & ex_acc_rspan_sec == FALSE)
 
-save(full_data, clean_data, file = "data/full_data.Rdata")
+
+save(full_data_sim, clean_data_sim, file = "data/full_data_sim.RData")
 

### 2024-01-23 scripts/1_sim_pipeline/4_sem.R

@@ -7,14 +7,14 @@ library(lavaan.survey)
 
 # 2. Load data ------------------------------------------------------------
 
-load("data/full_data.RData")
+load("data/full_data_sim.RData")
 
 # 2. Specify Clustering Design -----------------------------------------------
 
 # Some children are clustered within families.
 # We account for this using the lavaan.survey package
 
-cluster_design <- survey::svydesign(ids=~nohouse_encr, prob=~1, data = data_clean)
+cluster_design <- survey::svydesign(ids=~nohouse_encr, prob=~1, data = data_clean_sim)
 
 
 # 3. Model specification(s) --------------------------------------------------
@@ -35,7 +35,7 @@ mod_meas <-
 
 fit_meas1 <- lavaan::sem(
   model = mod_meas,
-  data = data_clean,
+  data = data_clean_sim,
   estimator = "MLR", # Robust SEs in case of non-normal data
   missing = "ML",    # FIML for missing data
   orthogonal = TRUE,

### 2024-01-23 data/README.Rmd
commit simulated data

---
title: ""
output: 
  html_document:
    template: ../assets/template.html
    css: ../assets/style.css
editor_options: 
  markdown: 
    wrap: sentence
---

## Data {#data}

This folder contains all the data that serves as input for the analysis scripts.
LISS data cannot be shared on open repositories.

**Full reproduction**. In order to fully reproduce the analysis described in the manuscript, the interested reader needs to have their own personal access to the LISS data archive.
After obtaining access, the required data files can be placed in the `data` folder (see [how to reproduce this repository](https://stefanvermeent.github.io/liss_wm_profiles_2023/) for more information).

**Computational reproduction**. To facilitate computational reproducibility, we provide simulated data files.
These files contain simulated data with the same variables as the original data.
They are generated under `scrips/1_sim_pipeline` and can be used as input to the data analysis functions.
Note that, because the values are not identical to the real data, the output will deviate from the statistics described in the manuscript.

### 2024-01-23 README.Rmd
update README

@@
 
-1.  [`Manuscript`](https://stefanvermeent.github.io/abcd_ddm/registered_report/README.html): The Registered Report written in Quarto.
-2.  [`Supplement`](https://stefanvermeent.github.io/abcd_ddm/supplement/README.html): a supplemental text (to be submitted with the manuscript) documenting all secondary analyses in detail.
-3.  [`Scripts`](https://stefanvermeent.github.io/abcd_ddm/scripts/README.html): R-scripts that read, analyze, and produce all analysis objects.
+1.  [`Manuscript`](https://stefanvermeent.github.io/liss_wm_profiles_2023/manuscript/README.html): The Registered Report written in Quarto.
+2.
+[`Supplement`](https://stefanvermeent.github.io/liss_wm_profiles_2023/supplement/README.html): a supplemental text (to be submitted with the manuscript) documenting all secondary analyses in detail.
+3.  [`Scripts`](https://stefanvermeent.github.io/liss_wm_profiles_2023/scripts/README.html): R-scripts that read, analyze, and produce all analysis objects.
+4.  [`Materials`](https://stefanvermeent.github.io/liss_wm_profiles_2023/materials/README.html): JsPsych scripts of the Working Memory tasks.
 4.
-[`Data`](https://stefanvermeent.github.io/abcd_ddm/data/README.html): Folder in which real LISS data can be placed to make the analyses fully reproducible.
+[`Data`](https://stefanvermeent.github.io/liss_wm_profiles_2023/data/README.html): Folder in which real LISS data can be placed to make the analyses fully reproducible.
 Note that we cannot openly share the raw data on the open repository.
-The folder contains synthetic data to facilitate computational reproducibility in the absence of LISS access.
-5.  [`Analysis Objects`](https://stefanvermeent.github.io/abcd_ddm/analysis_objects/README.html): Folder containing all analysis objects.
-6.  [`Codebooks`](https://stefanvermeent.github.io/abcd_ddm/codebooks/README.html): lists of variable names, labels, and value labels (where applicable).
+The folder contains simulated data to facilitate computational reproducibility in the absence of LISS access.
+5.  [`Analysis Objects`](https://stefanvermeent.github.io/liss_wm_profiles_2023/analysis_objects/README.html): Folder containing all analysis objects (as produced by the analysis scripts).
+6.  [`Codebooks`](https://stefanvermeent.github.io/liss_wm_profiles_2023/codebooks/README.html): lists of variable names, labels, and value labels (where applicable).
 

### 2024-01-24 .projectlog/405be71518cfc40e3de09f82e499412a9b29ee64.R
code scripts to .projectlog

@@ -1,0 +1,13 @@
+### Date: 2023-05-01 09:22:26
+
+### Description: LISS background variables March 2023, shuffled ids
+
+
+### For more information on this commit, see the README file, or go to https://github.com/StefanVermeent/liss_wm_profiles_2023/commit/405be71518cfc40e3de09f82e499412a9b29ee64
+
+### Below is the full code that was used to access the data:
+
+
+readr::read_delim('data/avars_202303_EN_1.0p.csv', col_select = NULL,  delim = ';') |>
+ dplyr::filter() |>
+ shuffle(data = _, shuffle_vars = 'nomem_encr', long_format = FALSE, seed = 3985843)

### 2024-01-24 .projectlog/4904078af6fd51f7430a2b25997b5ea307a51a19.R

@@ -1,0 +1,13 @@
+### Date: 2023-05-01 09:33:43
+
+### Description: LISS background variables March 2023, shuffled ids
+
+
+### For more information on this commit, see the README file, or go to https://github.com/StefanVermeent/liss_wm_profiles_2023/commit/4904078af6fd51f7430a2b25997b5ea307a51a19
+
+### Below is the full code that was used to access the data:
+
+
+readr::read_delim('data/avars_202302_EN_1.0p.csv', col_select = NULL,  delim = ';') |>
+ dplyr::filter() |>
+ shuffle(data = _, shuffle_vars = 'nomem_encr', long_format = FALSE, seed = 3985843)

### 2024-01-24 .projectlog/598922e78bf7abcc1b5160e06355e32a73d02be5.R

@@ -1,0 +1,13 @@
+### Date: 2023-05-01 09:40:53
+
+### Description: Crime Victimization Data Wave 5, shuffled ids
+
+
+### For more information on this commit, see the README file, or go to https://github.com/StefanVermeent/liss_wm_profiles_2023/commit/598922e78bf7abcc1b5160e06355e32a73d02be5
+
+### Below is the full code that was used to access the data:
+
+
+readr::read_delim('data/ac16e_EN_1.0p.csv', col_select = NULL,  delim = ';') |>
+ dplyr::filter() |>
+ shuffle(data = _, shuffle_vars = 'nomem_encr', long_format = FALSE, seed = 3985843)

### 2024-01-24 .projectlog/82f126ec06653b3794b6bb710b8a04b4e697d630.R

@@ -1,0 +1,13 @@
+### Date: 2023-05-01 09:41:18
+
+### Description: Crime Victimization Data Wave 4, shuffled ids
+
+
+### For more information on this commit, see the README file, or go to https://github.com/StefanVermeent/liss_wm_profiles_2023/commit/82f126ec06653b3794b6bb710b8a04b4e697d630
+
+### Below is the full code that was used to access the data:
+
+
+readr::read_delim('data/ac14d_EN_1.0p.csv', col_select = NULL,  delim = ';') |>
+ dplyr::filter() |>
+ shuffle(data = _, shuffle_vars = 'nomem_encr', long_format = FALSE, seed = 3985843)

### 2024-01-24 .projectlog/8f4ce9b0862b8107c7d02b93646c8d21374d40d2.R

@@ -1,0 +1,13 @@
+### Date: 2023-05-01 09:34:57
+
+### Description: LISS background variables January 2023, shuffled ids
+
+
+### For more information on this commit, see the README file, or go to https://github.com/StefanVermeent/liss_wm_profiles_2023/commit/8f4ce9b0862b8107c7d02b93646c8d21374d40d2
+
+### Below is the full code that was used to access the data:
+
+
+readr::read_delim('data/avars_202301_EN_1.0p.csv', col_select = NULL,  delim = ';') |>
+ dplyr::filter() |>
+ shuffle(data = _, shuffle_vars = 'nomem_encr', long_format = FALSE, seed = 3985843)

### 2024-01-24 .projectlog/fcaf6747766662ee20f9709a2d0262b766ea4500.R

@@ -1,0 +1,13 @@
+### Date: 2023-05-01 09:37:28
+
+### Description: Crime Victimization Data Wave 6, shuffled ids
+
+
+### For more information on this commit, see the README file, or go to https://github.com/StefanVermeent/liss_wm_profiles_2023/commit/fcaf6747766662ee20f9709a2d0262b766ea4500
+
+### Below is the full code that was used to access the data:
+
+
+readr::read_delim('data/ac18f_EN_1.0p.csv', col_select = NULL,  delim = ';') |>
+ dplyr::filter() |>
+ shuffle(data = _, shuffle_vars = 'nomem_encr', long_format = FALSE, seed = 3985843)

### 2024-01-24 .projectlog/generate_scripts.R

@@ -1,0 +1,34 @@
+gert::git_log(max = 10000) |> 
+  select(commit, author, time, message) |>
+  (function(.) filter(., str_detect(string = .$message, pattern = "^\\[\\[")))() |> 
+  separate(message, into = c("message", "hash", 'code'), sep = "\n") |> 
+  separate(message, into = c("milestone", "description"), sep = "\\]\\]\\s") |> 
+  mutate(milestone = str_remove(milestone, "\\[\\[")) |> 
+  pwalk(function(commit,author,time,message,hash,milestone,description, code){
+    
+    if(milestone == "data_access"){
+      code <- code |> 
+        str_replace_all(string = _, "\\|> ", "\\|>\n") |> 
+        str_remove(string = _, "^code ") 
+      
+      url <- gert::git_remote_list()$url |> 
+        str_remove(string = _, "\\.git$") |> 
+        paste0(paste0("/commit/", commit))
+      
+      #  url <- "<anonymized repository>"
+      
+      script <-
+        glue(
+          "
+          ### Date: {time}\n
+          ### Description: {description}\n\n
+          ### For more information on this commit, see the README file, or go to {url}\n
+          ### Below is the full code that was used to access the data:\n\n
+          {code}"
+        )
+      
+      writeLines(script, con = paste0(".projectlog/", commit, ".R"))
+    }
+  })
+
+

### 2024-01-24 manuscript/manuscript.qmd
in-progress manuscript

@@
 
-#### Stefan Vermeent^1,2^, Anna-Lena Schubert^3^, Meriah L. DeJoseph^4^, Jaap Denissen^1^, Jean-Louis van Gelder^2,5^, & Willem E. Frankenhuis^1,2,6^
+#### Stefan Vermeent^1,2^, Anna-Lena Schubert^3^, Meriah L. DeJoseph^4^, Jaap J. A. Denissen^1^, Jean-Louis van Gelder^2,5^, & Willem E. Frankenhuis^1,2,6^
 
@@
 
+# Corresponding author
+
+Correspondence concerning this article should be addressed to Stefan Vermeent, Department of Psychology, Utrecht University, Heidelberglaan 1, 3584 CS Utrecht, The Netherlands.
+Email: p.c.s.vermeent@gmail.com / p.c.s.vermeent@uu.nl.
+
 # Data Availability
 
-All scripts and materials needed to reproduce the findings are available on the article's Github repository ([https://github.com/stefanvermeent/liss-wm-profiles-2023/README.md](https://github.com/stefanvermeent/liss-wm-profiles-2023/README.md)).
-We also include instructions on how to reproduce each step of our analyses.
+All scripts and materials needed to reproduce the findings are available on the article's Github repository ([https://stefanvermeent.github.io/liss_wm_profiles_2023/](https://stefanvermeent.github.io/liss_wm_profiles_2023/)).
+We also include instructions on how to reproduce each step of our analyses, as well as simulated data to ensure computational reproducibility. 
 In this paper, we make use of data from the LISS panel (Longitudinal Internet studies for the Social Sciences) managed by the non-profit research institute Centerdata (Tilburg University, the Netherlands).
 All datasets are available in the LISS data archive. 
-Researchers who want to access the data are required to sign a statement confirming that information about individual persons, households, etc., will not be released to others (Go to [https://statements.centerdata.nl](https://statements.centerdata.nl) for more information).
+Researchers who want to access the data are required to sign a statement confirming that information about individual persons, households, etc., will not be released to others (go to [https://statements.centerdata.nl](https://statements.centerdata.nl) for more information).
 
 # Funding statement
-WEF’s contributions have been supported by the Dutch Research Council (V1.Vidi.195.130) and the James S.
-McDonnell Foundation (https://doi.org/10.37717/220020502).
-MLD was supported by the NICHD National Research Service Award (#1F32HD112065-01).
+WEF’s contributions have been supported by the Dutch Research Council (V1.Vidi.195.130) and the James S. McDonnell Foundation (https://doi.org/10.37717/220020502).
+MLD's contributions have been supported by the NICHD National Research Service Award (#1F32HD112065-01).
+JLvG's contributions have been supported by a Consolidator Grant from the European Research Council (772911– CRIMETIME).
 
 # Disclosures
+
 We declare no conflicts of interest.
@@
 
-XXX
+Decades of research have shown that adversity tends to lower working memory (WM) performance.
+This literature has mainly focused on impairments in the overall capacity to hold information available in WM for further processing.
+However, some recent adaptation-based studies suggest that certain types of adversity can leave intact, or even enhance, the ability to rapidly update information in WM.
+One key issue in evaluating deficits and adaptations in WM is that WM capacity and updating tasks tend to covary.
+This is likely due to the fact that both types of tasks require the creation and maintenance of bindings in WM; links between mental representations of information in WM.
+To estimate the associations between adversity and different processes in WM, we need to isolate variance in performance related to WM capacity from variance in performance related to updating ability.
+In this Registered Report, we combine archived and newly collected data in the Dutch Longitudinal Internet studies for the Social Sciences (LISS) panel, which includes a representative sample of the Dutch adult population.
+Participants completed three WM tasks: two complex span tasks and a task measuring both binding and updating of information.
+In addition, we will derive estimates of participants' exposure to neighborhood threat, material deprivation, and unpredictability.
+Using structural equation modeling, we will estimate associations between the three types of adversity and latent estimates of WM capacity and updating.
+These findings will advance our theoretical understanding of how adversity is associated with WM, which will aid the development of interventions aimed at alleviating performance difficulties and leveraging areas of strengths.
 
@@
 
-Living in adverse conditions, with prolonged exposure to intense stress, tends to have a substantial impact on cognitive functioning.
-A domain that seems to be particularly affected by adversity is working memory (WM), a system for mentally building, maintaining, and updating immediately relevant information [@oberauer_2018].
-WM is associated with a host of social and cognitive abilities, such as mathematics ability [@peng_2016], reading ability [@chiappe_2000], learning [@cowan_2014], general intelligence [@conway_2003], and mentalizing [@mutter_2006].
-Not surprisingly, then, deficits in WM have negative consequences for educational and professional outcomes [@ahmed_2018; @alloway_2010; @guo_2020; @spiegel_2021].
-Although decades of research show that adversity is generally negatively associated with WM [@goodman_2019], a small, emerging literature suggests that specific aspects of WM might be intact or even enhanced through developmental adaptations to adversity.
-So far, these literatures have tended to focus on related, but different aspects of WM, limiting a fuller integration.
-Here, we take a psychometric modeling approach to simultaneously examine potential decreases and enhancements in WM performance.
+Living in adverse conditions, with prolonged exposure to intense stress, tends to have a profound and enduring impact on cognitive functioning [@farah_2006; @sheridan_2022; @sheridan_2014].
+A domain that seems to be particularly affected by adversity is working memory (WM). 
+WM is a system for mentally building, maintaining, and updating immediately relevant information [@oberauer_2018].
+Performance on WM tasks is associated with a host of social and cognitive abilities, such as math [@peng_2016], reading [@chiappe_2000], learning [@cowan_2014], general intelligence [@conway_2003], and mentalizing [@mutter_2006].
+Not surprisingly, then, deficits in WM have negative consequences for both educational and professional outcomes [@ahmed_2018; @alloway_2010; @guo_2020; @spiegel_2021].
+Decades of research show that adversity is generally negatively associated with performance on WM tasks [@goodman_2019].
+However, emerging evidence suggests that specific aspects of WM might remain intact or even be enhanced through developmental adaptations to adversity.
+So far, the literature has tended to focus on related, but different aspects of WM in isolation, limiting a fuller integration.
+Here, we take a psychometric modeling approach to simultaneously examine potential decreases and enhancements in two WM components: capacity and updating.
 
@@
 
-A large literature has shown negative associations between exposures to adversity and WM, potentially attributable to the enduring influence of stress on several key brain regions that support WM [@duval_2017; @hanson_2012].
-Although not always mentioned explicitly, much of this work has focused on WM capacity, or the ability to keep multiple pieces of information simultaneously available for further processing.
+A large literature has shown negative associations between exposures to adversity and performance on WM tasks [@farah_2006; @sheridan_2022; @sheridan_2014]. 
+These associations may be potentially attributable to the enduring influence of stress on several key brain regions that support WM [@duval_2017; @hanson_2012].
+Much of this work has focused on WM capacity, or the ability to keep multiple pieces of information simultaneously available for further processing.
 For early-life adversity, this negative association is already present during childhood, and persists into adulthood [@bos_2009; @evans_2009; @hackman_2010; @farah_2006; @goodman_2019; @noble_2007; but see @nweze_2021].
-Other studies have found a similar link between recent adverse experiences and lowered WM capacity in college student samples [@klein_2001; @shields_2017; @shields_2019].
-The most often used tasks have been simple span tasks (repeating a string of stimuli of increasing length), complex span tasks (remembering a string of stimuli while being engaged by a secondary task), and *n*-back tasks (judging whether the current stimulus in a string is identical to the stimulus *n* steps ago)  [@goodman_2019].
-Performance on these tasks is assessed through the number of items that can be retained in WM, that is, its overall capacity [with the exception of *n*-back\; for concerns about the construct validity of this task, see @frost_2021; @kane_2007].
-Other research has shown that recent life stress is negatively associated with WM capacity, but not the precision of representations in WM [@shields_2019].
+Studies with college students have found a link between both recent and lifetime experiences of stressful major life events (discrete negative events that have a clear onset and offset, unlike chronic adversity) with lowered WM capacity [@klein_2001; @shields_2017; @shields_2019].
+The most common tasks used to examine this link are simple span tasks (repeating a string of stimuli of increasing length), complex span tasks (remembering a string of stimuli while being engaged by a secondary task), and *n*-back tasks (judging whether the current stimulus in a string is identical to the stimulus *n* steps ago)  [@goodman_2019].
+Performance on these tasks is assessed through the number of items that participants can retain in WM, that is, their overall capacity [with the exception of *n*-back\; for concerns about the construct validity of this task, see @frost_2021; @kane_2007].
 
-Although both early-life and recent adversity appear to lower WM capacity, a small set of studies suggest that exposure to adversity may leave intact, or even enhance, WM updating (i.e., the ability to rapidly replace old information in WM with new information) [@young_2018; @young_2022].
-These findings exemplify an emerging, theoretical framework grounded in adaptation reasoning that is complementary to deficit frameworks [@frankenhuis_2013; @ellis_2017; @ellis_2022; @frankenhuis_2020].
-This adaptation framework assumes that developmental processes tailor an individual's cognitive abilities to the unique challenges and opportunities posed by their environment.
+Although both early-life and recent adversity appear to be negatively associated with WM capacity, a small set of studies suggest that exposure to adversity may leave intact, or even enhance, the ability to update items in WM [@young_2018; @young_2022].
+Updating is defined as the ability to rapidly replace old information in WM with new information.
+The finding that updating may be left intact or even enhanced after exposure to adversity exemplifies emerging theoretical frameworks grounded in adaptive reasoning that are complementary to deficit frameworks [@frankenhuis_2013; @ellis_2017; @ellis_2022; @frankenhuis_2020].
+Adaptation frameworks assume that developmental processes tailor an individual's cognitive abilities to the unique challenges and opportunities posed by their environment.
 The link between adversity and cognitive abilities is further assumed to be specific; as different types of adversity (e.g., threat vs.
-deprivation) pose different challenges, they should (at least in part) shape different cognitive abilities.
-WM updating has been argued to be especially adaptive in environments that are unpredictable, which can be defined as random variation in harshness over space or time [@ellis_2009].
-In such environments, WM updating allows people to maintain an up-to-date overview of the current state of the environment [@young_2018].
-Improved WM updating performance has also been documented for threat exposure [@young_2022], and unpredictability might particularly shape cognition in circumstances of high deprivation, such as lower-resource contexts like poverty [@li_2018].
+deprivation) pose different challenges, they should (at least in part) shape cognitive abilities in different ways.
+WM updating may be especially adaptive in *unpredictable* environments, characterized by random variation in adversity exposure over space or time.
+WM updating allows people to maintain an up-to-date overview of the (changing) current state of the environment [@young_2018].
+Additionally, improved WM updating performance has also been documented for threat exposure [@young_2022].
+An enhanced WM updating ability could facilitate keeping track of and integrating signals that may potentially signal acutely threatening situations.
 
@@
 
-With deficit frameworks mainly focusing on WM capacity and adaptation frameworks on WM updating, we may wonder how these WM components are related to each other.
-Research to date has demonstrated substantial overlap between capacity and updating tasks using structural equation modeling (SEM).
-For example, correlations between latent capacity and updating factors can be high, to the extent that both complex span and updating paradigms are seen by some as valid measures of WM capacity [@wilhelm_2013].
-In fact, some research suggests that the amount of updating-specific variance on typical updating tasks might be as low as 15% [@frischkorn_2022].
-This overlap might stem from the fact that both WM updating and capacity tasks require maintaining information in WM for a short time [@ecker_2010; @frischkorn_2022; @oberauer_2000; @schmiedek_2009; @wilhelm_2013].
+With deficit frameworks focusing on WM capacity and adaptation frameworks on WM updating, we may wonder how capacity and updating are related to each other.
+Performance on tasks measuring WM capacity and updating tend to be substantially correlated [in the range of .20-.50\; @frischkorn_2022].
+This overlap appears to stem from shared demands of both types of tasks, in particular the need to create and maintain arbitrary bindings [@gruszka_2017; @oberauer_2009; @wilhelm_2013].
+The term *binding* refers to the process of mapping memory items to specific positions in WM (e.g., serial, spatial, or temporal positions, depending on the task) [@oberauer_2009; @oberauer_2019].
+For example, on most WM tasks, correct recall of memory items depends on remembering them in their correct serial position, or in relation to the location where they were presented.
 
-Nevertheless, updating tasks might involve processes other than maintenance that set them apart from WM capacity tasks.
-Updating tasks require different combinations of retrieval (making information available for immediate processing), transformation (changing a prior value into a new one, e.g., by addition or subtraction), and substitution (replacing a prior value for a new value) [@ecker_2010].
-@ecker_2010 measured three indicators of WM capacity as well as eight versions of a WM updating task that required different combinations of retrieval, transformation, and substitution.
-They found positive correlations of around .50 between WM capacity with retrieval and transformation accuracy, whereas substitution accuracy was not correlated with WM capacity.
-Thus, the ability to accurately substitute old with new information---a key aspect of WM updating---seems to be independent from WM capacity.
+The centrality of binding in WM is supported by theoretical models of WM and by empirical work showing that (latent) WM capacity is strongly related to the ability to maintain bindings [@oberauer_2000; @oberauer_2005; @oberauer_2009; @oberauer_2019; @wilhelm_2013].
+The number of bindings a person can create and maintain in WM might be the main limiting factor in WM capacity, as maintaining several bindings at the same time will increasingly lead to interference between them [@gruszka_2017; @oberauer_2009; @wilhelm_2013].
+This upper limit on WM capacity also affects performance on WM updating tasks.
+That is, updating items in WM requires not just dissolving old bindings and creating new ones, but also maintaining bindings of items that should not be updated.
+Thus, the overlap in performance on WM updating and capacity tasks likely stems from the need in both types of tasks to create and maintain bindings in WM [@ecker_2010; @frischkorn_2022; @oberauer_2000; @schmiedek_2009; @wilhelm_2013].
+
+Nevertheless, WM updating tasks additionally require the updating of established bindings, which sets them apart from WM capacity tasks [@ecker_2010; @frischkorn_2022].
+Different updating tasks require different combinations of retrieval (making information available for immediate processing), transformation (changing a prior value into a new one, e.g., by addition or subtraction), and substitution (replacing a prior value for a new value) [@ecker_2010].
+@ecker_2010 included three measures of WM capacity as well as eight versions of a WM updating measure that required different combinations of retrieval, transformation, and substitution.
+After accounting for overall updating accuracy (which was positively correlated with WM capacity), they found positive correlations of around .50 between WM capacity with latent estimates of retrieval and transformation accuracy, but not with a latent estimate of substitution accuracy.
+Thus, when the ability to accurately substitute old with new information---a key aspect of WM updating---is sufficiently isolated from WM capacity using latent modeling, capacity and updating seem to be independent components of WM.
 
@@
 Yet, to our knowledge, no previous research has analyzed both abilities within a single statistical model.
-This could lead us to underestimate---or altogether miss---ways in which adversity enhances WM components, with implications for basic and applied science.
-For basic science, it could bias inferences about WM abilities, especially when the negative association between adversity and WM capacity is stronger than the positive association with WM updating.
+This could lead to (1) an underestimation of the extent to which adversity undermines WM capacity, and (2) underestimation of the extent to which adversity can enhance WM updating. 
+This, in turn, has implications for basic and applied science.
+For basic science, it could bias inferences about individual differences in performance on WM tasks, especially when the negative association between adversity and WM capacity is stronger than the positive association with WM updating.
 For applied science, it could hide from view potential pathways to leverage people's existing strengths in school or work contexts.
@@
 Specifically, we will measure exposure to two forms of harshness: *neighborhood threat* (perceived neighborhood safety, personal crime victimization), and *material deprivation* (perceived scarcity, income-to-needs).
-We further include a measure of unpredictability *unpredictability*, indexed as random variation in deprivation-induced harshness over time.
-Together, these adversity types capture key dimensions in contemporary models of adversity [threat vs. deprivation vs. unpredictability\; @ellis_2009; @ellis_2022; @mclaughlin_2016; @mclaughlin_2019].
-Although it would have also been theoretically interesting to look at unpredictability in neighborhood threat, our dataset does not include enough timepoints to compute such estimates.
+We also include a measure of *unpredictability*, indexed as variability in material deprivation over time.
+Together, these adversity types capture key dimensions in contemporary models of adversity [threat vs. deprivation vs. unpredictability\; @ellis_2009; @ellis_2022; @mclaughlin_2016; @mclaughlin_2021].
+We will not calculate unpredictability in neighborhood threat given that participants have at most six timepoints, and often as few as one or two, which is insufficient to calculate variation over time [@walasek_2023].
 We will address three research questions.
-First, what is the association of adversity with WM capacity?
-Second, what is the association of adversity with WM updating *after* accounting for WM capacity?
-Third, are these two associations similar or different for neighborhood threat, material deprivation, and unpredictability?
+First, what is the association between adversity and WM capacity?
+Second, what is the association between adversity and WM updating *after* accounting for WM capacity?
+Third, are the directions and strengths of these associations similar or different for neighborhood threat, material deprivation, and unpredictability?
 
-Theoretically, we will evaluate evidence for deficit and adaptation frameworks.
-Figure 1 summarizes which combinations of lowered, enhanced, and intact WM capacity and updating performance would provide support for deficit and adaptation frameworks.
-Crucially, we assume that deficit and adaptation processes can operate simultaneously [@frankenhuis_2020], meaning that we could find (lack of) support for both frameworks in the same model.
-A hypothesis generated within a deficit framework would be consistent with lowered performance on WM capacity or updating, but not with enhanced or intact performance.
+We will evaluate evidence for deficit and adaptation frameworks.
+Rather than banking on specific predictions, we will describe how each of four different combinations of lowered, enhanced, and intact WM capacity and updating performance would be consistent with deficit and adaptation frameworks (For a summary, see Figure 1A).
+Crucially, we assume that deficit and adaptation processes can operate simultaneously [@frankenhuis_2020], meaning that we could find support or a lack of support for both frameworks in the same model.
+Lowered performance on WM capacity or updating, but not enhanced or intact performance, would be consistent with deficit frameworks.
 This follows from previous literature hypothesizing broad WM deficits [@farah_2006; @sheridan_2020].
-A hypothesis generated within an adaptation framework would be consistent with two different patterns.
-First, if performance on WM updating is enhanced.
+Two different patterns would be consistent with adaptation frameworks:
+First, if performance on WM updating is enhanced, regardless of WM capacity performance.
 Second, if performance on WM updating is intact, but only in combination with lowered WM capacity performance.
-If both updating and capacity are intact, this would instead suggest that the WM system is unaffected by adversity (i.e., neither impairment nor adaptation processes have shaped WM).
-
-Figure 1B depicts the theoretical possibilities in relation to deficit or adaptation frameworks (as formalized in Figure 1A) and how we hypothesize them to be linked to threat, material deprivation, and unpredictability.
-First, we expect that all three, and in particular material deprivation, will be associated with deficits in WM capacity, consistent with literature showing that deprivation is most strongly associated with cognitive deficits [@salhi_2021; @sheridan_2020].
-In this literature, most studies have focused on cognitive deprivation, which is not identical to material deprivation (as measured here).
-However, both are related to a person's access to stimulating resources (e.g., formal education, books\; @heppt_2022).
-Most research has focused on children, so it is unclear whether and how cognitive and material deprivation affects cognitive outcomes in adults.
-
-In contrast, we expect adaptations in WM updating to be most strongly associated with unpredictability [@young_2018; @young_2022], as well as with threat [in line with existing empirical findings\; @young_2022].
-In addition, we expect an interaction between unpredictability and mean levels of deprivation, such that WM updating shows the highest increase under conditions of high material deprivation *and* high unpredictability.
-Unpredictability is challenging to define, and has been measured in many different ways, ranging from number of residential changes, to perceived inconsistencies in household routines, to random variation in income [@ugarte_2023; @walasek_2023; @young_2020].
-
-One issue is that timescales of unpredictability measures may not always align with those of the cognitive abilities that are assumed to be shaped by them.
-For example, while residential changes occur on timescales of months or years, WM updating constitutes a process unfolding on the timescale of seconds to minutes.
-However, we expect that our measure of unpredictability (random variation in perceived resource scarcity and income-to-needs based on yearly measures) is a proxy for everyday situations that require rapid updating of information in WM.
-For example, yearly fluctuations in material deprivation, especially for someone living close to the poverty line, could reflect many everyday instances in which the person needs to update information about their current financial state in order to meet basic needs (e.g., because of an unpredictable income, or unforeseen costs or debts, that constrain everyday decisions like how much to spend on groceries).
+If both updating and capacity are intact, this would instead suggest that the WM system is unaffected by adversity; neither impairment nor adaptative processes have shaped WM.
+That pattern would be inconsistent with both deficit and adaptation frameworks.
 
@@
 Panel A depicts the most likely data patterns based on previous literature, and whether we would consider them consistent with deficit and adaptation frameworks (see the main text for more details).
-Panel B depicts an overview of the Structural Equation Model, including predictions for associations between different dimensions of adversity and WM components.
+Panel B depicts an overview of the Structural Equation Model.
+Ellipses represent latent variables, rectangles represent manifest variables, and circles represent residual variances.
+Unidirectional solid lines represent factor loadings, bidirectional solid lines represent covariances, and dashed lines represent regression paths.
 All four manifest WM measures load on a latent WM capacity factor, reflecting the fact that people have to hold information active in WM on all tasks.
+We fix the loading of WM capacity on the Binding Task to 1, reflecting the idea that the ability to create and maintain bindings is the main limiting factor in WM capacity [@gruszka_2017; @oberauer_2009; @wilhelm_2013].
 WM updating is modeled as a latent factor capturing the residual variance in the updating task after accounting for variance related to WM capacity.
@@
 Households without a computer or internet connection are provided with these facilities by LISS.
-Participants complete a yearly core battery of questionnaires about various aspects of their lives.
-In addition, participants can complete online questionnaires every month, which are more variable.
-The current study includes archived data as well as new data collected as part of a monthly data collection that took place between October 2nd and December 22nd, 2023.
-We signed a contract with LISS stipulating that we will receive access to the data only after Stage 1 acceptance of the Registered Report.
+Each year, participants complete the same core battery of questionnaires about---among other topics---their financial situation in the past year.
+In addition, participants can complete additional online questionnaires every month, with variable content.
+The current study integrates two data sources.
+First, we will use data that were previously collected in LISS (hereafter referred to as the LISS archive).
+Second, we will use new data that we collected ourselves in LISS between October 2nd and December 22nd, 2023 (hereafter referred to as newly collected data).
+We signed a contract with LISS stipulating that we will receive access to the data only after Stage 1 acceptance of this Registered Report.
 
 We based our power analysis on simulations reported by @kretzschmar_2019, determining the required sample size to detect a small effect size ($\beta$ = 0.1) with at least 90% power at $\alpha$ = 0.05.
-Assuming a reliability of at least 0.7 [which is typical for WM tasks with a number of trials similar to ours\; e.g., @wilhelm_2013], we would require a sample size of N = 730.
-Anticipating some exclusions, we therefore decided to include 800 participants.
-Participants were eligible for inclusion if they 1) were currently between 18 and 55 years old, 2) had completed at least one wave of the 'Conventional and computer crime victimization' survey ([https://doi.org/10.17026/dans-zch-j8xt](https://doi.org/10.17026/dans-zch-j8xt)) which ran between 2008 and 2018 at two-year intervals, and 3) had given permission to link their LISS data to government microdata (not relevant here).
-To ensure sufficient representation of people from lower socioeconomic backgrounds, half the total sample was sampled from participants who reported one or more of the following at least once in the three years: (1) a monthly income < €1,500, (2) HAVO or VWO as highest completed education (which are the two highest levels in Dutch secondary education), or (3) a score of 4 or lower on the 'ladder of life' ("If you imagine a ‘ladder of life’, where the first step represents the worst possible life, and the tenth (top) step the best possible life, on what step would you place yourself?").
-Participants will be excluded if they (1) switched to and interacted with other browser tabls *during* one or more of the cognitive tasks, except when this happened in between tasks or during a scheduled break, (2)
+Assuming a reliability of at least 0.7 [which is typical for WM tasks with a number of trials similar to ours\; e.g., @wilhelm_2013], we would require a sample size of *N* = 730.
+Anticipating some exclusions, we decided to include 800 participants.
+Participants were eligible for inclusion if they 1) were currently between 18 and 55 years old, 2) had completed at least one wave of an archived longitudinal LISS study containing measures that we use to operationalize crime neighborhood threat (see below), and 3) had given permission to link their LISS data to government microdata (not relevant here).
+
+To ensure sufficient representation of people from lower socioeconomic backgrounds, *half* the total sample was sampled from participants who reported one or more of the following at least once in the three years: (1) a monthly income < €1,500, (2) HAVO or VWO as highest completed education (which are the two highest levels in Dutch secondary education), or (3) a score of 4 or lower on the 'ladder of life' ("If you imagine a ‘ladder of life’, where the first step represents the worst possible life, and the tenth (top) step the best possible life, on what step would you place yourself?").
+Participants will be excluded if they (1) switched to and interacted with other browser tabs *during* one or more of the cognitive tasks, (2) did not perform above chance level.
 
@@
 
-**Neighborhood crime.** We included four items from existing LISS data collected across six waves ([https://doi.org/10.17026/dans-zch-j8xt](https://doi.org/10.17026/dans-zch-j8xt)), in which participants answered how often it happens that they 1) "avoid certain areas in your place of residence because you perceive them as
+**Perceived neighborhood crime.** We included four items from the LISS archive collected across six waves ([https://doi.org/10.17026/dans-zch-j8xt](https://doi.org/10.17026/dans-zch-j8xt)), in which participants answered how often it happens that they 1) "avoid certain areas in your place of residence because you perceive them as
 unsafe", 2) "do not respond to a call at the door because you feel that it is unsafe", 3) "leave valuable items at home to avoid theft or robbery in the street?", 4) "make a detour, by car or on foot, to avoid unsafe areas?" on a scale of 1 ("(Almost) never"), 2 ("Sometimes"), or 3 ("Often").
@@
 
-In addition, we will use the Neighborhood Violence Scale [NVS\; @frankenhuis_2018; @frankenhuis_2020] which was included in the current study.
+In addition, we implemented the Neighborhood Violence Scale [NVS\; @frankenhuis_2018; @frankenhuis_2020] in the newly collected data.
 The NVS includes seven items measuring perceived exposure to neighborhood violence (e.g., "Crime is common in the neighborhood where I live"; "Where I live, it is important to be able to defend yourself against physical harm").
 Participants answered these questions on a scale of 1 ("Completely disagree") to 7 ("Completely agree").
+We will compute an average of the seven items.
 
-**Crime victimization.** We used existing LISS data collected across six waves ([https://doi.org/10.17026/dans-zch-j8xt](https://doi.org/10.17026/dans-zch-j8xt)), in which participants indicated whether they fell victim to eight types of crime over the two years prior to a particular wave.
-We included seven items concerning physical crime: 1) burglary or attempted burglary; 2) theft from their car; 3) theft of their wallet or purse, handbag, or other personal possession; 4) wreckage of their car or other private property; 5) intimidation by any other means; 6) maltreatment of such serious nature that it required medical attention; 7) maltreatment that did not require medical attention.
-For each type of crime that participants indicated having been exposed to, they also indicated the frequency.
-We will average the frequencies for each type of crime separately across waves for which participants have data.
+**Crime victimization.** We used data from the LISS archive collected across six waves (same dataset as above), in which participants indicated whether they fell victim to eight types of crime over the two years prior to a particular wave (0 = no, 1 = yes).
+We included seven items concerning exposure to crime: (1) burglary or attempted burglary; (2) theft from their car; (3) theft of their wallet or purse, handbag, or other personal possession; (4) wreckage of their car or other private property; (5) intimidation by any other means; (6) maltreatment of such serious nature that it required medical attention; (7) maltreatment that did not require medical attention.
+We compute a variety score by summing the exposures to *unique* types of crime across all waves.
+Thus, if a participants reports exposure to the same type of crime on separate waves, this will count as one exposure in the total score [@sweeten_2012].
 
-**Overall composite.** To compute an overall measure of neighborhood threat exposure, we will submit the averages of each measure (the two measures of neighborhood crime and average frequencies of each crime exposure) to a Principal Component Analysis (PCA).
-We will extract a single factor score that explains the largest part of the variance.
-Higher scores will indicate higher levels of neighborhood threat.
+**Neighborhood threat composite.** We will compute a composite measure of neighborhood threat by aggregating across the perceived neighborhood crime and crime victimization measures exposure.
+First, we will create a single measure of perceived neighborhood crime by separately standardizing the two scales and averaging them.
+Second, we will average across perceived neighborhood crime and crime victimization, standardizing each before averaging.
 
@@
 
-**Perceived scarcity.** We will use a few items from LISS data that was collected on a yearly basis between 2007 and 2023 ([https://doi.org/10.57990/1gr4-bf42](https://doi.org/10.57990/1gr4-bf42)) to index perceived scarcity.
-First, participants indicated how hard or easy it currently is to live off the income of their household, on a scale of 0 (very hard) to 10 (very easy).
-Second, participants were asked to choose which of the following best applied to their current situation: 1) "we are accumulating debt"; 2) "we are somewhat eating into savings"; 3) "we are just managing to make ends meet"; 4) "we have a little bit of money to spare"; 5) "we have a lot of money to spare".
-Their response was reverse-coded, so that a higher score indicated a worse financial situation.
-Third, participants answered which of the following issues they were confronted with at present (0 = no, 1 = yes): 1) "having trouble making ends meet"; 2) unable to quickly replace things that break"; 3) "having to lend money for necessary expenditures"; 4) "running behind in paying rent/mortgage or general utilities"; 5) "debt collector/bailiff at the door in the last month"; 6) "received financial support from family or friends in the last month".
-Fourth, participants answered whether their household expenditure was more than, equal to, or less than their household income over the last 12 months, not counting large investments (e.g., buying a house).
-We will dummy-code this variable, using expenditure lower than income as the reference category.
+**Perceived scarcity.** We will use a few items from the LISS archive that were collected on a yearly basis between 2007 and 2023 ([https://doi.org/10.57990/1gr4-bf42](https://doi.org/10.57990/1gr4-bf42)) to index perceived scarcity.
+First, participants indicated how hard or easy it currently is to live off the income of their household, on a scale from 0 (very hard) to 10 (very easy).
+Second, participants were asked to choose which of the following best applied to their current situation: (1) "we are accumulating debt"; (2) "we are somewhat eating into savings"; (3) "we are just managing to make ends meet"; (4) "we have a little bit of money to spare"; (5) "we have a lot of money to spare".
+Responses will be reverse-coded, so that higher scores indicate a worse financial situation.
+Third, participants answered which of the following issues they were confronted with at present (0 = no, 1 = yes): (1) "having trouble making ends meet"; (2) unable to quickly replace things that break"; (3) "having to lend money for necessary expenditures"; (4) "running behind in paying rent/mortgage or general utilities"; (5) "debt collector/bailiff at the door in the last month"; (6) "received financial support from family or friends in the last month".
+We will sum across these six items, with higher scores indicating more financial difficulties.
 
-**Income-to-needs.** We will calculate an income-to-needs ratio for each year using monthly self-reported net household income ([https://doi.org/10.57990/qn3k-as78](https://doi.org/10.57990/qn3k-as78)).
+**Income-to-needs.** We will calculate an income-to-needs ratio for each year using monthly self-reported net household income from the LISS archive ([https://doi.org/10.57990/qn3k-as78](https://doi.org/10.57990/qn3k-as78)).
+Zero values in household income will be set to missing, as these could either indicate the lack of an income or an unwillingness to disclose the income.
+If monthly household income is missing (or zero) for an entire year for a participant, we will use, if available, the yearly net household income they reported in the LISS archive ([https://doi.org/10.57990/1gr4-bf42](https://doi.org/10.57990/1gr4-bf42)), dividing it by 12 to obtain a monthly estimate.
 First, we average monthly reported income across years for each participant.
-Then, we divide the average income per year by the *poverty threshold*, as determined by Statistics Netherlands (CBS) [@brakel_2023].
+Then, we divide the average income per year by the *poverty threshold*, as determined by Statistics Netherlands (CBS) [@brakel_2023\; CBS, personal communication, December 15, 2023].
 The poverty threshold changes across time to account for purchasing power.
-In addition, different thresholds are determined each year for different household compositions (single adult vs (married) couple and number of children (up to three)) by multiplying the threshold of a single-person household without children with a particular equivalence factor.
-As thresholds are only provided for households with up to three children, we will use the equivalence factor of a household with three children for households with more than three children.
-Likewise, we will use the equivalence factor of a household with a (married) couple for households with more than two adults that are not children.
+In addition, different thresholds are determined each year for different household compositions (single vs. two adults; zero, one, two, or three children).
+As thresholds are only provided for households with up to three children, we will apply the threshold of a household with three children to households with more than three children.
+Likewise, we will apply the threshold of a household with two adults for households that contain three or more adults.
 
-**Overall composite.** Similar to neighborhood threat, we will use PCA to extract a single estimate of material deprivation measured by perceived scarcity and income-to-needs.
-In contrast to neighborhood threat, we will apply the PCA to the most recent data (2023) and use the resulting weights to compute material deprivation scores for previous years.
-This approach will allow us to compute both the mean level of material deprivation, as well as random variation in material deprivation over time.
-If a participant has partial data for a particular year, we will use multiple imputation to impute the value using the *mice* package [@buuren_2011].
-We will not impute data for years where a participant has missings on all indicators.
-We will compute within-subject averages of material deprivation across all years for which participants data available.
-Higher scores will indicate more material deprivation.
+**Material deprivation composite.** We will compute a composite measure of material deprivation by aggregating across the perceived scarcity measures and the income-to-needs ratio.
+We will do so separately for each year.
+First, we will create a single measure of perceived scarcity by separately standardizing the three measures and averaging them.
+Second, we will calculate the average within-person income-to-needs ratio for each year by averaging across the monthly income-to-needs estimates.
+We will reverse code the yearly income-to-needs ratio so that higher scores indicate more deprivation.
+Third, we will average across perceived scarcity and the income-to-needs ratio for each year, standardizing each before averaging, and then finally calculate the average across all year.
 
@@
 
-**Overall composite.** Unpredictability will be based on the same PCA as used for material deprivation, involving the same indicators.
-we operationalize unpredictability as within-person variability in material deprivation over time [@ugarte_2023; @young_2020].
-Specifically, we will compute unpredictability by dividing the within-subject standard deviation of material deprivation by the within-subject mean-level of material deprivation. 
-Dividing by the mean accounts for the fact that the mean and standard deviation in income have been found to be strongly negatively correlated, indicating that people with lower incomes tend to experience less variability in income [@li_2018; @young_2023].
+**Unpredictability composite.** We will compute a composite measure of unpredictability based on the yearly unweighted averages of perceived scarcity and the income-to-needs ratio (reverse coded), as described above.
+We will compute unpredictability as the within-person standard deviation in material deprivation across years [@ugarte_2023; @walasek_2023; @young_2020].
+The mean and standard deviation in income have been found to be strongly negatively correlated, indicating that people with lower incomes tend to experience less variability in income [@li_2018; @young_2023].
 For that reason, the standard deviation alone has been called into question as a measure of adversity, as the same fluctuation in income can have a greater relative impact for people close to the poverty line than for people with high incomes.
+We will address this issue by including the covariance between material deprivation (the mean) and unpredictability (the standard deviation) into the SEM model (see below).
 
@@
 
-All materials and scripts for the cognitive tasks can be found at [https://github.com/StefanVermeent/liss_wm_profiles_2023/tree/master/materials](https://github.com/StefanVermeent/liss_wm_profiles_2023/tree/master/materials).
-Prior to collecting LISS data, we conducted a pilot study through Prolific in a Dutch sample of N = 100.
+The WM tasks were all part of the newly collected data.
+All materials and scripts for the cognitive tasks can be found at [https://stefanvermeent.github.io/liss_wm_profiles_2023/materials/README.html](https://stefanvermeent.github.io/liss_wm_profiles_2023/materials/README.html).
+Prior to collecting LISS data, we conducted a pilot study among in a Dutch sample (*N* = 100) through Prolific Academic.
 The main goals of this pilot study were to collect participant feedback (e.g., difficulty of instructions, whether we included sufficient breaks) and to analyze performance and correlations between tasks.
@@
 **Operation Span Task.** The Operation Span Task (Figure 2A) is a common measure of WM capacity [@conway_2005; @wilhelm_2013].
-In this task, participants alternated between a primary memorization task and a secondary processing task.
-On each trial, participant's task was to memorize a sequence of letters in the correct order (from a set of 12 letters).
-Each letter was presented for 1,000 ms in the center of the screen.
-Next, participants saw a simple mathematical equation including the outcome.
-Their task was to indicate whether the outcome was correct or incorrect by pressing either the 'a' or 'l' key on their keyboard.
-The equations always contained one addition or subtraction, with numbers ranging between one and 10.
-Outcomes were always positive integers.
-On each trial, participants had to memorize between four and six letters, with each set size repeated three times.
-At the end of each sequence, all letters were presented in a 3X4 grid, and participants clicked the letters in the correct order.
+In this task, participants alternate between a primary memorization task and a secondary processing task.
+On each trial, the task is to memorize a sequence of letters in the correct order (from a set of 12 letters).
+Each letter is presented for 1,000 ms in the center of the screen.
+Next, participants see a simple mathematical equation including the outcome.
+Their task is to indicate whether the outcome is correct or incorrect by pressing either the 'a' or 'l' key on their keyboard.
+The equations always contain one addition or subtraction, with numbers ranging between one and 10.
+Outcomes are always positive integers.
+On each trial, participants have to memorize between four and six letters, with each set size repeated three times.
+At the end of each sequence, all letters are presented in a 3 x 4 grid, and participants click the letters in the correct order.
 
-Participants first practiced the letter task (three times), then the math task (eight times), and then the full task (three times).
-If they performed at or below chance, they  had the opportunity to either repeat a part or advance to the next part.
-After practicing, participants completed 12 test trials.
+Participants first practice the letter task (three times), then the math task (eight times), and then the full task (three times).
+If they perform at or below chance, they have the opportunity to either repeat a part or advance to the next part.
+After practicing, participants complete 12 test trials.
+We will compute an operation span score by calculating the proportion of letters recalled in the correct sequential position across trials [@conway_2005].
 
-**Rotation Span Task.** The Rotation Span Task (Figure 2B) was similar to the Operation Span Task and was adopted from @wilhelm_2013.
-On each trial, participants' task was to memorize the orientation of a sequence of arrows in the correct order.
-Arrows could take on eight different orientations, with steps of 45$\circ$.
-Each arrow was presented for 1,000 ms in the center of the screen. 
-Next, participants saw a capital 'G' or 'F' that was rotated at one of eight different orientations, with steps of 45$\circ$.
-Their task was to indicate whether the letter was mirrored or not.
-On each trial, participants had to memorize between two and five arrows, with each set size repeated three times.
-At the end of each sequence, all arrows were presented simultaneously, and participants clicked the arrows in the correct order.
+**Rotation Span Task.** The Rotation Span Task (Figure 2B) is similar to the Operation Span Task and is adopted from @wilhelm_2013.
+On each trial, the task is to memorize the orientation of a sequence of arrows in the correct order.
+Arrows can take on eight different orientations, with increments of 45$\circ$.
+Each arrow is presented for 1,000 ms in the center of the screen. 
+Next, participants see a capital 'G' or 'F' that is rotated at one of eight different orientations, with increments of 45$\circ$.
+Their task is to indicate whether the letter is mirrored or not.
+On each trial, participants have to memorize between two to five arrows, with each set size repeated three times.
+At the end of each sequence, all arrows are presented simultaneously, and participants click the arrows in the correct order.
 
-Participants first practiced the arrow task (three times), then the letter task (eight times), and then the full task (three times).
-If they performed at or below chance, they  had the opportunity to either repeat a part or advance to the next part.
-After practicing, participants completed 12 test trials.
+Participants first practice the arrow task (three times), then the letter task (eight times), and then the full task (three times).
+If they perform at or below chance, they have the opportunity to either repeat a part or to advance to the next part.
+After practicing, participants complete 12 test trials.
+We will compute a rotation span score by calculating the proportion of arrows recalled in the correct sequential position across trials [@conway_2005.
 
-**Binding-Updating Task.** The Binding-Updating task (Figure 2C) was adopted from [XXX].
-On each trial, participants saw a 3x3 grid, with a fixation cross in the central cell.
-After 1,000 ms, they were presented with a sequence of numbers (0-9) in random locations of the grid.
-Each new number was presented for 1,500 ms, after which it disappeared for 500 ms before the next number was presented.
-The participants' task was to remember the last number they saw in each location.
-Memory set sizes (i.e., the number of unique locations in the grid) ranged between three and five.
-On half of the trials, only one number was presented in each location.
-These constituted the binding-only trials.
-On the other half of the trials, some letters were presented in the same location as previous numbers, requiring mentally replacing the old number with the new number.
-These constituted the updating trials.
-We used two, three, and four updating steps, each repeated in combination with the different set sizes.
-At the end of the trials, participants indicated which letter they saw last in each location in random order.
+**Binding-Updating Task.** The Binding-Updating task (Figure 2C) is adopted from [XXX].
+On each trial, participants see a 3 x 3 grid, with a fixation cross in the central cell.
+After 1,000 ms, they are presented with a sequence of numbers (0-9) in random locations of the grid.
+Each new number is presented for 1,500 ms, after which it disappears for 500 ms before the next number is presented.
+The task is to remember the last number they see in each location.
+Memory set sizes (i.e., the number of unique locations in the grid) ranges between three and five.
+On half of the trials, only one number is presented in each location.
+These constitute the binding trials.
+On the other half of the trials, some letters are presented in the same location as previous numbers, requiring mentally replacing the old number with the new number.
+These constitute the updating trials.
+We use two, three, and four updating steps, each repeated in combination with the different set sizes.
+At the end of the trials, participants indicate which letter they saw last in each location in random order.
 
-Participants first completed four practice trials.
-If they performed at or below chance, they had the opportunity to either repeat the practice trials or advance to the full task.
-After practicing, they completed 18 test trials.
-We will compute a binding score by calculating the overall accuracy (%) across trials with zero updating steps.
-We will compute an updating score by calculating the overall accuracy (%) across trials containing updating steps.
+Participants first complete four practice trials.
+If they perform at or below chance, they have the opportunity to either repeat the practice trials or to advance to the actual task.
+After practicing, they complete 18 test trials.
+We will compute a binding score by calculating the overall recall accuracy (%) across trials with zero updating steps.
+We will compute an updating score by calculating the overall recall accuracy (%) of updated letters (disregarding non-updated letters) across trials containing updating steps.
 
@@
 
-We received ethical approval from the first institutions ethical board.
+We received ethical approval from the first author's institutional ethical board.
 Upon starting the study, participants were informed that the study could only be completed on a laptop or desktop PC.
+If they attempted to start the study on a tablet or smartphone, they were unable to advance and prompted to switch to a suitable device.
 Participants started with the WM tasks, which on average took between 20 and 25 minutes.
@@
 If participants left fullscreen mode at any moment during the tasks, they saw instructions at the top of their screen that allowed them to return to fullscreen mode.
-the order of the WM tasks was counterbalanced, and participants had the opportunity to take breaks at regular intervals.
+The order of the WM tasks was counterbalanced, and participants had the opportunity to take breaks at regular intervals.
 
@@
 These data will only be made available to the first author after Stage 1 acceptance, as stipulated in a signed contract with LISS.
-During planning of the study, the first author received access to the LISS data archive and inspected three waves of the LISS data containing the items about neighborhood safety and crime exposure, as well as the three most recent monthly data collections containing basic demographic info.
-The reason was to discover how many people who completed the waves a few years back were still active in the panel (i.e., to see if we could reasonably create a link between these previous waves and newly collected data).
-All data access events were automatically detected and logged on the GitHub repository using the 'projectlog' R package [@vermeent_2023].
-We took the following measures to prevent bias: 1) we randomly shuffled the participant IDs in each data set using 'projectlog', so that we were unable to link participant data between (waves of) studies; 2) we did not inspect any of the measures that will be part of our adversity composites; 3) we did not know which participants would be selected for the new LISS study; 4) the primary analyses will be based on composite measures that combine measures from these previous studies with data that were collected at a later timepoint.
+During planning of the study, the first author accessed to the LISS data archive and inspected three waves of the LISS data containing the items about neighborhood safety and crime exposure, as well as the three most recent monthly data collections containing basic demographic info.
+The purpose was to ascertain the number of individuals who had finished the previous waves a few years ago and were presently still participating in the panel (i.e., to see if we could reasonably create a link between these previous waves and newly collected data).
+All data access events were automatically detected and logged on the GitHub repository using the *projectlog* R package [@vermeent_2023].
+We took the following measures to prevent bias: 1) we randomly shuffled the participant IDs in each data set using the *projectlog* R package, so that we were unable to link participant data between (waves of) studies; 2) we did not inspect any of the measures that will be part of our adversity composites; 3) we did not know which participants would be selected for the new LISS study; 4) the primary analyses will be based on composite measures that combine measures from these previous studies with data that were collected at a later timepoint.
 
@@
 See Figure 1B for an overview of the model specification.
-We will fit a single model containing all adversity measures using the *lavaan* package [@rosseel_2012].
+We will fit a single model containing all adversity measures using the *lavaan* R package [@rosseel_2012].
+We will use robust maximum likelihood estimation in case any variable is non-normally distributed.
 Missing data will be handled using full information maximum likelihood (FIML).
-If participants are from the same household, this clustering within families will be accounted for using the *lavaan.survey* package [@oberski_2014].
+If participants are from the same household, this clustering within families will be accounted for using the *lavaan.survey* R package [@oberski_2014].
 
@@
 We will estimate the effect of each adversity type (dashed lines in Figure 1B) through regression analyses.
-Each association is controlled for: (1) age in years ; (2) the quadratic effect of age; (2) environmental noise ("How noisy was your environment during the memory tasks", rated on a scale of 1 (very little noise) to 5 (a lot of noise)); (3) two items measuring interruptions ("Where you at any moment interrupted during the memory tasks?" and "Did you at any moment during the memory tasks leave your computer?", rated as yes or no).
+Each association will be controlled for: (1) age in years ; (2) the quadratic effect of age; (2) environmental noise ("How noisy was your environment during the memory tasks", rated on a scale of 1 (very little noise) to 5 (a lot of noise)); (3) two items measuring interruptions ("Where you at any moment interrupted during the memory tasks?" and "Did you at any moment during the memory tasks leave your computer?", rated as yes or no).
 Goodness of fit will be assessed using the comparative fit index (CFI) and the root mean square error of approximation (RMSEA).
-CFI values \> .90 and RMSEA values \< .08 will be interpreted as acceptable model fit, and CFI values \> .95 and RMSEA values \≤ .06 as good model fit.
+CFI values \> .90 and RMSEA values \< .08 will be interpreted as acceptable model fit, and CFI values \> .95 and RMSEA values \≤ .06 as good model fit [@hu_1999].
 
 We anticipate that we may have to optimize the model further in case of bad model fit, and will therefore estimate the model in two steps to prevent bias.
-First, we will construct the measurement model only, without including the adversity measures.
+First, we will construct the measurement model of WM, without including the adversity measures.
 This step will be carried out prior to accessing any of the adversity measures.

### 2024-01-24 README.Rmd
new README

@@
 ```{r results ='asis', echo = FALSE}
-gert::git_tag_list() |> 
+milestones <- gert::git_tag_list() |> 
   mutate(
@@
     ) |>
-  arrange(desc(time)) |> 
-  glue_data("  - **[{time}]({link_to_git}): {description}**
+  arrange(desc(time))
+
+# Write scripts including code to .projectlog folder
+milestones |> 
+  pwalk(function(name,commit,author,time,description,hash,code,link_to_git){
+
+    if(name == "Data Access"){
+      code <- code |> 
+        str_replace_all(string = _, "\\|> ", "\\|>\n") |> 
+        str_remove(string = _, "^code ")
+
+      url <- gert::git_remote_list()$url |> 
+        str_remove(string = _, "\\.git$") |> 
+        paste0(paste0("/commit/", commit))
+
+      #  url <- "<anonymized repository>"
+
+      script <-
+        glue(
+          "
+          ### Date: {time}\n
+          ### Description: {description}\n\n
+          ### For more information on this commit, see the README file, or go to {url}\n
+          ### Below is the full code that was used to access the data:\n\n
+          {code}"
+        )
+
+      writeLines(script, con = paste0(".projectlog/", commit, ".R"))
+    }
+  })
+
+# Print milestones
+glue_data(milestones, "  - **[{time}]({link_to_git}): {description}**
         - **Milestone:** {name}
@@
     ")
+
 ```

### 2024-01-24 scripts/1_sim_pipeline/4_sem.R

@@ -14,7 +14,7 @@ load("data/full_data_sim.RData")
 # Some children are clustered within families.
 # We account for this using the lavaan.survey package
 
-cluster_design <- survey::svydesign(ids=~nohouse_encr, prob=~1, data = data_clean_sim)
+cluster_design <- survey::svydesign(ids=~nohouse_encr, prob=~1, data = clean_data_sim)
 
 
 # 3. Model specification(s) --------------------------------------------------
@@ -35,7 +35,7 @@ mod_meas <-
 
 fit_meas1 <- lavaan::sem(
   model = mod_meas,
-  data = data_clean_sim,
+  data = clean_data_sim,
   estimator = "MLR", # Robust SEs in case of non-normal data
   missing = "ML",    # FIML for missing data
   orthogonal = TRUE,
@@ -68,3 +68,65 @@ mod_full <-
   upd_l ~ mat_dep + unp + threat + age + interrupt + leave
 '
 
+
+
+
+# Check model soundness ---------------------------------------------------
+
+mod_full_pop <-
+  '
+  # Measurement Model
+
+  ## Latent factors
+  wmc_l      =~ 1*bu_bind_acc + 0.7*rspan_cap + 0.7*ospan_cap + 0.7*bu_upd_acc
+  upd_l      =~ 1*bu_upd_acc
+
+  ## Covariances
+  bu_upd_acc ~~ 0*bu_upd_acc
+
+  mat_dep        ~~ 0.3*threat
+
+  # Structural Model
+  wmc_l ~ 0.1*mat_dep + 0.1*unp + 0.1*threat + 0.1*age + 0.1*interrupt + 0.1*leave
+  upd_l ~ 0.1*mat_dep + 0.1*unp + 0.1*threat + 0.1*age + 0.1*interrupt + 0.1*leave
+'
+
+
+sim_data <- lavaan::simulateData(
+  model = mod_full_pop,
+  sample.nobs = 800
+) |>
+  mutate(across(everything(), scale))
+
+
+mod_full_sample <-
+  '
+  # Measurement Model
+
+  ## Latent factors
+  wmc_l      =~ 1*bu_bind_acc + rspan_cap + ospan_cap + bu_upd_acc
+  upd_l      =~ 1*bu_upd_acc
+
+  ## Covariances
+  bu_upd_acc ~~ 0*bu_upd_acc
+
+  mat_dep        ~~ threat
+
+  # Structural Model
[…]

### 2024-01-24 README.Rmd
fix git url issue in README

@@
 
+branch <- gert::git_branch_list() |> filter(local == TRUE) |> pull(name)
+
 # Print milestones
@@
         - **Data MD5 hash**: {hash}
-        - [Link to code snippet](https://https://github.com/StefanVermeent/liss_wm_profiles_2023/tree/main/.projectlog/{commit}.R)
+        - [Link to code snippet](https://https://github.com/StefanVermeent/liss_wm_profiles_2023/blob/{branch}/.projectlog/{commit}.R)
         \n

### 2024-01-24 README.Rmd
fix additional issue with git url

@@
         - **Data MD5 hash**: {hash}
-        - [Link to code snippet](https://https://github.com/StefanVermeent/liss_wm_profiles_2023/blob/{branch}/.projectlog/{commit}.R)
+        - [Link to code snippet](https://github.com/StefanVermeent/liss_wm_profiles_2023/blob/{branch}/.projectlog/{commit}.R)
         \n

### 2024-01-29 supplement/supplement.qmd
supplemental materials

---
bibliography: ../manuscript/references.bib
csl: ../manuscript/apa.csl
format: 
  docx:
    reference-doc: ../manuscript/reference-doc.docx
    toc: true
    toc-location: left
    toc-title: Table of Contents
output:
  officedown::rdocx_document:
    page_margins:
      bottom: 1
      footer: 0
      gutter: 0
      header: 0.5
      left: 1
      right: 1
      top: 1
    plots:
      align: center
      caption:
        pre: 'Figure '
        sep: '. '
        style: Image Caption
    tables:
      caption:
        pre: 'Table '
        sep: '. '
        style: Table Caption
  pdf_document: default
  word_document: default
editor: 
  markdown: 
    wrap: sentence
---

```{r include = FALSE}
library(tidyverse)
library(flextable)
library(officer)

load('staged_results.RData')

# set up flextable for tables
set_flextable_defaults(
  font.family = "Times",
  font.size = 10,
  font.color = "black",
  line_spacing = 1,
  padding.bottom = 1,
  padding.top = 1,
  padding.left = 1,
  padding.right = 1
)

knitr::opts_chunk$set(
  echo = F,
  fig.align = "center",
  fig.pos = "!t", 
  out.extra = "",
  fig.show = "asis",
  message = FALSE,
  tab.topcaption = T,
  warning = FALSE
)
```

\pagebreak

# 1. Pilot study

A total of 99 Dutch adolescents (mean age = `r pilot_demo$m_age`, SD = `r pilot_demo$sd_age`, range = [`r pilot_demo$min_age`, `r pilot_demo$max_age`]) participated in the Pilot study via Prolific.
The main goal of the Pilot study was to obtain feedback on the tasks (e.g., difficulty, clarity of instructions), and to explore bivariate correlations between the measures.
Participants completed the same tasks as in the main study: (1) Operation Span Task, (2) Rotation Span Task, and (3) Binding-Updating Task.
The Rotation Span Task was administered in a second session, and was completed by a subsample of 50 participants.
All three tasks followed the exact same procedure as in the main study.

In addition, participants completed measures of material deprivation, neighborhood threat, and unpredictability.
These measures differed from the more comprehensive measures used in the main study, and were included to obtain quick, descriptive estimates.
Material deprivation was measured using seven items about perceived level of available resources.
Unpredictability was measured using a scale of perceived unpredictability [@mittal_2015; @young_2018].
Neighborhood violence exposure was measured using the Neighborhood Violence Scale [NVS; @frankenhuis_2018; @frankenhuis_deVries_2020] as well as two items measuring involvement in fights.
Participants responded to items of all questionnaires on a scale of 1 (never true) to 5 (very often true).
Finally, participants provided feedback on the difficulty of the tasks and the clarity of the task instructions.

Pilot data were collected sequentially to allow for intermediate changes to instructions based on participants' feedback.
The first session (including the Operation Span Task and the Binding-Updating Task) took approximately 35 minutes to complete, and participants were paid 5.25 GBP.
The second session (including the Rotation Span Task) took approximately 9 minutes to complete, and participants were paid 1.50 GBP.

Table S1 presents bivariate correlations among the WM tasks, and between the WM tasks and measures of adversity.
The WM tasks correlated moderately to strongly with each other.
The strongest correlation was between the Binding and Updating score (.80).
This is not surprising given that both scores are derived from the same task, and shows the importance of accounting for this association in the model.
Neither unpredictability nor material deprivation were significantly associated with performance on any of the WM tasks.
However, higher levels of experienced neighborhood threat were associated with lower performance on the Binding and Updating Task.
Note that these associations were based on raw task performance and not on latent estimates.

```{r}
pilot_table1
```

# 2. Study Design

Table S2 presents the study design plan, using the table provided by PCI-RR.

```{r}
table_s2 <- tibble(
  `Research question` = c(
    "1. what is the association between adversity and WM capacity?", 
    "2. what is the association between adversity and WM updating after accounting for WM capacity?", 
    "3. Are the directions and strengths of these associations similar or different for neighborhood threat, material deprivation, and unpredictability?"),
  `Hypotheses`        = c(
    "Deficit perspectives predict a negative association between all three types of adversity and WM capacity. Negative associations for one or two types of adversity would be partial support. In addition, deficit perspectives predict that none of the associations between adversity and WM capacity will be positive.",

    "Adaptation-based perspectives predict that the association between adversity and WM updating (after accounting for WM capacity) is either (1) positive, regardless of the association with WM capacity, or (2) practically equivalent to zero in combination with a negative association between adversity and WM capacity.

Deficit perspectives predict that the association between adversity and WM updating is negative.",

    "We have two expectations based on prior studies. First, we expect that the association between material deprivation and WM capacity is more strongly negative than the associations with unpredictability and neighborhood threat.
Second, we expect to find associations consistent with adaptations in WM updating (see RQ2) with either unpredictability or neighborhood threat, but not with material deprivation."
    ),
  `Sampling plan`     = c(
    "We are collecting data of 800 participants in the Dutch Longitudinal Internet studies for the Social Sciences (LISS) panel. First, we will use data that were previously collected in LISS. Second, we will use new data that we collected ourselves in LISS. Data collection started on October 2nd and is expected to be completed in February 2024. We signed a contract with LISS stipulating that we will receive access to the data only after Stage 1 acceptance of this Registered Report.

To ensure sufficient representation of people from lower socioeconomic backgrounds, roughly half the total sample will be sampled from participants who reported one or more of the following at least once in the three years: (1) a monthly income < €1,500, (2) HAVO or VWO as highest completed education (which are the two highest levels in Dutch secondary education), or (3) a score of 4 or lower on the 'ladder of life'", 
    NA, 
    NA),
  `Analysis plan`     =  c(
    "We will fit a single structural equation model (SEM) containing all adversity measures. We will use robust maximum likelihood estimation in case any variable is non-normally distributed. Missing data will be handled using full information maximum likelihood (FIML). If participants are from the same household, this clustering within families will be accounted for.

WM capacity will be estimated as a latent factor loading on all outcome measures. In addition, we will estimate WM updating as a latent factor capturing residual variance in the updating measure. We will estimate the effect of each adversity type through regression analyses. Each association will be controlled for: (1) age in years; (2) the quadratic effect of age; (2) environmental noise; (3) two items measuring interruptions.

We will estimate the model in two. First, we will construct the measurement model of WM, without including the adversity measures. Once we obtain at least acceptable model fit, we will access and add the adversity measures to the model. We will control for multiple testing using the false discovery rate 
We will use two one-sided tests (TOST) equivalence testing to test whether small effects—which we define as standardized effects between -.10 and .10—are practically equivalent, which we will interpret as evidence for intact performance.
", 
    NA, 
    NA),
  `Rationale for deciding the sensitivity of the test for confirming or disconfirming the hypothesis` = c(
    "We based our power analysis on simulations reported by Kretzschmar ad Gignac (2019), determining the required sample size to detect a small effect size (β = 0.1) with at least 90% power at α = 0.05.
Assuming a reliability of at least 0.7 (which is typical for WM tasks with a number of trials similar to ours; e.g., Wilhelm, et al., 2013), we would require a sample size of *N* = 730. Anticipating exclusions, we decided to include 800 participants.", 
    NA, 
    NA),
  `Interpretation given different outcomes` = c(
    "Contrary to predictions of deficit perspectives, we might find that all associations between adversity and WM capacity are either practically equivalent or positive. This would suggest that WM capacity is either unaffected or even enhanced by adversity.

If we find both a non-significant association and practical non-equivalence, we will conclude that our data neither support nor refute either framework.
",

    "Contrary to predictions of adaptation-based perspectives, we might find that the association between adversity and WM updating is negative. This would suggest that WM updating is impaired by adversity.

Contrary to predictions of adaptation-based perspectives, we might find a practically equivalent association with adversity for both WM capacity and updating. This would suggest that WM is unaffected by adversity.

If we find both a non-significant association and practical non-equivalence, we will conclude that our data neither support nor refute either framework.",

    "We might find that the association between threat or unpredictability with WM capacity is more strongly or equally strongly negative than with material deprivation. This would suggest that threat or unpredictability are more strongly associated with WM capacity than material deprivation.

We might also find that material deprivation, but not unpredictability or neighborhood threat, is positively associated with WM updating. This would suggest that an enhanced updating ability has an adaptive benefit for individuals experiencing material deprivation."
    ),
  `Theory that could be shown wrong by the outcomes` = c(
    "Theoretically, our analyses directly compare evidence in favor of deficit and adaptation-based perspectives. Both are established frameworks generating predictions that extend to other cognitive abilities beyond WM. Therefore, the current study could neither confirm nor disconfirm the frameworks in general.

However, our findings could be (partially) inconsistent with predictions derived from both frameworks. Deviating findings for RQ1 or RQ 2 would require revising theoretical predictions about the specific WM abilities that are adapted to/impaired by adversity.

In both cases, it would suggest that both frameworks need to be explicit in how they distinguish between different WM components.",

    NA, 
    "The hypotheses specified for RQ3 do not directly offer (non-) support for either framework. However, finding different patterns than hypothesized here would be inconsistent with findings of prior studies."),
) |> 
  flextable() |> 
  merge_at(i = 1:3, j = 3, part = "body") |> 
  merge_at(i = 1:3, j = 4, part = "body") |>
  merge_at(i = 1:3, j = 5, part = "body") |>
  merge_at(i = 1:2, j = 7, part = "body") |>  
  add_header_row(
    values = " ",
    colwidths = 7
  ) |>
  flextable::compose(
    i = 1, j = 1,
    as_paragraph(as_b("Table S2. "), "Study design plan."),
    part = "header"
  ) |> 
  border(border.top = fp_border(color = "black"), part = "body") |> 
  border(border.left = fp_border(color = "black"), part = "body") |> 
  border(border.right = fp_border(color = "black"), part = "body") |> 
  border(border.bottom = fp_border(color = "black"), part = "body") |> 
  border(i = 2, border.left = fp_border(color = "black"), part = "header") |>
  border(i = 2, j = 7, border.right = fp_border(color = "black"), part = "header") |>
  border(border.bottom = fp_border(color = "black"), part = "body") |> 
  border(i = 1, j = 1, border.left =  fp_border_default(style = "none", width = 0), part = "header") |>
  border(i = 1, j = 7, border.right =  fp_border_default(style = "none", width = 0), part = "header") |> 
  border(i = 1, border.top =  fp_border_default(style = "none", width = 0), part = "header") |> 
  valign(valign = "top") |> 
  width(width = 1)

table_s2
```

### 2024-01-29 supplement/scripts/stage_supp.R

@@ -1,0 +1,92 @@
+library(flextable)
+library(tidyverse)
+
+load("data/pilot_full_data.RData")
+source("scripts/0_custom_functions/binom_test.R")
+source("scripts/0_custom_functions/corr_plot.R")
+
+
+# set up flextable for tables
+set_flextable_defaults(
+  font.family = "Times",
+  font.size = 10,
+  font.color = "black",
+  line_spacing = 1,
+  padding.bottom = 1,
+  padding.top = 1,
+  padding.left = 1,
+  padding.right = 1
+)
+
+# 1. Pilot Study ----------------------------------------------------------
+
+pilot_demo <-
+  list(
+    m_age = pilot_data_full$dems_age |> mean(na.rm = T) |> round(1),
+    sd_age = pilot_data_full$dems_age |> sd(na.rm = T) |> round(1),
+    max_age = pilot_data_full$dems_age |> max(na.rm = T) |> round(1),
+    min_age = pilot_data_full$dems_age |> min(na.rm = T) |> round(1)
+  )
+
+pilot_exclusions <-
+  list(
+    acc_ospan_sec_n = pilot_data_full |> filter(ospan_sec_acc*100 < gbinom(45, 0.50)) |> nrow(),
+    acc_rspan_sec_n = pilot_data_full |> filter(rspan_sec_acc*100 < gbinom(45, 0.50)) |> nrow()
+  )
+
+pilot_clean <- pilot_full_data |>
+  filter(
+    ospan_sec_acc*100 > gbinom(45, 0.50),
+    (rspan_sec_acc*100 > gbinom(45, 0.50) | is.na(rspan_sec_acc))
+  )
+
+pilot_table1 <- pilot_clean |>
+  select(ospan_cap, rspan_cap, binding_number, updating_number, pcunp_mean, vio_comp, ses_subj_comp) |>
+  corr_table(
+    stats = c("mean", "sd", "median", "min", "max", "skew", "kurtosis"),
+    numbered = TRUE,
+    c.names = c(
+      "Operation Span Task",
+      "Rotation Span Task",
+      "Binding Task",
+      "Updating Task",
+      "Unpredictability",
+      "Threat",
+      "Material deprivation"
+    )
+  ) |>
+  add_row(Variable = "WM tasks", .before = 1) |>
+  add_row(Variable = "Adversity", .before = 6) |>
+  flextable() |>
+  set_header_labels(Variable = "") |>
[…]

### 2024-01-29 scripts/0_custom_functions/corr_plot.R
correlation plot function

@@ -30,8 +30,8 @@ corr_table <- function (data, sample_size = T, use = "pairwise", method = "pears
     ps <- 
       corr_data$p %>% 
       as.data.frame() %>% 
-      mutate(across(.cols = everything(), .fns = ~ my_ifelse(as.numeric(.x) <  .01, "", as.character(.x)))) %>% 
-      mutate(across(.cols = everything(), .fns = ~ my_ifelse(as.numeric(.x) <= .05 & !is.na(as.numeric(.x)), "", .x))) %>% 
+      mutate(across(.cols = everything(), .fns = ~ my_ifelse(as.numeric(.x) <  .01, "**", as.character(.x)))) %>% 
+      mutate(across(.cols = everything(), .fns = ~ my_ifelse(as.numeric(.x) <= .05 & !is.na(as.numeric(.x)), "*", .x))) %>% 
       mutate(across(.cols = everything(), .fns = ~ my_ifelse(as.numeric(.x) >  .05 & !is.na(as.numeric(.x)), " ", .x))) %>% 
       as.matrix()
     

### 2024-01-30 supplement/supplement.qmd
Final Stage 1 supplemental materials

@@
 Unpredictability was measured using a scale of perceived unpredictability [@mittal_2015; @young_2018].
-Neighborhood violence exposure was measured using the Neighborhood Violence Scale [NVS; @frankenhuis_2018; @frankenhuis_deVries_2020] as well as two items measuring involvement in fights.
+Neighborhood violence exposure was measured using the Neighborhood Violence Scale [NVS\; @frankenhuis_deVries_2020; @frankenhuis_2018] as well as two items measuring involvement in fights.
 Participants responded to items of all questionnaires on a scale of 1 (never true) to 5 (very often true).
@@
   `Hypotheses`        = c(
-    "Deficit perspectives predict a negative association between all three types of adversity and WM capacity.
-Negative associations for one or two types of adversity would be partial support.
-In addition, deficit perspectives predict that none of the associations between adversity and WM capacity will be positive.",
+    "Deficit frameworks predict a negative association between all three types of adversity and WM capacity as well as WM updating.
+This follows from the hypothesis that adversity leads to broad WM deficits.
+Deficit frameworks are partially supported if we find negative associations with only one (or two) types of adversity.
+",
 
-    "Adaptation-based perspectives predict that the association between adversity and WM updating (after accounting for WM capacity) is either (1) positive, regardless of the association with WM capacity, or (2) practically equivalent to zero in combination with a negative association between adversity and WM capacity.
+    "Within adaptation-based frameworks, theories make two predictions.
 
-Deficit perspectives predict that the association between adversity and WM updating is negative.",
+First, if adaptive processes enhance WM updating and there are no impairment processes operating, we can expect a positive association between adversity and WM updating. 
+Second, if, adaptive processes operate in concert with general impairment processes, we can expect intact WM updating in combination with lowered WM capacity.
+If neither impairment nor adaptative processes are operating, we can expect both WM updating and capacity to be intact.",
 
     "We have two expectations based on prior studies.
-First, we expect that the association between material deprivation and WM capacity is more strongly negative than the associations with unpredictability and neighborhood threat.
-Second, we expect to find associations consistent with adaptations in WM updating (see RQ2) with either unpredictability or neighborhood threat, but not with material deprivation."
+First, we expect the association between material deprivation and WM capacity to be more negative than the associations with unpredictability and neighborhood threat.
+Second, we expect WM updating to be associated with unpredictability and neighborhood threat, but not with material deprivation."
     ),
@@
 ```
+
+\pagebreak
+
+# References
+
+::: {#refs}
+:::
+
+\pagebreak

### 2024-01-30 manuscript/manuscript.qmd
Stage 1 manuscript submitted to PCI-RR

@@
 The finding that updating may be left intact or even enhanced after exposure to adversity exemplifies emerging theoretical frameworks grounded in adaptive reasoning that are complementary to deficit frameworks [@frankenhuis_2013; @ellis_2017; @ellis_2022; @frankenhuis_2020].
-Adaptation frameworks assume that developmental processes tailor an individual's cognitive abilities to the unique challenges and opportunities posed by their environment.
+Adaptation-based theories assume that developmental processes tailor an individual's cognitive abilities to the unique challenges and opportunities posed by their environment.
 The link between adversity and cognitive abilities is further assumed to be specific; as different types of adversity (e.g., threat vs.
@@
 
-With deficit frameworks focusing on WM capacity and adaptation frameworks on WM updating, we may wonder how capacity and updating are related to each other.
-Performance on tasks measuring WM capacity and updating tend to be substantially correlated [in the range of .20-.50\; @frischkorn_2022].
+With deficit theories focusing on WM capacity and adaptation-based theories on WM updating, we may wonder how capacity and updating are related to each other.
+Performance on tasks measuring WM capacity and updating tend to be substantially correlated [in the range of .20-.50\; @frischkorn_2022; @loffler_2022].
 This overlap appears to stem from shared demands of both types of tasks, in particular the need to create and maintain arbitrary bindings [@gruszka_2017; @oberauer_2009; @wilhelm_2013].
@@
 
-We will evaluate evidence for deficit and adaptation frameworks.
-Rather than banking on specific predictions, we will describe how each of four different combinations of lowered, enhanced, and intact WM capacity and updating performance would be consistent with deficit and adaptation frameworks (For a summary, see Figure 1A).
-Crucially, we assume that deficit and adaptation processes can operate simultaneously [@frankenhuis_2020], meaning that we could find support or a lack of support for both frameworks in the same model.
-Lowered performance on WM capacity or updating, but not enhanced or intact performance, would be consistent with deficit frameworks.
-This follows from previous literature hypothesizing broad WM deficits [@farah_2006; @sheridan_2020].
-Two different patterns would be consistent with adaptation frameworks:
-First, if performance on WM updating is enhanced, regardless of WM capacity performance.
-Second, if performance on WM updating is intact, but only in combination with lowered WM capacity performance.
-If both updating and capacity are intact, this would instead suggest that the WM system is unaffected by adversity; neither impairment nor adaptative processes have shaped WM.
-That pattern would be inconsistent with both deficit and adaptation frameworks.
+We will evaluate evidence for deficit- and adaptation-based frameworks (see Figure 1A for a visual summary, and supplemental Table S2 for the study design plan). 
+Crucially, as deficit and adaptation processes can operate in concert [@frankenhuis_2020], we could find support (or lack thereof) for both frameworks in the same model.
+
+Deficit frameworks predict a negative association between all three types of adversity and WM capacity as well as WM updating. 
+This follows from the hypothesis that adversity leads to broad WM deficits [@farah_2006; @sheridan_2020]. 
+Deficit frameworks are partially supported if we find negative associations with only one (or two) types of adversity.
+
+Within adaptation-based frameworks, theories make two predictions. 
+First, if adaptive processes enhance WM updating and there are no impairment processes operating, we can expect a positive association between adversity and WM updating. 
+Second, if, adaptive processes operate in concert with general impairment processes, we can expect intact WM updating in combination with lowered WM capacity. 
+If neither impairment nor adaptative processes are operating, we can expect both WM updating and capacity to be intact.
+
+We also have two expectations based on prior studies. 
+First, we expect the association between material deprivation and WM capacity to be more negative than the associations with unpredictability and neighborhood threat. 
+This follows from findings showing that cognitive abilities are more negatively associated with cognitive deprivation than threat [@salhi_2021; @sheridan_2020].
+Although cognitive and material deprivation are distinct types of deprivation, they tend to be correlated, and are both associated with limited access to resources that stimulate cognitive development and functioning [@bradley_2001; @lurie_2024; @rosen_2019].
+Therefore, we expect that their associations with WM will have comparable effect sizes. 
+Second, researchers have hypothesized that WM updating is particularly adaptive in unpredictable and threatening environments, as it facilitates keeping track of unpredictable changes and sudden threats.
+Therefore, we expect WM updating to be associated with unpredictability and neighborhood threat, but not with material deprivation [@young_2018; but see @young_2022].
 
@@
 We will sum across these six items, with higher scores indicating more financial difficulties.
+Finally, we will create a single measure of perceived scarcity for each year by separately standardizing the three measures and averaging them.
 
@@
 If monthly household income is missing (or zero) for an entire year for a participant, we will use, if available, the yearly net household income they reported in the LISS archive ([https://doi.org/10.57990/1gr4-bf42](https://doi.org/10.57990/1gr4-bf42)), dividing it by 12 to obtain a monthly estimate.
-First, we average monthly reported income across years for each participant.
-Then, we divide the average income per year by the *poverty threshold*, as determined by Statistics Netherlands (CBS) [@brakel_2023\; CBS, personal communication, December 15, 2023].
+First, we will divide the average income per year by the *poverty threshold*, as determined by Statistics Netherlands (CBS) [@brakel_2023\; CBS, personal communication, December 15, 2023].
 The poverty threshold changes across time to account for purchasing power.
@@
 Likewise, we will apply the threshold of a household with two adults for households that contain three or more adults.
-
-**Material deprivation composite.** We will compute a composite measure of material deprivation by aggregating across the perceived scarcity measures and the income-to-needs ratio.
-We will do so separately for each year.
-First, we will create a single measure of perceived scarcity by separately standardizing the three measures and averaging them.
 Second, we will calculate the average within-person income-to-needs ratio for each year by averaging across the monthly income-to-needs estimates.
 We will reverse code the yearly income-to-needs ratio so that higher scores indicate more deprivation.
-Third, we will average across perceived scarcity and the income-to-needs ratio for each year, standardizing each before averaging, and then finally calculate the average across all year.
+
+**Material deprivation composite.** We plan to compute a composite measure of material deprivation by aggregating across the perceived scarcity measures and the income-to-needs ratio, unless the correlation between these two variables is less than .30.
+First, we will separately calculate the average income-to-needs ratio and perceived scarcity across years.
+If the correlation between these measures is < .30, we will use perceived scarcity alone in the main model, and will explore a model using the income-to-needs ratio in the supplement.
+If the correlation is > .30, we will standardize both measures separately and average them together.
 
@@
 
-**Unpredictability composite.** We will compute a composite measure of unpredictability based on the yearly unweighted averages of perceived scarcity and the income-to-needs ratio (reverse coded), as described above.
+**Unpredictability composite.** We will compute a composite measure of unpredictability based on the same measures as material deprivation, as described above.
 We will compute unpredictability as the within-person standard deviation in material deprivation across years [@ugarte_2023; @walasek_2023; @young_2020].
@@
 For that reason, the standard deviation alone has been called into question as a measure of adversity, as the same fluctuation in income can have a greater relative impact for people close to the poverty line than for people with high incomes.
-We will address this issue by including the covariance between material deprivation (the mean) and unpredictability (the standard deviation) into the SEM model (see below).
+We will address this issue by including the covariance between material deprivation (the mean) and unpredictability (the standard deviation) into the model (see below).
 
@@
 The main goals of this pilot study were to collect participant feedback (e.g., difficulty of instructions, whether we included sufficient breaks) and to analyze performance and correlations between tasks.
-The results of this pilot study are described in more detail in the Supplemental Materials.
+The results of this pilot study are described in more detail in the Supplemental Materials [https://stefanvermeent.github.io/liss_wm_profiles_2023/supplement/README.html](https://stefanvermeent.github.io/liss_wm_profiles_2023/supplement/README.html).
 
@@
 
-**Binding-Updating Task.** The Binding-Updating task (Figure 2C) is adopted from [XXX].
+**Binding-Updating Task.** The Binding-Updating task (Figure 2C) is adopted from @wilhelm_2013.
 On each trial, participants see a 3 x 3 grid, with a fixation cross in the central cell.
@@
 We will compute a binding score by calculating the overall recall accuracy (%) across trials with zero updating steps.
-We will compute an updating score by calculating the overall recall accuracy (%) of updated letters (disregarding non-updated letters) across trials containing updating steps.
+We will compute an updating score by calculating the overall recall accuracy (%) across trials with one or more updating steps.
 
@@
 The purpose was to ascertain the number of individuals who had finished the previous waves a few years ago and were presently still participating in the panel (i.e., to see if we could reasonably create a link between these previous waves and newly collected data).
+
 All data access events were automatically detected and logged on the GitHub repository using the *projectlog* R package [@vermeent_2023].

### 2024-01-31 manuscript/manuscript.qmd
Resubmission of Stage 1 manuscript - inclusion of Study Design Plan

@@
 library(lavaan)
+library(officer)
 
@@
 
-We will evaluate evidence for deficit- and adaptation-based frameworks (see Figure 1A for a visual summary, and supplemental Table S2 for the study design plan). 
+We will evaluate evidence for deficit- and adaptation-based frameworks (see Figure 1A for a visual summary, and Appendix 1 for the study design plan). 
 Crucially, as deficit and adaptation processes can operate in concert [@frankenhuis_2020], we could find support (or lack thereof) for both frameworks in the same model.
@@
 \pagebreak
+
+# Appendix I: Study Design Plan
+
+```{r}
+appendix1 <- tibble(
+  `Research question` = c(
+    "1. what is the association between adversity and WM capacity?", 
+    "2. what is the association between adversity and WM updating after accounting for WM capacity?", 
+    "3. Are the directions and strengths of these associations similar or different for neighborhood threat, material deprivation, and unpredictability?"),
+  `Hypotheses`        = c(
+    "Deficit frameworks predict a negative association between all three types of adversity and WM capacity as well as WM updating.
+This follows from the hypothesis that adversity leads to broad WM deficits.
+Deficit frameworks are partially supported if we find negative associations with only one (or two) types of adversity.
+",
+
+    "Within adaptation-based frameworks, theories make two predictions.
+
+First, if adaptive processes enhance WM updating and there are no impairment processes operating, we can expect a positive association between adversity and WM updating. 
+Second, if, adaptive processes operate in concert with general impairment processes, we can expect intact WM updating in combination with lowered WM capacity.
+If neither impairment nor adaptative processes are operating, we can expect both WM updating and capacity to be intact.",
+
+    "We have two expectations based on prior studies.
+First, we expect the association between material deprivation and WM capacity to be more negative than the associations with unpredictability and neighborhood threat.
+Second, we expect WM updating to be associated with unpredictability and neighborhood threat, but not with material deprivation."
+    ),
+  `Sampling plan`     = c(
+    "We are collecting data of 800 participants in the Dutch Longitudinal Internet studies for the Social Sciences (LISS) panel.
+First, we will use data that were previously collected in LISS.
+Second, we will use new data that we collected ourselves in LISS.
+Data collection started on October 2nd and is expected to be completed in February 2024.
+We signed a contract with LISS stipulating that we will receive access to the data only after Stage 1 acceptance of this Registered Report.
+
+To ensure sufficient representation of people from lower socioeconomic backgrounds, roughly half the total sample will be sampled from participants who reported one or more of the following at least once in the three years: (1) a monthly income < €1,500, (2) HAVO or VWO as highest completed education (which are the two highest levels in Dutch secondary education), or (3) a score of 4 or lower on the 'ladder of life'",
+    NA, 
+    NA),
+  `Analysis plan`     =  c(
+    "We will fit a single structural equation model (SEM) containing all adversity measures.
+We will use robust maximum likelihood estimation in case any variable is non-normally distributed.
+Missing data will be handled using full information maximum likelihood (FIML).
+If participants are from the same household, this clustering within families will be accounted for.
+
+WM capacity will be estimated as a latent factor loading on all outcome measures.
+In addition, we will estimate WM updating as a latent factor capturing residual variance in the updating measure.
+We will estimate the effect of each adversity type through regression analyses.
+Each association will be controlled for: (1) age in years; (2) the quadratic effect of age; (2) environmental noise; (3) two items measuring interruptions.
+
+We will estimate the model in two.
+First, we will construct the measurement model of WM, without including the adversity measures.
+Once we obtain at least acceptable model fit, we will access and add the adversity measures to the model.
+We will control for multiple testing using the false discovery rate
+We will use two one-sided tests (TOST) equivalence testing to test whether small effects—which we define as standardized effects between -.10 and .10—are practically equivalent, which we will interpret as evidence for intact performance.
+", 
+    NA, 
+    NA),
+  `Rationale for deciding the sensitivity of the test for confirming or disconfirming the hypothesis` = c(
+    "We based our power analysis on simulations reported by Kretzschmar ad Gignac (2019), determining the required sample size to detect a small effect size (β = 0.1) with at least 90% power at α = 0.05.
+Assuming a reliability of at least 0.7 (which is typical for WM tasks with a number of trials similar to ours; e.g., Wilhelm, et al., 2013), we would require a sample size of *N* = 730.
+Anticipating exclusions, we decided to include 800 participants.",
+    NA, 
+    NA),
+  `Interpretation given different outcomes` = c(
+    "Contrary to predictions of deficit perspectives, we might find that all associations between adversity and WM capacity are either practically equivalent or positive.
+This would suggest that WM capacity is either unaffected or even enhanced by adversity.
+
+If we find both a non-significant association and practical non-equivalence, we will conclude that our data neither support nor refute either framework.
+",
+
+    "Contrary to predictions of adaptation-based perspectives, we might find that the association between adversity and WM updating is negative.
+This would suggest that WM updating is impaired by adversity.
+
+Contrary to predictions of adaptation-based perspectives, we might find a practically equivalent association with adversity for both WM capacity and updating.
+This would suggest that WM is unaffected by adversity.
+
+If we find both a non-significant association and practical non-equivalence, we will conclude that our data neither support nor refute either framework.",
+
+    "We might find that the association between threat or unpredictability with WM capacity is more strongly or equally strongly negative than with material deprivation.
+This would suggest that threat or unpredictability are more strongly associated with WM capacity than material deprivation.
+
+We might also find that material deprivation, but not unpredictability or neighborhood threat, is positively associated with WM updating.
+This would suggest that an enhanced updating ability has an adaptive benefit for individuals experiencing material deprivation."
+    ),
+  `Theory that could be shown wrong by the outcomes` = c(
+    "Theoretically, our analyses directly compare evidence in favor of deficit and adaptation-based perspectives.
+Both are established frameworks generating predictions that extend to other cognitive abilities beyond WM.
+Therefore, the current study could neither confirm nor disconfirm the frameworks in general.
+
+However, our findings could be (partially) inconsistent with predictions derived from both frameworks.
+Deviating findings for RQ1 or RQ 2 would require revising theoretical predictions about the specific WM abilities that are adapted to/impaired by adversity.
+
+In both cases, it would suggest that both frameworks need to be explicit in how they distinguish between different WM components.",
+
+    NA, 
+    "The hypotheses specified for RQ3 do not directly offer (non-) support for either framework.
+However, finding different patterns than hypothesized here would be inconsistent with findings of prior studies."),
+) |> 
+  flextable() |> 
+  merge_at(i = 1:3, j = 3, part = "body") |> 
+  merge_at(i = 1:3, j = 4, part = "body") |>
+  merge_at(i = 1:3, j = 5, part = "body") |>
+  merge_at(i = 1:2, j = 7, part = "body") |>  
+  add_header_row(
+    values = " ",
+    colwidths = 7
+  ) |>
+  flextable::compose(
+    i = 1, j = 1,
+    as_paragraph(as_b("Table S2. "), "Study design plan."),
+    part = "header"
+  ) |> 
+  border(border.top = fp_border(color = "black"), part = "body") |> 
+  border(border.left = fp_border(color = "black"), part = "body") |> 
+  border(border.right = fp_border(color = "black"), part = "body") |> 
+  border(border.bottom = fp_border(color = "black"), part = "body") |> 
+  border(i = 2, border.left = fp_border(color = "black"), part = "header") |>
+  border(i = 2, j = 7, border.right = fp_border(color = "black"), part = "header") |>
+  border(border.bottom = fp_border(color = "black"), part = "body") |> 
+  border(i = 1, j = 1, border.left =  fp_border_default(style = "none", width = 0), part = "header") |>
+  border(i = 1, j = 7, border.right =  fp_border_default(style = "none", width = 0), part = "header") |> 
+  border(i = 1, border.top =  fp_border_default(style = "none", width = 0), part = "header") |> 
+  valign(valign = "top") |> 
+  width(width = 1)
+
+appendix1
+```

### 2024-03-12 manuscript/manuscript.qmd
Revision 1 for PCI-RR

@@
 
-Decades of research have shown that adversity tends to lower working memory (WM) performance.
-This literature has mainly focused on impairments in the overall capacity to hold information available in WM for further processing.
+Decades of research have shown that adversity tends to be associated with lower working memory (WM) performance.
+This literature has mainly focused on impairments in the capacity to hold information available in WM for further processing.
 However, some recent adaptation-based studies suggest that certain types of adversity can leave intact, or even enhance, the ability to rapidly update information in WM.
-One key issue in evaluating deficits and adaptations in WM is that WM capacity and updating tasks tend to covary.
+One key challenge is that WM capacity and updating tasks tend to covary.
 This is likely due to the fact that both types of tasks require the creation and maintenance of bindings in WM; links between mental representations of information in WM.
@@
 Participants completed three WM tasks: two complex span tasks and a task measuring both binding and updating of information.
-In addition, we will derive estimates of participants' exposure to neighborhood threat, material deprivation, and unpredictability.
+In addition, we will estimate participants' exposure to neighborhood threat, material deprivation, and unpredictability.
 Using structural equation modeling, we will estimate associations between the three types of adversity and latent estimates of WM capacity and updating.
-These findings will advance our theoretical understanding of how adversity is associated with WM, which will aid the development of interventions aimed at alleviating performance difficulties and leveraging areas of strengths.
+These findings will advance our theoretical understanding of how adversity is associated with WM, which will aid interventions aimed at alleviating performance difficulties and leveraging areas of strength.
 
@@
 Living in adverse conditions, with prolonged exposure to intense stress, tends to have a profound and enduring impact on cognitive functioning [@farah_2006; @sheridan_2022; @sheridan_2014].
+Although adversity can be described in many ways, we follow contemporary models focusing on threat, deprivation, and unpredictability as key dimensions of adversity [@ellis_2009; @ellis_2022; @mclaughlin_2016; @mclaughlin_2021].
 A domain that seems to be particularly affected by adversity is working memory (WM). 
@@
 Studies with college students have found a link between both recent and lifetime experiences of stressful major life events (discrete negative events that have a clear onset and offset, unlike chronic adversity) with lowered WM capacity [@klein_2001; @shields_2017; @shields_2019].
-The most common tasks used to examine this link are simple span tasks (repeating a string of stimuli of increasing length), complex span tasks (remembering a string of stimuli while being engaged by a secondary task), and *n*-back tasks (judging whether the current stimulus in a string is identical to the stimulus *n* steps ago)  [@goodman_2019].
+
+The most common tasks used to examine the negative association between adversity and WM are simple span tasks (repeating a string of stimuli of increasing length), complex span tasks (remembering a string of stimuli while being engaged by a secondary task), and *n*-back tasks (judging whether the current stimulus in a string is identical to the stimulus *n* steps ago)  [@goodman_2019].
 Performance on these tasks is assessed through the number of items that participants can retain in WM, that is, their overall capacity [with the exception of *n*-back\; for concerns about the construct validity of this task, see @frost_2021; @kane_2007].
 
-Although both early-life and recent adversity appear to be negatively associated with WM capacity, a small set of studies suggest that exposure to adversity may leave intact, or even enhance, the ability to update items in WM [@young_2018; @young_2022].
+Although both early-life and recent adversity appear to be negatively associated with WM capacity, a small set of studies suggest that exposure to adversity may leave intact, or even enhance, the ability to update items in WM in adolescents [@young_2022] and adults [@young_2018].
 Updating is defined as the ability to rapidly replace old information in WM with new information.
 The finding that updating may be left intact or even enhanced after exposure to adversity exemplifies emerging theoretical frameworks grounded in adaptive reasoning that are complementary to deficit frameworks [@frankenhuis_2013; @ellis_2017; @ellis_2022; @frankenhuis_2020].
+
 Adaptation-based theories assume that developmental processes tailor an individual's cognitive abilities to the unique challenges and opportunities posed by their environment.
@@
 deprivation) pose different challenges, they should (at least in part) shape cognitive abilities in different ways.
-WM updating may be especially adaptive in *unpredictable* environments, characterized by random variation in adversity exposure over space or time.
+For example, with regards to executive functioning, some previous studies have found that children and adults with more exposure to unpredictability (characterized by random variation in adversity exposure over space or time) and threat tend to be better at rapidly shifting their attention between tasks [@fields_2021; @mittal_2015; @steudte_2014; @young_2022; but see @nweze_2021].
+WM updating may be especially adaptive in unpredictable environments.
 WM updating allows people to maintain an up-to-date overview of the (changing) current state of the environment [@young_2018].
@@
 
-In this study, we will estimate associations between three types of adversity exposure and latent estimates of WM capacity and updating.
-Specifically, we will measure exposure to two forms of harshness: *neighborhood threat* (perceived neighborhood safety, personal crime victimization), and *material deprivation* (perceived scarcity, income-to-needs).
-We also include a measure of *unpredictability*, indexed as variability in material deprivation over time.
-Together, these adversity types capture key dimensions in contemporary models of adversity [threat vs. deprivation vs. unpredictability\; @ellis_2009; @ellis_2022; @mclaughlin_2016; @mclaughlin_2021].
+In this study, we will estimate associations between latent estimates of WM capacity and updating with three types of adversity: threat, deprivation, and unpredictability. 
+Together, these adversity types capture key dimensions in contemporary models of adversity [@ellis_2009; @ellis_2022; @mclaughlin_2016; @mclaughlin_2021]. 
+Threat refers to experiences involving the potential for harm imposed by others. 
+We focus on perceived neighborhood violence, the extent to which an individual reports having been exposed to acts of violence in their neighborhood. 
+Deprivation refers to having a low level of resources. 
+We focus on perceived material deprivation, a (perceived) lack of access to material resources. 
+Unpredictability refers to variation in material deprivation over time. 
+This definition is inspired by, but deviates from the harshness-unpredictability framework, in which unpredictability is defined as stochastic variation in harshness (age-specific rates in morbidity and mortality) over space and time [@ellis_2009; @ellis_2022].
 We will not calculate unpredictability in neighborhood threat given that participants have at most six timepoints, and often as few as one or two, which is insufficient to calculate variation over time [@walasek_2023].
+
 We will address three research questions.
@@
 Crucially, as deficit and adaptation processes can operate in concert [@frankenhuis_2020], we could find support (or lack thereof) for both frameworks in the same model. 
+We distinguish between three between-person data patterns: (1) lowered performance, (2) enhanced performance, and (3) intact performance.
+We define lowered performance as a statistically significant negative association between a type of adversity and WM capacity or updating (irrespective of effect size).
+We define enhanced performance as a statistically significant positive association between a type of adversity and WM capacity or updating (irrespective of effect size).
+We define intact performance as an association between a type of adversity and WM capacity or updating that has a standardized effect smaller than 0.1 *and* larger than -0.1---even if the effect is statistically different from zero---which we will test using Two One-Sided T-Tests (TOST) equivalence testing [see the 'Primary analyses' section\; @lakens_2018].
 
@@
 #|   **Figure 1.** Overview of predictions derived from deficit and adaptation frameworks.
-Panel A depicts the most likely data patterns based on previous literature, and whether we would consider them consistent with deficit and adaptation frameworks (see the main text for more details).
+Panel A depicts the most likely between-person data patterns based on previous literature, and whether we would consider them consistent with deficit and adaptation frameworks (see the main text for more details).
 Panel B depicts an overview of the Structural Equation Model.
@@
 WM updating is modeled as a latent factor capturing the residual variance in the updating task after accounting for variance related to WM capacity.
+INR = income-to-needs ratio; Perc.
+Scarcity = perceived scarcity; SD = standard deviation.
 knitr::include_graphics("fig/fig1.png")
@@
 
-We collected data in the Longitudinal Internet studies for the Social Sciences (LISS) panel [@scherpenzeel_2011].
+Our study will include 800 participant who were randomly sampled from the Longitudinal Internet studies for the Social Sciences (LISS) panel [@scherpenzeel_2011].
 The LISS panel is a representative probability sample of roughly 5,000 Dutch households (~7,500 individuals) drawn from the population register by Statistics Netherlands on an invite-only basis.
@@
 The current study integrates two data sources.
-First, we will use data that were previously collected in LISS (hereafter referred to as the LISS archive).
-Second, we will use new data that we collected ourselves in LISS between October 2nd and December 22nd, 2023 (hereafter referred to as newly collected data).
-We signed a contract with LISS stipulating that we will receive access to the data only after Stage 1 acceptance of this Registered Report.
+First, our sample of 800 participants participated in a new LISS study between October 2023 and February 2024 (hereafter referred to as 'newly collected data'), in which we included a measure of neighborhood threat and multiple measures of working memory.
+Second, we will access data that were previously collected in LISS (hereafter referred to as 'the LISS archive').
+See Figure 2 for a visual overview of the data sources and their measurement timepoints.
+We signed a contract with LISS stipulating that we will receive access to the newly collected data only after Stage 1 acceptance of this Registered Report.
+
+```{r}
+#| label: Figure2
+#| fig-width: 6.5
+#| fig-height: 7
+#| dpi: 600
+#| out-width: 6in
+#| fig-cap: | 
+#|   **Figure 2.**.
+Overview of the different data sources used in this study.
+We distinguish between measures taken from the LISS data archive and measures that were newly collected in our own study between October 2023 and February 2024.
+Perceived scarcity and income were collected yearly in the full panel from 2008 – 2023.
+Neighborhood crime and crime victimization were collected across six waves between 2008 and 2018.
+In the newly collected data, we collected data on a measure of neighborhood threat and multiple measures of working memory.
+Note that participants may not have data across all timepoints of the archived studies because they joined the LISS panel more recently or because they did not participate in each wave.
+knitr::include_graphics("fig/fig2.png")
+```
 
@@
 
-**Neighborhood threat composite.** We will compute a composite measure of neighborhood threat by aggregating across the perceived neighborhood crime and crime victimization measures exposure.
-First, we will create a single measure of perceived neighborhood crime by separately standardizing the two scales and averaging them.
-Second, we will average across perceived neighborhood crime and crime victimization, standardizing each before averaging.
+**Neighborhood threat composite.** We will first compute an average across time for each measure separately (i.e., the two measures of neighborhood crime and the measure of crime victimization).
+If all correlations between averaged measures are equal to or larger than .60 (i.e., indicating a "strong" correlation), then we will compute a uniformly weighted average.
+If the correlation is lower than .60, we will apply Principal Component Analysis (PCA) to the averaged measures and extract only the first principal component score.
 
@@
 
-**Perceived scarcity.** We will use a few items from the LISS archive that were collected on a yearly basis between 2007 and 2023 ([https://doi.org/10.57990/1gr4-bf42](https://doi.org/10.57990/1gr4-bf42)) to index perceived scarcity.
+We will measure material deprivation with two separate indicators: perceived scarcity and the income-to-needs ratio.
+
+**Perceived scarcity (mean).** We will use a few items from the LISS archive that were collected on a yearly basis between 2008 and 2023 ([https://doi.org/10.57990/1gr4-bf42](https://doi.org/10.57990/1gr4-bf42)) to index perceived scarcity.
 First, participants indicated how hard or easy it currently is to live off the income of their household, on a scale from 0 (very hard) to 10 (very easy).
@@
 Third, participants answered which of the following issues they were confronted with at present (0 = no, 1 = yes): (1) "having trouble making ends meet"; (2) unable to quickly replace things that break"; (3) "having to lend money for necessary expenditures"; (4) "running behind in paying rent/mortgage or general utilities"; (5) "debt collector/bailiff at the door in the last month"; (6) "received financial support from family or friends in the last month".
-We will sum across these six items, with higher scores indicating more financial difficulties.
-Finally, we will create a single measure of perceived scarcity for each year by separately standardizing the three measures and averaging them.
 
-**Income-to-needs.** We will calculate an income-to-needs ratio for each year using monthly self-reported net household income from the LISS archive ([https://doi.org/10.57990/qn3k-as78](https://doi.org/10.57990/qn3k-as78)).
+We will first compute the average across time for each item separately, and examine correlation between the item averages.
+We will follow the same guidelines as with neighborhood threat to determine if we will compute a uniformly weighted average or extract the first principal component using PCA.
+
+**Income-to-needs (mean).** We will calculate an income-to-needs ratio for each year using monthly self-reported net household income from the LISS archive ([https://doi.org/10.57990/qn3k-as78](https://doi.org/10.57990/qn3k-as78)).
 Zero values in household income will be set to missing, as these could either indicate the lack of an income or an unwillingness to disclose the income.
@@
 First, we will divide the average income per year by the *poverty threshold*, as determined by Statistics Netherlands (CBS) [@brakel_2023\; CBS, personal communication, December 15, 2023].
-The poverty threshold changes across time to account for purchasing power.
-In addition, different thresholds are determined each year for different household compositions (single vs. two adults; zero, one, two, or three children).
 As thresholds are only provided for households with up to three children, we will apply the threshold of a household with three children to households with more than three children.
@@
 
-**Material deprivation composite.** We plan to compute a composite measure of material deprivation by aggregating across the perceived scarcity measures and the income-to-needs ratio, unless the correlation between these two variables is less than .30.
-First, we will separately calculate the average income-to-needs ratio and perceived scarcity across years.
-If the correlation between these measures is < .30, we will use perceived scarcity alone in the main model, and will explore a model using the income-to-needs ratio in the supplement.
-If the correlation is > .30, we will standardize both measures separately and average them together.
-
 ### Unpredictability
 
-**Unpredictability composite.** We will compute a composite measure of unpredictability based on the same measures as material deprivation, as described above.
-We will compute unpredictability as the within-person standard deviation in material deprivation across years [@ugarte_2023; @walasek_2023; @young_2020].
+**Perceived scarcity (SD/mean).** This measure is based on the same items as outlined above (see Perceived scarcity (mean)).
+We will compute unpredictability over time in perceived scarcity using the coefficient of variation, which is the within-person standard deviation across years divided by the mean [@key_2017; @liu_2022; @ugarte_2023; @walasek_2024; @young_2020].
 The mean and standard deviation in income have been found to be strongly negatively correlated, indicating that people with lower incomes tend to experience less variability in income [@li_2018; @young_2023].
 For that reason, the standard deviation alone has been called into question as a measure of adversity, as the same fluctuation in income can have a greater relative impact for people close to the poverty line than for people with high incomes.
-We will address this issue by including the covariance between material deprivation (the mean) and unpredictability (the standard deviation) into the model (see below).
+
+To compute an overall measure of unpredictability in perceived scarcity, we will follow the same procedure as with mean perceived scarcity, but based on the coefficient of variation.
+
+**Income-to-needs (SD/mean).** Similar to perceived scarcity, we will compute unpredictability over time in the income-to-needs ratio using the coefficient of variation.
 
@@
 On each trial, participants have to memorize between four and six letters, with each set size repeated three times.
-At the end of each sequence, all letters are presented in a 3 x 4 grid, and participants click the letters in the correct order.
+At the end of each sequence, all letters are presented in a 3×4 grid, and participants click the letters in the correct order.
 
@@
 If they perform at or below chance, they have the opportunity to either repeat a part or advance to the next part.
-After practicing, participants complete 12 test trials.
+After practicing, participants complete 9 test trials, with a total of 45 recall items and 45 math items.
 We will compute an operation span score by calculating the proportion of letters recalled in the correct sequential position across trials [@conway_2005].
@@
 If they perform at or below chance, they have the opportunity to either repeat a part or to advance to the next part.
-After practicing, participants complete 12 test trials.
-We will compute a rotation span score by calculating the proportion of arrows recalled in the correct sequential position across trials [@conway_2005.
+After practicing, participants complete 12 test trials, with a total of 45 recall items and 45 letter items.
+We will compute a rotation span score by calculating the proportion of arrows recalled in the correct sequential position across trials [@conway_2005].
 
 **Binding-Updating Task.** The Binding-Updating task (Figure 2C) is adopted from @wilhelm_2013.
-On each trial, participants see a 3 x 3 grid, with a fixation cross in the central cell.
+On each trial, participants see a 3×3 grid, with a fixation cross in the central cell.
 After 1,000 ms, they are presented with a sequence of numbers (0-9) in random locations of the grid.
@@
 If they perform at or below chance, they have the opportunity to either repeat the practice trials or to advance to the actual task.
-After practicing, they complete 18 test trials.
+After practicing, they complete 18 test trials, of which nine are binding-only (24 recall items in total) and nine are updating trials (24 recall items in total).
 We will compute a binding score by calculating the overall recall accuracy (%) across trials with zero updating steps.
@@
 ```{r}
-#| label: Figure2
+#| label: Figure3
 #| fig-width: 6.5
@@
 #| fig-cap: | 
-#|   **Figure 2.** Overview of the working memory tasks.
+#|   **Figure 3.** Overview of the working memory tasks.
 Panel A: Operation Span Task.
-Participants had to memorize letters in the correct order, while engaging in a secondary math task.
+Participants memorized letters in the correct order, while engaging in a secondary math task.
 Panel B: Rotation Span Task.
-Participants had to memorize the orientation of arrows, while judging whether letters were mirrored or normal in a secondary task.
-Panel C: Participants had to memorize numbers in the correct location in a 3x3 grid.
+Participants memorized the orientation of arrows, while judging whether letters were mirrored or normal in a secondary task.
+Panel C: Participants memorized numbers in the correct location in a 3×3 grid.
 On half of the trials, all numbers were presented in unique locations, only requiring binding the numbers to the correct position.
 On the other half, some numbers were presented in the same location as a previously presented number, requiring updating.
-Note that stimuli are not to scale.
-knitr::include_graphics("fig/fig2.png")
+Note: stimuli are not to scale.
+knitr::include_graphics("fig/fig3.png")
 ```
@@
 During planning of the study, the first author accessed to the LISS data archive and inspected three waves of the LISS data containing the items about neighborhood safety and crime exposure, as well as the three most recent monthly data collections containing basic demographic info.
-The purpose was to ascertain the number of individuals who had finished the previous waves a few years ago and were presently still participating in the panel (i.e., to see if we could reasonably create a link between these previous waves and newly collected data).
+The purpose was to ascertain the number of individuals who had finished the previous waves in the LISS data archive and were presently still participating in the panel (i.e., to see if we could reasonably create a link between the LISS data archive and newly collected data).
 
 All data access events were automatically detected and logged on the GitHub repository using the *projectlog* R package [@vermeent_2023].
-We took the following measures to prevent bias: 1) we randomly shuffled the participant IDs in each data set using the *projectlog* R package, so that we were unable to link participant data between (waves of) studies; 2) we did not inspect any of the measures that will be part of our adversity composites; 3) we did not know which participants would be selected for the new LISS study; 4) the primary analyses will be based on composite measures that combine measures from these previous studies with data that were collected at a later timepoint.
+We took the following measures to prevent bias: 1) we randomly shuffled the participant IDs in each data set using the *projectlog* R package, so that we were unable to link participant data between (waves of) studies in the LISS data archive; 2) we did not inspect any of the measures that will be part of our adversity composites; 3) we did not know which participants would be selected for the newly collected data; 4) the primary analyses will be based on composite measures that combine measures from the LISS data archive with measures from the newly collected data.
 
@@
 We will control for multiple testing using the false discovery rate [@benjamini_1995; @cribbie_2007].
-We will use two one-sided tests (TOST) equivalence testing to test whether small effects---which we define as standardized effects between -.10 and .10---are practically equivalent, which we will interpret as evidence for intact performance.
+
+To statistically test whether small effects are practically equivalent to zero---suggesting intact performance---we will use Two One-Sided T-tests (TOST) equivalence testing [@lakens_2018], using -0.1 and 0.1 as equivalence bounds.
+TOST equivalence testing allows us to conclude intact performance based on a significant effect, rather than erroneously interpreting a non-significant effect as evidence for the absence of an effect.
+We consider any effect that falls within this region to reflect practical equivalence, that is, a between-person difference in performance that is practically equivalent to zero. 
+TOST provides two *p*-values, one testing against the upper bound and one testing against the lower bound; we will report only the largest of the two *p*-values.
 

### 2024-03-18 manuscript/manuscript.qmd
fix reference error in revised manuscript

@@
 This definition is inspired by, but deviates from the harshness-unpredictability framework, in which unpredictability is defined as stochastic variation in harshness (age-specific rates in morbidity and mortality) over space and time [@ellis_2009; @ellis_2022].
-We will not calculate unpredictability in neighborhood threat given that participants have at most six timepoints, and often as few as one or two, which is insufficient to calculate variation over time [@walasek_2023].
+We will not calculate unpredictability in neighborhood threat given that participants have at most six timepoints, and often as few as one or two, which is insufficient to calculate variation over time [@walasek_2024].
 

### 2024-06-25 scripts/0_custom_functions/corr_plot.R
updates to analysis scripts

@@ -1,47 +1,48 @@
-corr_table <- function (data, sample_size = T, use = "pairwise", method = "pearson", 
-                        stats = NULL, c.names = names(data), change = F, numbered = F, 
+corr_table <- function (data, sample_size = T, use = "pairwise", method = "pearson",
+                        stats = NULL, c.names = names(data), change = F, numbered = F,
                         flagged = T) {
-  
+
   my_ifelse <- function(...) {
     suppressWarnings(ifelse(...))
   }
-  
-  descriptives <- psych::describe(data) %>% as.data.frame() %>% round(2) 
+
+  descriptives <- psych::describe(data) %>% as.data.frame() %>% round(2)
   corr_data <- psych::corr.test(x = data, use = use, method = method)
   ns <- corr_data$n
   rs <- corr_data$r %>% round(2) %>% formatC(digits = 2, width = 3, flag = '0', format = 'f')
-  
-  descriptives <- 
-    descriptives %>% 
-    select(-vars) %>% 
-    rownames_to_column(var = "var") %>% 
-    mutate(across(3:13, ~ formatC(.x, digits = 2, width = 3, flag = '0', format = 'f'))) %>% 
-    pivot_longer(cols = c(-var), names_to = "stat", values_to = "value", values_transform = list(value = as.character)) %>% 
+
+  descriptives <-
+    descriptives %>%
+    select(-vars) %>%
+    rownames_to_column(var = "var") %>%
+    mutate(across(3:13, ~ formatC(.x, digits = 2, width = 3, flag = '0', format = 'f'))) %>%
+    pivot_longer(cols = c(-var), names_to = "stat", values_to = "value", values_transform = list(value = as.character)) %>%
     pivot_wider(names_from = var, values_from = value)
-  
+
   if (!is.null(stats)) {
-    descriptives <- 
-      descriptives %>% 
-      filter(stat %in% stats) %>% 
-      slice(match(stats, stat)) %>% 
+    descriptives <-
+      descriptives %>%
+      filter(stat %in% stats) %>%
+      slice(match(stats, stat)) %>%
       mutate(stat = str_to_title(stat), stat = ifelse(stat=="Sd","SD",stat))
   }
   if (flagged) {
[…]

### 2024-06-25 scripts/0_custom_functions/impute_income.R

@@ -1,0 +1,35 @@
+
+impute_income <- function(month_data, year_data) {
+
+  imputed_data <- unique(month_data$nomem_encr) |>
+    map(function(x){
+
+      month_data_i <- month_data |>
+        filter(nomem_encr == x) |>
+        group_by(year) |>
+        mutate(all_missing = ifelse(all(is.na(nettohh_f)), TRUE, FALSE))
+
+      if(any(month_data_i$all_missing)) {
+        missing_years <- month_data_i |>
+          filter(all_missing) |>
+          pull(year) |>
+          unique()
+
+        year_data_i <- year_data |>
+          filter(nomem_encr == x, year %in% missing_years) |>
+          select(nomem_encr, year, net_inc_y)
+
+        if(nrow(year_data_i) == 0) {return(month_data_i)}
+
+        message("data imputed for ", x)
+        return(month_data_i |> left_join(year_data_i))
+
+      } else
+      {
+        return(month_data_i)
+      }
+
+    })
+
+  return(imputed_data |> bind_rows())
+}

### 2024-06-25 scripts/0_custom_functions/liss-aggregate-data.R

@@ -1,0 +1,37 @@
+library(tidyverse)
+library(haven)
+
+liss_income_zips <- list.files("income", full.names = TRUE)
+
+liss_income <-
+  map_df(liss_income_zips, function(x){
+
+    liss_wave <- str_extract(x, "ci\\d\\d")
+    liss_sav <- str_remove(x, "income.") |> str_replace("zip$", "sav")
+
+    print(liss_wave)
+    print(liss_sav)
+
+    unzip(x)
+    liss_data <-
+      read_sav(liss_sav) |>
+      rename_with(~str_remove(.x, liss_wave)) |>
+      rename_with(.cols = matches("_m$"), .fn =  ~paste("period")) |>
+      mutate(
+        wave = liss_wave
+      ) |>
+      select(
+        wave, period, nomem_encr,
+        matches("00([2-5]|[8:9])$|23[0-6]$|a24[3-9]$|25[0-3]$")
+      ) |>
+      rename_with(.cols = c(-wave, -period, -nomem_encr), .fn =  ~str_replace(.x, "^.", "income"))
+
+    unlink(liss_sav)
+
+    liss_data
+  })
+
+
+liss_income |>
+  arrange(nomem_encr, wave) |>
+  print(n = 100)

### 2024-06-25 scripts/0_custom_functions/liss-aggretaion.R

@@ -1,0 +1,52 @@
+# some trials?
+unzip_read_sav("liss/background/avars_200711_EN_3.0p.zip") |>
+  select(nomem_encr, nohouse_encr, aantalhh, positie, gebjaar, leeftijd, lftdhhh, oplcat, doetmee)
+
+unzip_read_sav("liss/background/avars_202209_EN_1.0p.zip") |>
+  select(nomem_encr, nohouse_encr, aantalhh, positie, gebjaar, leeftijd, lftdhhh, oplcat, doetmee) |>
+  filter(nohouse_encr == 583404)
+
+liss_background <-
+  map(
+    list.files("liss/background", pattern = ".zip$", full.names = T),
+    unzip_read_sav
+  )
+
+liss_health <-
+  map(
+    list.files("liss/health", pattern = ".zip$", full.names = T),
+    function(x){
+      unzip_read_sav(x) |>
+        select(nomem_encr, starts_with("ch")) |>
+        rename_with(.cols = c(starts_with("ch")), ~str_replace(.x, "^ch\\d\\d[a-n]","q")) |>
+        rename_with(.cols = c(matches("^q_m$")), ~c("wave")) |>
+        select(-any_of(c("q256","q258")))
+    }
+  )
+
+
+liss_background_data <- list_rbind(liss_background)
+liss_health_data <- reduce(liss_health, bind_rows)
+
+liss_health_data |> var_lookup(collapse = ", ") |> View()
+
+liss_health[[10]]$q016
+
+d2 <- liss_health |>
+  map(function(x){
+    x |>
+      select(matches("q0(16|17|19)$"))
+  })
+d
+
+glue::glue(
+  "d2[[1]] |> sjlabelled::var_labels(",
+  glue::glue_data(
+    d[[1]],
+    "{Variable} = '{str_trim(Label)}',"
+  ) |> paste(collapse = " "),
+  ")"
+) |>
+  rlang::parse_expr() |>
+  rlang::eval_tidy() |>
+  var_lookup(collapse = ", ")

### 2024-06-25 scripts/0_custom_functions/liss-catalogs.R

@@ -1,0 +1,115 @@
+library(tidyverse)
+library(haven)
+
+#script number 1 of the liss scripts for unzupping the data and working with them
+
+source("scripts/0-functions.R")
+
+liss_health <-
+  unzip_read_sav("liss/health/ch07a_2p_EN.zip") |>
+  var_lookup(collapse = ", ")
+
+liss_neigh <-
+  unzip_read_sav("liss/neighborhood-perceptions/sr20a_EN_1.0p.zip") |>
+  var_lookup(collapse = ", ")
+
+liss_saftey <-
+  unzip_read_sav("liss/personal-saftey/oc17a_EN_1.0p.zip") |>
+  var_lookup(collapse = ", ")
+
+liss_safteymonitor <-
+  unzip_read_sav("liss/saftey-monitor/ic13a_EN_1.0p.zip") |>
+  var_lookup(collapse = ", ")
+
+liss_crime08 <-
+  unzip_read_sav("liss/conventional-computer-crime/ac08a_2.0p_EN.zip") |>
+  var_lookup(collapse = ", ")
+
+liss_crime10 <-
+  unzip_read_sav("liss/conventional-computer-crime/ac10b_1.0p_EN.zip") |>
+  var_lookup(collapse = ", ")
+
+liss_crime12 <-
+  unzip_read_sav("liss/conventional-computer-crime/ac12c_EN_1.0p.zip") |>
+  var_lookup(collapse = ", ")
+
+liss_crime14 <-
+  unzip_read_sav("liss/conventional-computer-crime/ac14d_EN_1.0p.zip") |>
+  var_lookup(collapse = ", ")
+
+liss_crime16 <-
+  unzip_read_sav("liss/conventional-computer-crime/ac16e_EN_1.0p.zip") |>
+  var_lookup(collapse = ", ")
+
+liss_crime18 <-
+  unzip_read_sav("liss/conventional-computer-crime/ac18f_EN_1.0p.zip") |>
+  var_lookup(collapse = ", ")
+
+liss_lifehistory12 <-
+  unzip_read_sav("liss/life-history/cb12a_EN_1.0p.zip") |>
+  var_lookup(collapse = ", ")
+
+liss_lifehistory13 <-
+  unzip_read_sav("liss/life-history/cb13b_EN_1.0p.zip") |>
+  var_lookup(collapse = ", ")
+
+liss_victims18a <-
[…]

### 2024-06-25 scripts/0_custom_functions/liss-check.R

@@ -1,0 +1,15 @@
+# test whether the unzipping has worked
+list.files("liss/background") |>
+  str_extract("\\d\\d\\d\\d\\d\\d") |>
+  as_tibble() |>
+  mutate(
+    year = str_extract(value, "^\\d\\d\\d\\d"),
+    month = str_extract(value, "\\d\\d$")
+  ) |>
+  group_by(year) |>
+  count(
+  )
+
+list.files("liss/background") |>
+  str_extract("\\d\\d\\d\\d\\d\\d") |>
+  str_subset("2015")

### 2024-06-25 scripts/0_custom_functions/liss-functions.R

@@ -1,0 +1,55 @@
+# Unzipping data from liss downloads --------------------------------------
+unzip_liss <- function(zip_path){
+
+  zip_exdir <-
+    ifelse(
+      str_detect(zip_path, "\\/"),
+      str_extract(zip_path, "^(.*\\/)+"),
+      "."
+    )
+
+  zip_sav <- str_replace(zip_path, ".zip$", ".sav")
+  zip_dta <- str_replace(zip_path, ".zip$", ".dta")
+
+  unzip(zip_path, overwrite = TRUE,  exdir = zip_exdir)
+
+  unlink(zip_dta)
+}
+
+# look up table ------------------------------------------------------------
+var_lookup <- function (data, ...) {
+  lp_data <-
+    tibble::tibble(
+      Variable = names(data),
+      Label = purrr::map_chr(data,
+                             function(x) {
+                               extracted_label <- attr(x, which = "label")
+                               if (is.null(extracted_label)) {
+                                 extracted_label <- ""
+                               }
+                               extracted_label
+                             }), Values = purrr::map_chr(data, function(x) {
+                               extracted_values <- attr(x, which = "labels")
+                               if (is.null(extracted_values)) {
+                                 extracted_values <- ""
+                               }
+                               else {
+                                 extracted_values <- paste(names(extracted_values),
+                                                           "=", extracted_values, ...)
+                               }
+                               extracted_values
+                             }))
+  lp_table <- lp_data
+  no_labels <- lp_table %>% dplyr::filter(Label == "")
+  no_values <- lp_table %>% dplyr::filter(Values == "")
+  if (nrow(no_labels) > 0) {
+    message("The following variables have no labels:\n",
+            paste(no_labels %>% dplyr::pull(Variable), collapse = "\n"))
+  }
+  if (nrow(no_values) > 0) {
+    message("The following variables have no value labels:\n",
+            paste(no_values %>% dplyr::pull(Variable), collapse = "\n"))
+  }
+  lp_table
+}
+

### 2024-06-25 scripts/1_sim_pipeline/0_ivs.R

@@ -1,342 +1,0 @@
-# 1. Load libraries ----------------------------------------------------------
-
-library(tidyverse)
-
-source('scripts/0_custom_functions/codebooks.R')
-
-# 2. Functions ---------------------------------------------------------------
-
-# INCOME-TO-NEEDS RATIO FOR DUTCH HOUSEHOLDS
-# the information up until 2021 was derived from van den Brakel et al. 2023 (See main text).
-# The information for 2022 was derived from https://www.cbs.nl/nl-nl/nieuws/2023/45/laagste-armoederisico-in-45-jaar-door-energietoeslag#:~:text=De%20lage%2Dinkomensgrens%20voor%20een,de%20grens%201%20830%20euro.
-# At the moment of writing, there was no final data for 2023, so we used the same values as 2022.
-inr_conv_table <-
-  tibble(
-    year            = rep(str_c(20, c("07", "08", "09", 10:23)), each = 8),
-    adult           = rep(c(1, 1, 1, 1, 2, 2, 2, 2), 17),
-    aantalki        = rep(c(0, 1, 2, 3, 0, 1, 2, 3), 17),
-    eq_fct_pre2018  = rep(c(1, 1.33, 1.51, 1.76, 1.37, 1.67, 1.88, 2.06), 17),
-    eq_fct_post2018 = rep(c(1, 1.32, 1.52, 1.73, 1.40, 1.69, 1.91, 2.09), 17)
-  ) |>
-  mutate(
-    # `base` is the low-income threshold for a single-person household without children.
-    # All other threshold are derived from the base by multiplying it with the equivalence factor.
-    base = case_when(
-      year %in% c("2007", "2008", "2009")                 ~ 870,
-      year %in% c("2010", "2011", "2012", "2013", "2014") ~ 940,
-      year %in% c("2015", "2016", "2017")                 ~ 1030,
-      year == "2018"                                      ~ 1060,
-      year == "2019"                                      ~ 1090,
-      year == "2020"                                      ~ 1100,
-      year == "2021"                                      ~ 1130,
-      year == "2022"                                      ~ 1200,
-      year == "2023"                                      ~ 1200
-    ),
-    threshold = ifelse(str_detect(year, "(18|19|20|21|22|23)$"), base*eq_fct_post2018, base*eq_fct_pre2018) |> plyr::round_any(x = _, 10, f = ceiling))
-
-
[…]

### 2024-06-25 scripts/1_sim_pipeline/1_dvs.R

@@ -1,38 +1,0 @@
-# 1. Libraries ----------------------------------------------------------------
-
-library(tidyverse)
-
-
-# 2. Stage 1: Simulate data for computational reproducibility ----------------
-
-dvs_sim <-
-  tibble(
-    nomem_encr = 1:800 |> as.character(),
-    rspan_cap     = rnorm(800, 0.75, 0.15) %>% ifelse(. > 1, 1, .),
-    rspan_sec_acc = rnorm(800, 0.88, 0.11) %>% ifelse(. > 1, 1, .),
-    rspan_sec_rt  = rnorm(800, 0.88, 0.11) %>% ifelse(. < 0, 300, .),
-    ospan_cap     = rnorm(800, 0.75, 0.15) %>% ifelse(. > 1, 1, .),
-    ospan_sec_acc = rnorm(800, 0.88, 0.11) %>% ifelse(. > 1, 1, .),
-    ospan_sec_rt  = rnorm(800, 1500, 476)  %>% ifelse(. < 0, 300, .),
-    bu_bind_acc   = rnorm(800, 0.88, 0.11) %>% ifelse(. > 1, 1, .),
-    bu_upd_acc    = rnorm(800, 0.70, 0.11) %>% ifelse(. > 1, 1, .),
-    brows_int     = sample(c(0, 1), size = 800, replace = T, prob = c(0.95, 0.05))
-  ) |>
-  sjlabelled::var_labels(
-    rspan_cap     = "Rspan Capacity. Average number of arrows recalled in the correct serial position on each trial (Proportion).",
-    rspan_sec_acc = "Rspan Accuracy on secondary rotation task. Average accuracy across trials (Proportion)",
-    rspan_sec_rt  = "Rspan RT on secondary rotation task. Average RT across trials (RT in miliseconds)",
-    ospan_cap     = "Ospan Capacity. Average number of letters recalled in the correct serial position on each trial. (Proportion)",
-    ospan_sec_acc = "Ospan Accuracy on secondary math task. Average accuracy across trials (Proportion)",
-    ospan_sec_rt  = "Ospan RT on secondary math task. Average RT across trials (RT in miliseconds)",
-    bu_bind_acc   = "Binding-Updating, Binding accuracy. Average amount of numbers recalled in the correct location across trials that did not involve an updating component (Proportion)",
[…]

### 2024-06-25 scripts/1_sim_pipeline/2_exclusions.R

@@ -1,28 +1,0 @@
-# 1. Libraries ---------------------------------------------------------------
-
-library(tidyverse)
-
-source('scripts/0_custom_functions/codebooks.R')
-source('scripts/0_custom_functions/binom_test.R')
-
-# 2. Load data ---------------------------------------------------------------
-
-load("data/dvs_sim.RData")
-load("data/ivs_sim.RData")
-
-
-# 3. Exclusions -----------------------------------------------------------
-
-full_data_sim <- full_join(ivs_sim, dvs_sim) |>
-  mutate(
-    ex_brows_int       = ifelse(brows_int == 0, FALSE, TRUE),
-    ex_acc_ospan_sec   = ifelse(ospan_sec_acc*100 < gbinom(45, 0.50), TRUE, FALSE),
-    ex_acc_rspan_sec   = ifelse(rspan_sec_acc*100 < gbinom(45, 0.50), TRUE, FALSE),
-  )
-
-clean_data_sim <- full_data_sim |>
-  filter(ex_brows_int == FALSE & ex_acc_ospan_sec == FALSE & ex_acc_rspan_sec == FALSE)
-
-
-save(full_data_sim, clean_data_sim, file = "data/full_data_sim.RData")
-

### 2024-06-25 scripts/1_sim_pipeline/4_sem.R

@@ -1,132 +1,0 @@
-
-# 1. Libraries and Data ------------------------------------------------------
-
-library(lavaan)
-library(tidyverse)
-library(lavaan.survey)
-
-# 2. Load data ------------------------------------------------------------
-
-load("data/full_data_sim.RData")
-
-# 2. Specify Clustering Design -----------------------------------------------
-
-# Some children are clustered within families.
-# We account for this using the lavaan.survey package
-
-cluster_design <- survey::svydesign(ids=~nohouse_encr, prob=~1, data = clean_data_sim)
-
-
-# 3. Model specification(s) --------------------------------------------------
-
-## 3.1 Measurement Model ----
-mod_meas <-
-  '
-  # Measurement Model
-
-  ## Latent factors
-  wmc_l      =~ 1*bu_bind_acc + rspan_cap + ospan_cap + bu_upd_acc
-  upd_l      =~ 1*bu_upd_acc
-
-  ## Covariances
-  bu_upd_acc ~~ 0*bu_upd_acc
-'
-
-
-fit_meas1 <- lavaan::sem(
-  model = mod_meas,
-  data = clean_data_sim,
-  estimator = "MLR", # Robust SEs in case of non-normal data
-  missing = "ML",    # FIML for missing data
-  orthogonal = TRUE,
-  auto.cov.lv.x = FALSE,
-  auto.cov.y = FALSE,
-  std.lv = FALSE,
-  fixed.x = FALSE
-)
-
-fit_meas1_cluster <- lavaan.survey::lavaan.survey(lavaan.fit = fit_meas1, survey.design = cluster_design, estimator = "MLR")
-summary(fit_meas1_cluster, fit.measures = TRUE, standardized = TRUE)
-
-## 3.2 Full Structural model
-
-mod_full <-
-  '
-  # Measurement Model
-
-  ## Latent factors
-  wmc_l      =~ 1*bu_bind_acc + rspan_cap + ospan_cap + bu_upd_acc
-  upd_l      =~ 1*bu_upd_acc
-
-  ## Covariances
-  bu_upd_acc ~~ 0*bu_upd_acc
-
-  mat_dep        ~~ threat
-
-  # Structural Model
-  wmc_l ~ mat_dep + unp + threat + age + interrupt + leave
[…]

### 2024-06-25 scripts/2_pipeline/0_parse_liss_files.R

@@ -1,0 +1,32 @@
+
+# Libraries and functions -------------------------------------------------
+
+library(tidyverse)
+library(haven)
+library(glue)
+
+source("scripts/0_custom_functions/liss-functions.R")
+
+
+# LISS Background Variables -----------------------------------------------
+
+walk(
+    list.files("data/liss_data/background", pattern = ".zip$", full.names = T),
+    unzip_liss
+  )
+
+
+# LISS Income Data --------------------------------------------------------
+
+walk(
+  list.files("data/liss_data/income", pattern = ".zip$", full.names = T),
+  unzip_liss
+)
+
+# LISS Crime Data ---------------------------------------------------------
+
+walk(
+  list.files("data/liss_data/crime", pattern = ".zip$", full.names = T),
+  unzip_liss
+)
+

### 2024-06-25 scripts/2_pipeline/1_ivs.R

@@ -71,31 +71,36 @@ background_parsed <- bind_rows(background_vars01, background_vars02) |>
 write_csv(background_parsed, "data/liss_data/background/background_parsed.csv")
 
 
-
 ## 1.3 Material deprivation ----
 
 core_income <- projectlog::read_data(file = "list.files('data/liss_data/income', pattern = '.sav', full.names = TRUE)", read_fun = "haven::read_sav")
 
 
-# in wave 1 to 11, net income was asked for each employer separately
+
 core_income01 <- core_income[1:11] |>
   map(function(x){
     x |>
+      sjlabelled::remove_all_labels() |>
       as_tibble() |>
       select(nomem_encr, matches("(_m|012|021|030|244|245|246|247|248|249|250|252)$")) |>
       rename_with(
         .cols = -nomem_encr,
         ~str_replace_all(string = ., pattern = "ci\\d\\d[a-z]_|ci\\d\\d[a-z]", "q_")
       ) %>%
+      #in wave 1 to 11, net income was asked for each employer separately
      mutate(net_inc_y = across(c(q_012, q_021, q_030)) %>% rowSums(na.rm = T))
   }) |>
   bind_rows() |>
-  separate(col = q_m, into = c("year", "month"), sep = 4)
+  separate(col = q_m, into = c("year", "month"), sep = 4) |>
+  left_join(background_parsed |> select(nomem_encr, year, month, nohouse_encr)) |>
+  group_by(year, nohouse_encr) |>
+  mutate(net_inc_y = sum(net_inc_y, na.rm = T))
 
 # in wave 12 to 16, net income was asked across all employers
 core_income02 <- core_income[12:16] |>
   map(function(x){
     x |>
+      sjlabelled::remove_all_labels() |>
       as_tibble() |>
       select(nomem_encr, matches("(_m|339|244|245|246|247|248|249|250|252)$")) |>
       rename_with(
@@ -108,8 +113,18 @@ core_income02 <- core_income[12:16] |>
   separate(col = q_m, into = c("year", "month"), sep = 4)
 
 core_income_parsed <- bind_rows(core_income01, core_income02) |>
-  select(-c(q_012, q_021, q_030))
[…]

### 2024-06-25 scripts/2_pipeline/3_confirmatory.R

@@ -1,0 +1,250 @@
+
+# Libraries ---------------------------------------------------------------
+
+library(tidyverse)
+library(lavaan)
+
+# 2. Load data ------------------------------------------------------------
+
+clean_data <- read_csv("data/liss_data/clean_data.csv") %>%
+  mutate(
+    inr_m = log(inr_m),
+    inr_cv = log(inr_cv)
+  ) |>
+  mutate(
+    across(c(ends_with('cap'), binding_number_acc, updating_number_acc), ~scale(x = .) |> as.numeric()),
+    across(c(inr_m, inr_cv, threat_comp, p_scar_m, p_scar_cv), ~scale(x = .) |> as.numeric()),
+    age = scale(age) |> as.numeric(),
+    edu = scale(edu) |> as.numeric(),
+    noise = scale(noise) |> as.numeric(),
+    interrupt = case_when(
+      interrupt == 1 ~ 0,
+      interrupt == 2 ~ 1,
+      TRUE ~ interrupt
+    ),
+    leave = case_when(
+      leave == 1 ~ 0,
+      leave == 2 ~ 1,
+      TRUE ~ leave
+    ),
+    age_sq = age^2
+  ) |>
+  mutate(nohouse_encr = as.character(nohouse_encr))
+
+
+# 3. Model specification(s) --------------------------------------------------
+
+## 3.1 Starting specification (does not converge) ----
+mod_meas01 <-
+  '
+   # Measurement Model
+
+   ## Latent factors
+   wmc_l      =~ 1*binding_number_acc + rspan_cap + ospan_cap + updating_number_acc
+   upd_l      =~ 1*updating_number_acc
+
+   ## Covariances
+    updating_number_acc ~~ binding_number_acc
+
+   rspan_cap ~ 0
+   ospan_cap ~ 0
+   binding_number_acc ~ 0
+   updating_number_acc ~ 0
+'
+
+
+fit_meas01 <- lavaan::sem(
+  model = mod_meas01,
+  data = clean_data,
+  estimator = "MLR", # Robust SEs in case of non-normal data
+  missing = "ML",    # FIML for missing data
+  cluster = 'nohouse_encr', # Clustering within
+  se = "robust",
+  orthogonal = TRUE,
[…]

### 2024-06-25 scripts/2_pipeline/3_sem.R

@@ -1,128 +1,0 @@
-
-# Libraries ---------------------------------------------------------------
-
-library(tidyverse)
-library(lavaan)
-
-# 2. Load data ------------------------------------------------------------
-
-clean_data <- read_csv("data/liss_data/clean_data.csv") %>%
-  mutate(
-    across(c(ends_with('cap'), binding_number_acc, updating_number_acc), ~scale(x = .) |> as.numeric())
-  ) |>
-  mutate(nohouse_encr = as.character(nohouse_encr))
-
-
-
-# 3. Model specification(s) --------------------------------------------------
-
-## 3.1 Starting specification (does not converge) ----
-mod_meas01 <-
-  '
-   # Measurement Model
-
-   ## Latent factors
-   wmc_l      =~ 1*binding_number_acc + rspan_cap + ospan_cap + updating_number_acc
-   upd_l      =~ 1*updating_number_acc
-
-   ## Covariances
-    updating_number_acc ~~ 0*updating_number_acc
-    updating_number_acc ~~ binding_number_acc
-
-   rspan_cap ~ 0
-   ospan_cap ~ 0
-   binding_number_acc ~ 0
-   updating_number_acc ~ 0
-'
-
-
-fit_meas01 <- lavaan::sem(
-  model = mod_meas01,
-  data = clean_data,
-  estimator = "MLR", # Robust SEs in case of non-normal data
-  missing = "ML",    # FIML for missing data
-  cluster = 'nohouse_encr', # Clustering within
-  se = "robust",
-  orthogonal = TRUE,
-  auto.cov.lv.x = FALSE,
-  auto.cov.y = TRUE,
-  std.lv = FALSE,
-  fixed.x = FALSE
-)
-
-fit = lavaan::summary(fit_meas01, fit.measures = TRUE, standardized = TRUE)
-
-
-
-## 3.2 Remove covariance ----
-
-mod_meas02 <-
-  '
-   # Measurement Model
-
-   ## Latent factors
-   wmc_l      =~ 1*binding_number_acc + rspan_cap + ospan_cap + updating_number_acc
-   upd_l      =~ 1*updating_number_acc
-
-   ## Covariances
-    updating_number_acc ~~ 0*updating_number_acc
-
-   rspan_cap ~ 0
-   ospan_cap ~ 0
-   binding_number_acc ~ 0
-   updating_number_acc ~ 0
-'
[…]

### 2024-06-25 scripts/2_pipeline/4_exploratory.R

@@ -1,0 +1,204 @@
+
+# Libraries ---------------------------------------------------------------
+
+library(tidyverse)
+library(lavaan)
+
+# 2. Load data ------------------------------------------------------------
+
+clean_data <- read_csv("data/liss_data/clean_data.csv") %>%
+  mutate(
+    inr_m = log(inr_m),
+    inr_cv = log(inr_cv)
+  ) |>
+  mutate(
+    across(c(ends_with('cap'), binding_number_acc, updating_number_acc), ~scale(x = .) |> as.numeric()),
+    across(c(inr_m, inr_cv, threat_comp, p_scar_m, p_scar_cv), ~scale(x = .) |> as.numeric()),
+    age = scale(age) |> as.numeric(),
+    edu = scale(edu) |> as.numeric(),
+    noise = scale(noise) |> as.numeric(),
+    interrupt = case_when(
+      interrupt == 1 ~ 0,
+      interrupt == 2 ~ 1,
+      TRUE ~ interrupt
+    ),
+    leave = case_when(
+      leave == 1 ~ 0,
+      leave == 2 ~ 1,
+      TRUE ~ leave
+    ),
+    age_sq = age^2
+  ) |>
+  mutate(nohouse_encr = as.character(nohouse_encr))
+
+
+load('analysis_objects/confirmatory_results.RData')
+
+
+# 2. Exploratory analysis 1: Associations with separate tasks --------------------------
+
+# These results are presented in the main text.
+
+lm_reg_coef_expl <- bind_rows(
+  lm(rspan_cap ~ inr_cv + inr_m + threat_comp + p_scar_m + p_scar_cv + age + age_sq + noise + interrupt + leave, data = clean_data) |>
+    broom::tidy(conf.int = T) |>
+    mutate(task = "rspan", domain = "WMC"),
+  lm(ospan_cap ~ inr_cv + inr_m + threat_comp + p_scar_m + p_scar_cv + age + age_sq + noise + interrupt + leave, data = clean_data) |>
+    broom::tidy(conf.int = T) |>
+    mutate(task = "Ospan", domain = "WMC"),
[…]

### 2024-06-25 .projectlog/1c2d2e6b017aeb8ce96b1022cc62393e66e7cb55.R
.projectlog data access files

@@ -1,0 +1,13 @@
+### Date: 2024-04-03 16:13:42
+
+### Description: Read subject IDs only of new data collection to filter IV data
+
+
+### For more information on this commit, see the README file, or go to https://github.com/StefanVermeent/liss_wm_profiles_2023/commit/1c2d2e6b017aeb8ce96b1022cc62393e66e7cb55
+
+### Below is the full code that was used to access the data:
+
+
+haven::read_sav('data/liss_data/wm/L_Cognitieve_Vaardigheden_v2_1.0p.sav', col_select = 'nomem_encr') |>
+ dplyr::filter() |>
+ shuffle(data = _, shuffle_vars = 'NULL', long_format = FALSE, seed = 3985843)

### 2024-06-25 .projectlog/71d70ab79326ada925a66e4e695c2e1ae703389a.R

@@ -1,0 +1,13 @@
+### Date: 2024-03-30 16:25:37
+
+### Description: access to all income waves in LISS archive
+
+
+### For more information on this commit, see the README file, or go to https://github.com/StefanVermeent/liss_wm_profiles_2023/commit/71d70ab79326ada925a66e4e695c2e1ae703389a
+
+### Below is the full code that was used to access the data:
+
+
+purrr::map(.x = list.files('data/liss_data/crime', pattern = '.sav', full.names = TRUE), function(x) {haven::read_sav(file = x, col_select = NULL) |>
+ dplyr::filter() |>
+ shuffle(data = _, shuffle_vars = 'NULL', long_format = FALSE, seed = 3985843)})

### 2024-06-25 .projectlog/76539b6a145eef99ba237f8837ee40ba966158ae.R

@@ -1,0 +1,13 @@
+### Date: 2024-04-01 09:58:39
+
+### Description: access to all background variable waves in LISS archive
+
+
+### For more information on this commit, see the README file, or go to https://github.com/StefanVermeent/liss_wm_profiles_2023/commit/76539b6a145eef99ba237f8837ee40ba966158ae
+
+### Below is the full code that was used to access the data:
+
+
+purrr::map(.x = list.files('data/liss_data/background', pattern = '.sav', full.names = TRUE), function(x) {haven::read_sav(file = x, col_select = NULL) |>
+ dplyr::filter() |>
+ shuffle(data = _, shuffle_vars = 'NULL', long_format = FALSE, seed = 3985843)})

### 2024-06-25 .projectlog/869bedb88af275fd01f5dc0a72e2776cea975e1c.R

@@ -1,0 +1,13 @@
+### Date: 2024-03-30 13:21:56
+
+### Description: access to all crime victimization waves in LISS archive
+
+
+### For more information on this commit, see the README file, or go to https://github.com/StefanVermeent/liss_wm_profiles_2023/commit/869bedb88af275fd01f5dc0a72e2776cea975e1c
+
+### Below is the full code that was used to access the data:
+
+
+purrr::map(.x = list.files('data/liss_data/crime', pattern = '.sav', full.names = TRUE), function(x) {haven::read_sav(file = x, col_select = NULL) |>
+ dplyr::filter() |>
+ shuffle(data = _, shuffle_vars = 'NULL', long_format = FALSE, seed = 3985843)})

### 2024-06-25 .projectlog/9dbbb90a82566421d7967010dba755b0faae773f.R

@@ -1,0 +1,11 @@
+### Date: 2024-04-29 11:52:56
+
+### Description: final SEM specification
+
+
+### For more information on this commit, see the README file, or go to https://github.com/StefanVermeent/liss_wm_profiles_2023/commit/9dbbb90a82566421d7967010dba755b0faae773f
+
+### Below is the full code that was used to access the data:
+
+
+NA

### 2024-06-25 .projectlog/a0be5183db79da2d3f752565d29097b8c4a72186.R

@@ -1,0 +1,13 @@
+### Date: 2024-04-04 09:20:20
+
+### Description: access to all working memory measures of new study
+
+
+### For more information on this commit, see the README file, or go to https://github.com/StefanVermeent/liss_wm_profiles_2023/commit/a0be5183db79da2d3f752565d29097b8c4a72186
+
+### Below is the full code that was used to access the data:
+
+
+haven::read_sav('data/liss_data/wm/L_Cognitieve_Vaardigheden_v2_1.0p.sav', col_select = NULL) |>
+ dplyr::filter() |>
+ shuffle(data = _, shuffle_vars = 'NULL', long_format = FALSE, seed = 3985843)

### 2024-06-25 .projectlog/d7dd554d9061e035af12a0268034fbfebb623e71.R

@@ -1,0 +1,13 @@
+### Date: 2024-03-30 16:15:43
+
+### Description: access to all crime victimization waves in LISS archive, v2
+
+
+### For more information on this commit, see the README file, or go to https://github.com/StefanVermeent/liss_wm_profiles_2023/commit/d7dd554d9061e035af12a0268034fbfebb623e71
+
+### Below is the full code that was used to access the data:
+
+
+purrr::map(.x = list.files('data/liss_data/crime', pattern = '.sav', full.names = TRUE), function(x) {haven::read_sav(file = x, col_select = NULL) |>
+ dplyr::filter() |>
+ shuffle(data = _, shuffle_vars = 'NULL', long_format = FALSE, seed = 3985843)})

### 2024-06-25 supplement/supplement.qmd
supplemental materials for Stage 2 manuscript

@@
 load('staged_results.RData')
+load('../analysis_objects/exploratory_results.RData')
+load('../analysis_objects/fig6.RData')
 
@@
 
+# 3. Exploratory analyses
+
+To contextualize our confirmatory (preregistered) findings, we conducted twoexploratory (non-preregistered) analyses.
+Below, we present the results of analyses exploring the associations between adversity and performance on the separate WM tasks using linear regression.
+The other exploratory analyses, which were based on alternative SEM specifications, are presented in the supplemental materials.
+
+We estimated a total of five linear regression models, one per WM task.
+Each model included the same independent variables and covariates as the primary analysis.
+We adjusted for multiple testing across models involving the Rotation Span Task, Operation Span Task, and the binding trials of the Binding-Updating Task, and separately for the updating trials of the Binding-Updating Task, as the former three tasks are primarily conceptualized as WM capacity tasks.
+We also tested for practical equivalence in the same way as for the confirmatory analyses
+
+The results are summarized in Figure 6.
+Threat was negatively associated with performance on the Rotation Span Task ($\beta$ = `r lm_reg_coef_expl |> filter(term == "threat_comp", task == "rspan") |> pull(estimate) |> formatC(digits = 2, width = 2, format = 'f')`, *p* = `r lm_reg_coef_expl |> filter(term == "threat_comp", task == "rspan") |> pull(pvalue_adj) |> formatC(digits = 3, width = 3, format = 'f') |> str_remove(string = _, pattern = "^0")`), Operation Span Task ($\beta$ = `r lm_reg_coef_expl |> filter(term == "threat_comp", task == "Ospan") |> pull(estimate) |> formatC(digits = 2, width = 2, format = 'f')`, *p* = `r lm_reg_coef_expl |> filter(term == "threat_comp", task == "Ospan") |> pull(pvalue_adj) |> formatC(digits = 3, width = 3, format = 'f') |> str_remove(string = _, pattern = "^0")`), and binding trials of the Binding-Updating Task ($\beta$ = `r lm_reg_coef_expl |> filter(term == "threat_comp", task == "binding") |> pull(estimate) |> formatC(digits = 2, width = 2, format = 'f')`, *p* = `r lm_reg_coef_expl |> filter(term == "threat_comp", task == "binding") |> pull(pvalue_adj) |> formatC(digits = 3, width = 3, format = 'f') |> str_remove(string = _, pattern = "^0")`).
+Unpredictability in perceived scarcity was positively associated with performance on the Rotation Span Task ($\beta$ = `r lm_reg_coef_expl |> filter(term == "p_scar_cv", task == "rspan") |> pull(estimate) |> formatC(digits = 2, width = 2, format = 'f')`, *p* = `r lm_reg_coef_expl |> filter(term == "p_scar_cv", task == "rspan") |> pull(pvalue_adj) |> formatC(digits = 3, width = 3, format = 'f') |> str_remove(string = _, pattern = "^0")`).
+None of the types of adversity were significantly associated with performance on the updating trials of the Binding-Updating Task.
+
+In addition, there was some limited evidence for practical equivalence, especially for unpredictability in the income-to-needs ratio, which showed a practically equivalent association with the Operation Span Task, Rotation Span Task, and Updating Task.
+We also found a practically equivalent association between unpredictability in perceived scarcity and the Binding Task.
+
+<br>
+
+```{r}
+#| label: Figure6
+#| fig-width: 6.5
+#| fig-height: 6
+#| dpi: 600
+#| out-width: 6in
+#| fig-cap: | 
+#|   **Figure 6.** Exploratory (non-preregistered) results testing the association between threat, deprivation, and unpredictability on residual variances of separate WM tasks.
+The gray area shows the area of practical equivalence.
+Solid points indicate effects outside the area of practical equivalence, which was true for all effects.
+Standard errors represent the 95% confidence intervals.
+CV = coefficient of variation; INR = income-to-needs ratio; M = mean; WM = working memory.
+fig6
+```
+<br>
+
+# 4. Histograms of independent measures
+
+```{r}
+#| label: Figure7
+#| fig-width: 6.5
+#| fig-height: 6
+#| dpi: 600
+#| out-width: 6in
+#| fig-cap: | 
+#|   **Figure 7.** Histograms for mean perceived material deprivation over time.
+p_scar_m_combn_hist
+```
+
+<br>
+
+```{r}
+#| label: Figure8
+#| fig-width: 6.5
+#| fig-height: 6
+#| dpi: 600
+#| out-width: 6in
+#| fig-cap: | 
+#|   **Figure 8.** Histogram for the mean income-to-needs ratio over time.
+p_scar_m_combn_hist
+```
+
+<br>
+
+```{r}
+#| label: Figure9
+#| fig-width: 6.5
+#| fig-height: 6
+#| dpi: 600
+#| out-width: 6in
+#| fig-cap: | 
+#|   **Figure 9.** Histograms for mean threat exposure over time.
+threat_combn_hist
+
+```
+
+<br>
+
+```{r}
+#| label: Figure10
+#| fig-width: 6.5
+#| fig-height: 6
+#| dpi: 600
+#| out-width: 6in
+#| fig-cap: | 
+#|   **Figure 10.** Histograms for unpredictability in measures of perceived scarcity over time (coefficient of change).
+p_scar_cv_combn_hist
+```
+
+<br>
+
+```{r}
+#| label: Figure11
+#| fig-width: 6.5
+#| fig-height: 6
+#| dpi: 600
+#| out-width: 6in
+#| fig-cap: | 
+#|   **Figure 11.** Histograms for unpredictability in the income-to-needs ratio over time (coefficient of change).
+inr_cv_hist
+```
+
+\pagebreak
+
 # References

### 2024-06-25 supplement/scripts/stage_supp.R

@@ -1,10 +1,11 @@
 library(flextable)
 library(tidyverse)
+library(patchwork)
 
 load("data/pilot_full_data.RData")
 source("scripts/0_custom_functions/binom_test.R")
 source("scripts/0_custom_functions/corr_plot.R")
-
+clean_data <- read_csv("data/liss_data/clean_data.csv")
 
 # set up flextable for tables
 set_flextable_defaults(
@@ -22,16 +23,16 @@ set_flextable_defaults(
 
 pilot_demo <-
   list(
-    m_age = pilot_data_full$dems_age |> mean(na.rm = T) |> round(1),
-    sd_age = pilot_data_full$dems_age |> sd(na.rm = T) |> round(1),
-    max_age = pilot_data_full$dems_age |> max(na.rm = T) |> round(1),
-    min_age = pilot_data_full$dems_age |> min(na.rm = T) |> round(1)
+    m_age = pilot_full_data$dems_age |> mean(na.rm = T) |> round(1),
+    sd_age = pilot_full_data$dems_age |> sd(na.rm = T) |> round(1),
+    max_age = pilot_full_data$dems_age |> max(na.rm = T) |> round(1),
+    min_age = pilot_full_data$dems_age |> min(na.rm = T) |> round(1)
   )
 
 pilot_exclusions <-
   list(
-    acc_ospan_sec_n = pilot_data_full |> filter(ospan_sec_acc*100 < gbinom(45, 0.50)) |> nrow(),
-    acc_rspan_sec_n = pilot_data_full |> filter(rspan_sec_acc*100 < gbinom(45, 0.50)) |> nrow()
+    acc_ospan_sec_n = pilot_full_data |> filter(ospan_sec_acc*100 < gbinom(45, 0.50)) |> nrow(),
+    acc_rspan_sec_n = pilot_full_data |> filter(rspan_sec_acc*100 < gbinom(45, 0.50)) |> nrow()
   )
 
 pilot_clean <- pilot_full_data |>
@@ -90,3 +91,221 @@ pilot_table1 <- pilot_clean |>
 
 save(pilot_demo, pilot_exclusions, pilot_table1, file = "supplement/staged_results.RData")
 
+
+
+# 2. IV histograms --------------------------------------------------------
+
+## 2.1 Perceived deprivation (mean) ----
+
+hist_data <- clean_data |>
[…]

### 2024-06-25 README.Rmd
new top-level README file

@@
 Data access events were automatically captured using custom code, which over the course of this project was collected in the R package `projectlog` [https://stefanvermeent.github.io/projectlog/](https://stefanvermeent.github.io/projectlog/).
-For more information about how tracking worked, Go to the [Open science workflow tab](https://stefanvermeent.github.io/abcd_ddm/opensci_workflow/README.html).
 
@@
   mutate(
+    branch = gert::git_branch_list() |> filter(local == TRUE) |> pull(name),
     name = map_chr(name, function(name) {
       name |> 
-        str_remove("\\[\\[") |>
         str_remove("\\d*$") |> 
@@
         str_to_title()
-    })
-  ) |> 
-  mutate(
-    code = str_remove_all(code, "^code\\s") |> str_replace_all(string = _, "\\|\\>", "\\|\\>\\\n"),
-    link_to_git = paste0(
-       gert::git_remote_list()$url |> str_remove(string = _, pattern = "\\.git$"), "/tree/", commit)
+    }),
+    code = ifelse(name == "Data Access", str_remove_all(code, "^code\\s") |> str_replace_all(string = _, "\\|\\>", "\\|\\>\\\n"), "Not applicable"),
+    link_to_git = ifelse(name == 'Data Access', 
+                         paste0(get_git_url(), "/blob/", branch, "/", commit, ".R"),
+                         paste0(get_git_url(), "/tree/", commit)
+    ),
+    hash = ifelse(name != "Data Access", "Not applicable", hash),
+
     ) |>
@@
 milestones |> 
-  pwalk(function(name,commit,author,time,description,hash,code,link_to_git){
+  pwalk(function(name,commit,author,time,description,hash,code,branch,link_to_git){
 
@@
 
-      url <- gert::git_remote_list()$url |> 
-        str_remove(string = _, "\\.git$") |> 
-        paste0(paste0("/commit/", commit))
-
-      #  url <- "<anonymized repository>"
+      url <- get_git_url() |> 
+        paste0("/commit/", commit)
 
@@
 
-branch <- gert::git_branch_list() |> filter(local == TRUE) |> pull(name)
-
 # Print milestones

### 2024-06-25 README.Rmd
new rendering of README file

@@
 ```{r results ='asis', echo = FALSE}
-milestones <- gert::git_tag_list() |> 
-  mutate(
-    tags = pmap(list(name = name, ref = ref, commit = commit), function(name, ref, commit) {
-      gert::git_commit_info(commit) |> 
-        as_tibble() |> 
-        select(commit = id, author, time, message) 
-    })
-  ) |> 
-  select(-commit, -ref) |> 
-  unnest(tags) |> 
-  separate(message, into = c("description", "hash", "code"), sep = "\\n") |> 
-  mutate(
-    branch = gert::git_branch_list() |> filter(local == TRUE) |> pull(name),
-    name = map_chr(name, function(name) {
-      name |> 
-        str_remove("\\d*$") |> 
-        str_replace("_", " ") |> 
-        str_to_title()
-    }),
-    code = ifelse(name == "Data Access", str_remove_all(code, "^code\\s") |> str_replace_all(string = _, "\\|\\>", "\\|\\>\\\n"), "Not applicable"),
-    link_to_git = ifelse(name == 'Data Access', 
-                         paste0(get_git_url(), "/blob/", branch, "/", commit, ".R"),
-                         paste0(get_git_url(), "/tree/", commit)
-    ),
-    hash = ifelse(name != "Data Access", "Not applicable", hash),
+milestones <- projectlog::generate_milestone_overview()
 
-    ) |>
-  arrange(desc(time))
-
-# Write scripts including code to .projectlog folder
-milestones |> 
-  pwalk(function(name,commit,author,time,description,hash,code,branch,link_to_git){
-
-    if(name == "Data Access"){
-      code <- code |> 
-        str_replace_all(string = _, "\\|> ", "\\|>\n") |> 
-        str_remove(string = _, "^code ")
-
-      url <- get_git_url() |> 
-        paste0("/commit/", commit)
-
-      script <-
-        glue(
-          "
-          ### Date: {time}\n
-          ### Description: {description}\n\n
-          ### For more information on this commit, see the README file, or go to {url}\n
-          ### Below is the full code that was used to access the data:\n\n
-          {code}"
-        )
-
-      writeLines(script, con = paste0(".projectlog/", commit, ".R"))
-    }
-  })
-
-# Print milestones
-glue_data(milestones, "  - **[{time}]({link_to_git}): {description}**
-        - **Milestone:** {name}
-        - **Data MD5 hash**: {hash}
-        - [Link to code snippet](https://github.com/StefanVermeent/liss_wm_profiles_2023/blob/{branch}/.projectlog/{commit}.R)
-        \n
-    ")
-
+milestones
 ```

### 2024-06-25 manuscript/manuscript.qmd
Add submission of stage 2 manuscript to README

@@
 library(tidyr)
+library(readr)
 library(tibble)
@@
 
+clean_data <- read_csv('../data/liss_data/clean_data.csv')
+load("../analysis_objects/confirmatory_results.RData") # SEM results
+load("../analysis_objects/exploratory_results.RData") # SEM results
+load("../analysis_objects/pca.RData") # PCA results for threat and unpredictability in perceived scarcity
+load("../analysis_objects/tables.RData")
+load("../analysis_objects/fig5.RData")
+load("../analysis_objects/fig6.RData")
 ```
@@
 All scripts and materials needed to reproduce the findings are available on the article's Github repository ([https://stefanvermeent.github.io/liss_wm_profiles_2023/](https://stefanvermeent.github.io/liss_wm_profiles_2023/)).
-We also include instructions on how to reproduce each step of our analyses, as well as simulated data to ensure computational reproducibility. 
+We also include instructions on how to reproduce each step of our analyses.
 In this paper, we make use of data from the LISS panel (Longitudinal Internet studies for the Social Sciences) managed by the non-profit research institute Centerdata (Tilburg University, the Netherlands).
@@
 
-# Proposal abstract
+# Abstract
 
@@
 However, some recent adaptation-based studies suggest that certain types of adversity can leave intact, or even enhance, the ability to rapidly update information in WM.
-One key challenge is that WM capacity and updating tasks tend to covary.
-This is likely due to the fact that both types of tasks require the creation and maintenance of bindings in WM; links between mental representations of information in WM.
+One key challenge is that WM capacity and updating tasks tend to covary, as both types of tasks require the creation and maintenance of bindings in WM; links between mental representations of information in WM.
 To estimate the associations between adversity and different processes in WM, we need to isolate variance in performance related to WM capacity from variance in performance related to updating ability.
-In this Registered Report, we combine archived and newly collected data in the Dutch Longitudinal Internet studies for the Social Sciences (LISS) panel, which includes a representative sample of the Dutch adult population.
-Participants completed three WM tasks: two complex span tasks and a task measuring both binding and updating of information.
-In addition, we will estimate participants' exposure to neighborhood threat, material deprivation, and unpredictability.
-Using structural equation modeling, we will estimate associations between the three types of adversity and latent estimates of WM capacity and updating.
-These findings will advance our theoretical understanding of how adversity is associated with WM, which will aid interventions aimed at alleviating performance difficulties and leveraging areas of strength.
+In this Registered Report, participants from the Dutch Longitudinal Internet studies for the Social Sciences (LISS) panel completed three WM tasks: two complex span tasks and a task measuring both binding and updating of information.
+In addition, we estimated participants' exposure to neighborhood threat, material deprivation, and unpredictability.
+We estimated associations between the three types of adversity and latent estimates of WM capacity and updating using structural equation modeling.
+We did not find consistent associations between adversity and WM capacity or updating, nor did we find evidence that the associations were practically equivalent to zero.
+Our results show that adversity researchers should account for overlap in WM tasks when estimating specific WM abilities.
 
@@
 With deficit theories focusing on WM capacity and adaptation-based theories on WM updating, we may wonder how capacity and updating are related to each other.
-Performance on tasks measuring WM capacity and updating tend to be substantially correlated [in the range of .20-.50\; @frischkorn_2022; @loffler_2022].
+Performance on tasks measuring WM capacity and updating tend to be substantially correlated [in the range of .20-.50\; @frischkorn_2022; @loffler_2024].
 This overlap appears to stem from shared demands of both types of tasks, in particular the need to create and maintain arbitrary bindings [@gruszka_2017; @oberauer_2009; @wilhelm_2013].
@@
 
-In this study, we will estimate associations between latent estimates of WM capacity and updating with three types of adversity: threat, deprivation, and unpredictability. 
+In this study, we estimated associations between latent estimates of WM capacity and updating with three types of adversity: threat, deprivation, and unpredictability. 
 Together, these adversity types capture key dimensions in contemporary models of adversity [@ellis_2009; @ellis_2022; @mclaughlin_2016; @mclaughlin_2021]. 
 Threat refers to experiences involving the potential for harm imposed by others. 
-We focus on perceived neighborhood violence, the extent to which an individual reports having been exposed to acts of violence in their neighborhood. 
+We focused on perceived neighborhood violence, the extent to which an individual reports having been exposed to acts of violence in their neighborhood. 
 Deprivation refers to having a low level of resources. 
-We focus on perceived material deprivation, a (perceived) lack of access to material resources. 
+We focused on perceived material deprivation, a (perceived) lack of access to material resources. 
 Unpredictability refers to variation in material deprivation over time. 
 This definition is inspired by, but deviates from the harshness-unpredictability framework, in which unpredictability is defined as stochastic variation in harshness (age-specific rates in morbidity and mortality) over space and time [@ellis_2009; @ellis_2022].
-We will not calculate unpredictability in neighborhood threat given that participants have at most six timepoints, and often as few as one or two, which is insufficient to calculate variation over time [@walasek_2024].
+We did not calculate unpredictability in neighborhood threat given that participants had at most six timepoints, and often as few as one or two, which is insufficient to calculate variation over time [@walasek_2024].
 
-We will address three research questions.
+We addressed three research questions.
 First, what is the association between adversity and WM capacity?
@@
 
-We will evaluate evidence for deficit- and adaptation-based frameworks (see Figure 1A for a visual summary, and Appendix 1 for the study design plan). 
+We evaluated evidence for deficit- and adaptation-based frameworks (see Figure 1A for a visual summary, and Appendix 1 for the study design plan). 
 Crucially, as deficit and adaptation processes can operate in concert [@frankenhuis_2020], we could find support (or lack thereof) for both frameworks in the same model. 
-We distinguish between three between-person data patterns: (1) lowered performance, (2) enhanced performance, and (3) intact performance.
-We define lowered performance as a statistically significant negative association between a type of adversity and WM capacity or updating (irrespective of effect size).
-We define enhanced performance as a statistically significant positive association between a type of adversity and WM capacity or updating (irrespective of effect size).
-We define intact performance as an association between a type of adversity and WM capacity or updating that has a standardized effect smaller than 0.1 *and* larger than -0.1---even if the effect is statistically different from zero---which we will test using Two One-Sided T-Tests (TOST) equivalence testing [see the 'Primary analyses' section\; @lakens_2018].
+We distinguished between three between-person data patterns: (1) lowered performance, (2) enhanced performance, and (3) intact performance.
+We defined lowered performance as a statistically significant negative association between a type of adversity and WM capacity or updating (irrespective of effect size).
+We defined enhanced performance as a statistically significant positive association between a type of adversity and WM capacity or updating (irrespective of effect size).
+We defined intact performance as an association between a type of adversity and WM capacity or updating that has a standardized effect smaller than 0.1 *and* larger than -0.1---even if the effect is statistically different from zero---which we tested using Two One-Sided T-Tests (TOST) equivalence testing [see the 'Primary analyses' section\; @lakens_2018].
 
@@
 
-We also have two expectations based on prior studies. 
-First, we expect the association between material deprivation and WM capacity to be more negative than the associations with unpredictability and neighborhood threat. 
+We also had two expectations based on prior studies. 
+First, we expected the association between material deprivation and WM capacity to be more negative than the associations with unpredictability and neighborhood threat. 
 This follows from findings showing that cognitive abilities are more negatively associated with cognitive deprivation than threat [@salhi_2021; @sheridan_2020].
 Although cognitive and material deprivation are distinct types of deprivation, they tend to be correlated, and are both associated with limited access to resources that stimulate cognitive development and functioning [@bradley_2001; @lurie_2024; @rosen_2019].
-Therefore, we expect that their associations with WM will have comparable effect sizes. 
+Therefore, we expected that their associations with WM would have comparable effect sizes. 
 Second, researchers have hypothesized that WM updating is particularly adaptive in unpredictable and threatening environments, as it facilitates keeping track of unpredictable changes and sudden threats.
-Therefore, we expect WM updating to be associated with unpredictability and neighborhood threat, but not with material deprivation [@young_2018; but see @young_2022].
+Therefore, we expected WM updating to be associated with unpredictability and neighborhood threat, but not with material deprivation [@young_2018; but see @young_2022].
 
@@
 Panel A depicts the most likely between-person data patterns based on previous literature, and whether we would consider them consistent with deficit and adaptation frameworks (see the main text for more details).
-Panel B depicts an overview of the Structural Equation Model.
+Panel B depicts an overview of the preregistered Structural Equation Model.
+Note that this model differs slightly from the final model (see Figure 4).
 Ellipses represent latent variables, rectangles represent manifest variables, and circles represent residual variances.
 Unidirectional solid lines represent factor loadings, bidirectional solid lines represent covariances, and dashed lines represent regression paths.
-All four manifest WM measures load on a latent WM capacity factor, reflecting the fact that people have to hold information active in WM on all tasks.
-We fix the loading of WM capacity on the Binding Task to 1, reflecting the idea that the ability to create and maintain bindings is the main limiting factor in WM capacity [@gruszka_2017; @oberauer_2009; @wilhelm_2013].
-WM updating is modeled as a latent factor capturing the residual variance in the updating task after accounting for variance related to WM capacity.
+All four manifest WM measures loaded on a latent WM capacity factor, reflecting the fact that people have to hold information active in WM on all tasks.
+We fixed the loading of WM capacity on the Binding Task to 1, reflecting the idea that the ability to create and maintain bindings is the main limiting factor in WM capacity [@gruszka_2017; @oberauer_2009; @wilhelm_2013].
+WM updating was modeled as a latent factor capturing the residual variance in the updating task after accounting for variance related to WM capacity.
 INR = income-to-needs ratio; Perc.
@@
 
-Our study will include 800 participant who were randomly sampled from the Longitudinal Internet studies for the Social Sciences (LISS) panel [@scherpenzeel_2011].
+Our study included 800 participant who were randomly sampled from the Longitudinal Internet studies for the Social Sciences (LISS) panel [@scherpenzeel_2011].
 The LISS panel is a representative probability sample of roughly 5,000 Dutch households (~7,500 individuals) drawn from the population register by Statistics Netherlands on an invite-only basis.
@@
 In addition, participants can complete additional online questionnaires every month, with variable content.
-The current study integrates two data sources.
+The current study integrated two data sources.
 First, our sample of 800 participants participated in a new LISS study between October 2023 and February 2024 (hereafter referred to as 'newly collected data'), in which we included a measure of neighborhood threat and multiple measures of working memory.
-Second, we will access data that were previously collected in LISS (hereafter referred to as 'the LISS archive').
+Second, we accessed data that were previously collected in LISS (hereafter referred to as 'the LISS archive').
 See Figure 2 for a visual overview of the data sources and their measurement timepoints.
-We signed a contract with LISS stipulating that we will receive access to the newly collected data only after Stage 1 acceptance of this Registered Report.
+We signed a contract with LISS stipulating that we would receive access to the newly collected data only after Stage 1 acceptance of this Registered Report.
 
@@
 Overview of the different data sources used in this study.
-We distinguish between measures taken from the LISS data archive and measures that were newly collected in our own study between October 2023 and February 2024.
+We distinguished between measures taken from the LISS data archive and measures that were newly collected in our own study between October 2023 and February 2024.
 Perceived scarcity and income were collected yearly in the full panel from 2008 – 2023.
@@
 In the newly collected data, we collected data on a measure of neighborhood threat and multiple measures of working memory.
-Note that participants may not have data across all timepoints of the archived studies because they joined the LISS panel more recently or because they did not participate in each wave.
+Note that participants did not have data across all timepoints of the archived studies because they joined the LISS panel more recently or because they did not participate in each wave.
 knitr::include_graphics("fig/fig2.png")
@@
 We based our power analysis on simulations reported by @kretzschmar_2019, determining the required sample size to detect a small effect size ($\beta$ = 0.1) with at least 90% power at $\alpha$ = 0.05.
-Assuming a reliability of at least 0.7 [which is typical for WM tasks with a number of trials similar to ours\; e.g., @wilhelm_2013], we would require a sample size of *N* = 730.
+Assuming a reliability of at least 0.7 [which is typical for WM tasks with a number of trials similar to ours\; e.g., @wilhelm_2013], we required a sample size of *N* = 730.
 Anticipating some exclusions, we decided to include 800 participants.
@@
 To ensure sufficient representation of people from lower socioeconomic backgrounds, *half* the total sample was sampled from participants who reported one or more of the following at least once in the three years: (1) a monthly income < €1,500, (2) HAVO or VWO as highest completed education (which are the two highest levels in Dutch secondary education), or (3) a score of 4 or lower on the 'ladder of life' ("If you imagine a ‘ladder of life’, where the first step represents the worst possible life, and the tenth (top) step the best possible life, on what step would you place yourself?").
-Participants will be excluded if they (1) switched to and interacted with other browser tabs *during* one or more of the cognitive tasks, (2) did not perform above chance level.
+Participants were excluded if they (1) switched to and interacted with other browser tabs *during* one or more of the cognitive tasks, (2) did not perform above chance level on the secondary processing tasks.
+The final sample consisted of `r nrow(clean_data)` participants.
+
+```{r}
+table1 |> 
+  flextable::autofit() |> 
+  add_header_row(
+    values = " ",
+    colwidths = 2
+  ) |> 
+  flextable::compose(
+    i = 1, j = 1, 
+    as_paragraph(as_b("Table 1. "), "Descriptive statistics."),
+    part = "header"
+  ) |> 
+  border_remove() |> 
+  border(i = 2, border.top = fp_border_default(), part = "header") |>  
+  border(i = 2, border.bottom = fp_border_default(), part = "header") |> 
+  border(i = 15, border.bottom = fp_border_default(), part = "body") 
+```
 
 ## Measures
+
+All independent variables, except for the income-to-needs ratio (INR) consisted of multiple items and/or scales.
+If all correlations between the items/scales were equal to or larger than .60 (i.e., indicating a "strong" correlation), then we computed a uniformly weighted average.
+If the correlation was lower than .60, we applied Principal Component Analysis (PCA) to the averaged measures and extracted only the first principal component score.
+We present bivariate correlations in Table 2, and histograms for all independent measures in the supplemental materials.
 
@@
 unsafe", 2) "do not respond to a call at the door because you feel that it is unsafe", 3) "leave valuable items at home to avoid theft or robbery in the street?", 4) "make a detour, by car or on foot, to avoid unsafe areas?" on a scale of 1 ("(Almost) never"), 2 ("Sometimes"), or 3 ("Often").
-We will recode these items so that 0 indicates "(Almost) never". 
-We then sum the responses within each wave for which participants have data, and calculate an average across the waves.
+We recoded these items so that 0 indicated "(Almost) never". 
+We then summed the responses within each wave for which participants had data, and calculated an average across the waves.
 
@@
 Participants answered these questions on a scale of 1 ("Completely disagree") to 7 ("Completely agree").
-We will compute an average of the seven items.
+We computed an average of the seven items.
 
@@
 We included seven items concerning exposure to crime: (1) burglary or attempted burglary; (2) theft from their car; (3) theft of their wallet or purse, handbag, or other personal possession; (4) wreckage of their car or other private property; (5) intimidation by any other means; (6) maltreatment of such serious nature that it required medical attention; (7) maltreatment that did not require medical attention.
-We compute a variety score by summing the exposures to *unique* types of crime across all waves.
-Thus, if a participants reports exposure to the same type of crime on separate waves, this will count as one exposure in the total score [@sweeten_2012].
+We computed a variety score by summing the exposures to *unique* types of crime across all waves.
+Thus, if a participants reported exposure to the same type of crime on separate waves, this counted as one exposure in the total score [@sweeten_2012].
 
-**Neighborhood threat composite.** We will first compute an average across time for each measure separately (i.e., the two measures of neighborhood crime and the measure of crime victimization).
-If all correlations between averaged measures are equal to or larger than .60 (i.e., indicating a "strong" correlation), then we will compute a uniformly weighted average.
-If the correlation is lower than .60, we will apply Principal Component Analysis (PCA) to the averaged measures and extract only the first principal component score.
+**Neighborhood threat composite.** We first computed an average across time for each measure separately (i.e., the two measures of neighborhood crime and the measure of crime victimization).
+Because correlations were below .60 (see Table 2), we then used PCA to extract only the first principal component score (R^2 = .20).
+The threat component was most strongly determined by the NVS (`r pca_threat_fit$loadings[3] |> formatC(digits = 2, width = 2, format = 'f')`), followed by the perceived neighborhood crime scale from the LISS archive (`r pca_threat_fit$loadings[1] |> formatC(digits = 2, width = 2, format = 'f')`) and crime victimization (`r pca_threat_fit$loadings[2] |> formatC(digits = 2, width = 2, format = 'f')`).
 
@@
 
-We will measure material deprivation with two separate indicators: perceived scarcity and the income-to-needs ratio.
+We measured material deprivation with two separate indicators: perceived scarcity and the income-to-needs ratio.
 
-**Perceived scarcity (mean).** We will use a few items from the LISS archive that were collected on a yearly basis between 2008 and 2023 ([https://doi.org/10.57990/1gr4-bf42](https://doi.org/10.57990/1gr4-bf42)) to index perceived scarcity.
+**Perceived scarcity (mean).** We used a few items from the LISS archive that were collected on a yearly basis between 2008 and 2023 ([https://doi.org/10.57990/1gr4-bf42](https://doi.org/10.57990/1gr4-bf42)) to index perceived scarcity.
 First, participants indicated how hard or easy it currently is to live off the income of their household, on a scale from 0 (very hard) to 10 (very easy).
 Second, participants were asked to choose which of the following best applied to their current situation: (1) "we are accumulating debt"; (2) "we are somewhat eating into savings"; (3) "we are just managing to make ends meet"; (4) "we have a little bit of money to spare"; (5) "we have a lot of money to spare".
-Responses will be reverse-coded, so that higher scores indicate a worse financial situation.
+Responses were reverse-coded, so that higher scores indicated a worse financial situation.
 Third, participants answered which of the following issues they were confronted with at present (0 = no, 1 = yes): (1) "having trouble making ends meet"; (2) unable to quickly replace things that break"; (3) "having to lend money for necessary expenditures"; (4) "running behind in paying rent/mortgage or general utilities"; (5) "debt collector/bailiff at the door in the last month"; (6) "received financial support from family or friends in the last month".
 
-We will first compute the average across time for each item separately, and examine correlation between the item averages.
-We will follow the same guidelines as with neighborhood threat to determine if we will compute a uniformly weighted average or extract the first principal component using PCA.
+We first computed the average across time for each item separately.
+Because correlations were all above .60, we calculated a uniformly weighted average.
 
-**Income-to-needs (mean).** We will calculate an income-to-needs ratio for each year using monthly self-reported net household income from the LISS archive ([https://doi.org/10.57990/qn3k-as78](https://doi.org/10.57990/qn3k-as78)).
-Zero values in household income will be set to missing, as these could either indicate the lack of an income or an unwillingness to disclose the income.
-If monthly household income is missing (or zero) for an entire year for a participant, we will use, if available, the yearly net household income they reported in the LISS archive ([https://doi.org/10.57990/1gr4-bf42](https://doi.org/10.57990/1gr4-bf42)), dividing it by 12 to obtain a monthly estimate.
-First, we will divide the average income per year by the *poverty threshold*, as determined by Statistics Netherlands (CBS) [@brakel_2023\; CBS, personal communication, December 15, 2023].
-As thresholds are only provided for households with up to three children, we will apply the threshold of a household with three children to households with more than three children.
-Likewise, we will apply the threshold of a household with two adults for households that contain three or more adults.
-Second, we will calculate the average within-person income-to-needs ratio for each year by averaging across the monthly income-to-needs estimates.
-We will reverse code the yearly income-to-needs ratio so that higher scores indicate more deprivation.
+**Income-to-needs (mean).** We calculated an income-to-needs ratio for each year using monthly self-reported net household income from the LISS archive ([https://doi.org/10.57990/qn3k-as78](https://doi.org/10.57990/qn3k-as78)).
+Zero values in household income were set to missing, as these could either indicate the lack of an income or an unwillingness to disclose the income.
+If monthly household income is missing (or zero) for an entire year for a participant, we used, if available, the yearly net household income they reported in the LISS archive ([https://doi.org/10.57990/1gr4-bf42](https://doi.org/10.57990/1gr4-bf42)), dividing it by 12 to obtain a monthly estimate.
+First, we divided the average income per year by the *poverty threshold*, as determined by Statistics Netherlands (CBS) [@brakel_2023\; CBS, personal communication, December 15, 2023].
+As thresholds are only provided for households with up to three children, we applied the threshold of a household with three children to households with more than three children.
+Likewise, we applied the threshold of a household with two adults for households that contained three or more adults.
+Second, we calculated the average within-person income-to-needs ratio for each year by averaging across the monthly income-to-needs estimates.
 
@@
 
-**Perceived scarcity (SD/mean).** This measure is based on the same items as outlined above (see Perceived scarcity (mean)).
-We will compute unpredictability over time in perceived scarcity using the coefficient of variation, which is the within-person standard deviation across years divided by the mean [@key_2017; @liu_2022; @ugarte_2023; @walasek_2024; @young_2020].
+**Perceived scarcity (SD/mean).** This measure was based on the same items as outlined above (see Perceived scarcity (mean)).
+We computed unpredictability over time in perceived scarcity using the coefficient of variation, which is the within-person standard deviation across years divided by the mean [@key_2017; @liu_2022; @ugarte_2023; @walasek_2024; @young_2020].
 The mean and standard deviation in income have been found to be strongly negatively correlated, indicating that people with lower incomes tend to experience less variability in income [@li_2018; @young_2023].
@@
 
-To compute an overall measure of unpredictability in perceived scarcity, we will follow the same procedure as with mean perceived scarcity, but based on the coefficient of variation.
+We first computed the standard deviation across time for each item separately.
+because correlations were below .60 (see Table 2), we then used PCA to extract only the first principal component score (R^2 = .38).
+The perceived unpredictability component was almost fully determined by the item about people's current situation (`r pca_p_scar_fit$loadings[3] |> formatC(digits = 2, width = 2, format = 'f')`), followed by difficulties to live off income (`r pca_p_scar_fit$loadings[2] |> formatC(digits = 2, width = 2, format = 'f')`) and financial troubles (`r  pca_p_scar_fit$loadings[1] |> formatC(digits = 2, width = 2, format = 'f')`).
 
-**Income-to-needs (SD/mean).** Similar to perceived scarcity, we will compute unpredictability over time in the income-to-needs ratio using the coefficient of variation.
+**Income-to-needs (SD/mean).** Similar to perceived scarcity, we computed unpredictability over time in the income-to-needs ratio using the coefficient of variation.
+
+\pagebreak
+
+```{r}
+table2 |> 
+  flextable() |> 
+  flextable::set_header_labels(Variable = "") |> 
+  flextable::autofit() |> 
+  border_remove() %>% 
+  border(i = 1, border.top = fp_border_default(), part = "header") %>% 
+  border(i = 1, border.bottom = fp_border_default(), part = "header") %>% 
+  border(i = 20, border.bottom = fp_border_default(), part = "body") %>% 
+  add_header_row(
+    values = " ",
+    colwidths = 15
+  ) |> 
+  flextable::compose(
+    i = 1, j = 1, 
+    as_paragraph(as_b("Table 2. "), "Spearman correlation between the main independent variables."),
+    part = "header"
+  ) |> 
+  add_footer_row(
+    values = " ",
+    colwidths = 15
+  ) |> 
+  flextable::compose(
+    i = 1, j = 1, 
+    as_paragraph(as_i("Note: "), "* = ", as_i("p"), " < .05, ** = ", as_i("p"), " < .01, *** = ", as_i("p"), " < .001.
+", "CV = coefficient of variance, INR = income-to-needs ratio, M = mean, Perc.
+Scarcity = perceived scarcity"),
+    part = "footer"
+  ) |> 
+  autofit()
+```
+
+<br>
 
@@
 
-Participants first practice the letter task (three times), then the math task (eight times), and then the full task (three times).
-If they perform at or below chance, they have the opportunity to either repeat a part or advance to the next part.
-After practicing, participants complete 9 test trials, with a total of 45 recall items and 45 math items.
-We will compute an operation span score by calculating the proportion of letters recalled in the correct sequential position across trials [@conway_2005].
+Participants first practiced the letter task (three times), then the math task (eight times), and then the full task (three times).
+If they performed at or below chance, they had the opportunity to either repeat a part or advance to the next part.
+After practicing, participants completed 9 test trials, with a total of 45 recall items and 45 math items.
+We computed an operation span score by calculating the proportion of letters recalled in the correct sequential position across trials [@conway_2005].
 
-**Rotation Span Task.** The Rotation Span Task (Figure 2B) is similar to the Operation Span Task and is adopted from @wilhelm_2013.
+**Rotation Span Task.** The Rotation Span Task (Figure 2B) is similar to the Operation Span Task and was adopted from @wilhelm_2013.
 On each trial, the task is to memorize the orientation of a sequence of arrows in the correct order.
@@
 
-Participants first practice the arrow task (three times), then the letter task (eight times), and then the full task (three times).
-If they perform at or below chance, they have the opportunity to either repeat a part or to advance to the next part.
-After practicing, participants complete 12 test trials, with a total of 45 recall items and 45 letter items.
-We will compute a rotation span score by calculating the proportion of arrows recalled in the correct sequential position across trials [@conway_2005].
+Participants first practiced the arrow task (three times), then the letter task (eight times), and then the full task (three times).
+If they performed at or below chance, they had the opportunity to either repeat a part or to advance to the next part.
+After practicing, participants completed 12 test trials, with a total of 45 recall items and 45 letter items.
+We computed a rotation span score by calculating the proportion of arrows recalled in the correct sequential position across trials [@conway_2005].
 
-**Binding-Updating Task.** The Binding-Updating task (Figure 2C) is adopted from @wilhelm_2013.
+**Binding-Updating Task.** The Binding-Updating task (Figure 2C) was adopted from @wilhelm_2013.
 On each trial, participants see a 3×3 grid, with a fixation cross in the central cell.
@@
 
-Participants first complete four practice trials.
-If they perform at or below chance, they have the opportunity to either repeat the practice trials or to advance to the actual task.
-After practicing, they complete 18 test trials, of which nine are binding-only (24 recall items in total) and nine are updating trials (24 recall items in total).
-We will compute a binding score by calculating the overall recall accuracy (%) across trials with zero updating steps.
-We will compute an updating score by calculating the overall recall accuracy (%) across trials with one or more updating steps.
+Participants first completed four practice trials.
+If they performed at or below chance, they had the opportunity to either repeat the practice trials or to advance to the actual task.
+After practicing, they completed 18 test trials, of which nine were binding-only (24 recall items in total) and nine were updating trials (24 recall items in total).
+We computed a binding score by calculating the overall recall accuracy (%) across trials with zero updating steps.
+We computed an updating score by calculating the overall recall accuracy (%) across trials with one or more updating steps.
 
@@
 The working memory data and one of the neighborhood threat indices were collected through October-December 2023, prior to submitting the Stage 1 protocol.
-These data will only be made available to the first author after Stage 1 acceptance, as stipulated in a signed contract with LISS.
-During planning of the study, the first author accessed to the LISS data archive and inspected three waves of the LISS data containing the items about neighborhood safety and crime exposure, as well as the three most recent monthly data collections containing basic demographic info.
+These data were only made available to the first author after Stage 1 acceptance, as stipulated in a signed contract with LISS.
+During planning of the study, the first author accessed the LISS data archive and inspected three waves of the LISS data containing the items about neighborhood safety and crime exposure, as well as the three most recent monthly data collections containing basic demographic info.
 The purpose was to ascertain the number of individuals who had finished the previous waves in the LISS data archive and were presently still participating in the panel (i.e., to see if we could reasonably create a link between the LISS data archive and newly collected data).
@@
 See Figure 1B for an overview of the model specification.
-We will fit a single model containing all adversity measures using the *lavaan* R package [@rosseel_2012].
-We will use robust maximum likelihood estimation in case any variable is non-normally distributed.
-Missing data will be handled using full information maximum likelihood (FIML).
-If participants are from the same household, this clustering within families will be accounted for using the *lavaan.survey* R package [@oberski_2014].
+We fitted a single model containing all adversity measures using the *lavaan* R package [@rosseel_2012].
+We used robust maximum likelihood estimation to account for non-normality.
+Missing data were handled using full information maximum likelihood (FIML).
+We accounted for clustering within families using the *lavaan.survey* R package [@oberski_2014].
 
-WM capacity will be estimated as a latent factor loading on all outcome measures.
-In addition, we will estimate WM updating as a latent factor capturing residual variance in the updating measure.
-Thus, this factor accounts for updating-specific variance after accounting for WM capacity.
-We will estimate the effect of each adversity type (dashed lines in Figure 1B) through regression analyses.
-Each association will be controlled for: (1) age in years ; (2) the quadratic effect of age; (2) environmental noise ("How noisy was your environment during the memory tasks", rated on a scale of 1 (very little noise) to 5 (a lot of noise)); (3) two items measuring interruptions ("Where you at any moment interrupted during the memory tasks?" and "Did you at any moment during the memory tasks leave your computer?", rated as yes or no).
-Goodness of fit will be assessed using the comparative fit index (CFI) and the root mean square error of approximation (RMSEA).
-CFI values \> .90 and RMSEA values \< .08 will be interpreted as acceptable model fit, and CFI values \> .95 and RMSEA values \≤ .06 as good model fit [@hu_1999].
+WM capacity was estimated as a latent factor loading on all outcome measures.
+In addition, we estimated WM updating as a latent factor capturing residual variance in the updating measure.
+Thus, this factor accounted for updating-specific variance after accounting for WM capacity.
+We estimated the effect of each adversity type (dashed lines in Figure 1B) through regression analyses.
+Each association was controlled for: (1) age in years ; (2) the quadratic effect of age; (2) environmental noise ("How noisy was your environment during the memory tasks", rated on a scale of 1 (very little noise) to 5 (a lot of noise)); (3) two items measuring interruptions ("Where you at any moment interrupted during the memory tasks?" and "Did you at any moment during the memory tasks leave your computer?", rated as yes or no).
+Goodness of fit was assessed using the comparative fit index (CFI) and the root mean square error of approximation (RMSEA).
+CFI values \> .90 and RMSEA values \< .08 were interpreted as acceptable model fit, and CFI values \> .95 and RMSEA values \≤ .06 as good model fit [@hu_1999].
 
-We anticipate that we may have to optimize the model further in case of bad model fit, and will therefore estimate the model in two steps to prevent bias.
-First, we will construct the measurement model of WM, without including the adversity measures.
-This step will be carried out prior to accessing any of the adversity measures.
-Once we obtain at least acceptable model fit, we will access and add the adversity measures to the model.
-This procedure will be tracked and timestamped on the GitHub repository using the procedure outlined above.
-We will control for multiple testing using the false discovery rate [@benjamini_1995; @cribbie_2007].
+We anticipated that we may have to optimize the model further in case of bad model fit, and therefore planned to estimated the model in two steps to prevent bias.
+First, we constructed the measurement model of WM, without including the adversity measures.
+This step was planned to be carried out prior to accessing any of the adversity measures.
+Once we obtained at least acceptable model fit, we accessed and added the adversity measures to the model.
+This procedure was tracked and timestamped on the GitHub repository using the procedure outlined above.
+We controlled for multiple testing using the false discovery rate [@benjamini_1995; @cribbie_2007].
 
-To statistically test whether small effects are practically equivalent to zero---suggesting intact performance---we will use Two One-Sided T-tests (TOST) equivalence testing [@lakens_2018], using -0.1 and 0.1 as equivalence bounds.
+To statistically test whether small effects were practically equivalent to zero---suggesting intact performance---we used Two One-Sided T-tests (TOST) equivalence testing [@lakens_2018], using -0.1 and 0.1 as equivalence bounds.
 TOST equivalence testing allows us to conclude intact performance based on a significant effect, rather than erroneously interpreting a non-significant effect as evidence for the absence of an effect.
-We consider any effect that falls within this region to reflect practical equivalence, that is, a between-person difference in performance that is practically equivalent to zero. 
-TOST provides two *p*-values, one testing against the upper bound and one testing against the lower bound; we will report only the largest of the two *p*-values.
+We considered any effect that fell within this region to reflect practical equivalence, that is, a between-person difference in performance that is practically equivalent to zero. 
+TOST provides two *p*-values, one testing against the upper bound and one testing against the lower bound; we report only the largest of the two *p*-values.
+
+# Results
+
+## Confirmatory analyses
+
+### Model fit
+
+The preregistered measurement model specification did not converge. 
+A model version excluding the covariance between manifest binding and updating did converge, but resulted in suboptimal fit (Robust CFI = `r fitstats_meas02[["cfi.robust"]] |> round(2)`, robust RMSEA = `r fitstats_meas02[["rmsea.robust"]] |> round(2)`, 95% CI = [`r fitstats_meas02[["rmsea.ci.lower.scaled"]] |> round(2)`, `r fitstats_meas02[["rmsea.ci.upper.scaled"]] |> round(2)`]).
+Modification indices indicated that model fit would improve most from estimating the covariance between Rotation Span and Operation Span, which is in line with previous factor models of working memory containing span tasks as a subset of other working memory tasks [e.g., @loffler_2024].
+A model incorporating an estimate of this covariance provided a good fit to the data (Robust CFI = `r fitstats_meas03[["cfi.robust"]] |> round(2)`, robust RMSEA = `r fitstats_meas03[["rmsea.robust"]] |> round(2)`, 95% CI = [`r fitstats_meas03[["rmsea.ci.lower.scaled"]] |> round(2)`, `r fitstats_meas03[["rmsea.ci.upper.scaled"]] |> round(2)`]).
+After finalizing the measurement model, we constructed the final structural model by adding all predictors and covariates to the model, which resulted in a good model fit (Robust CFI = `r fitstats_struc01[["cfi.robust"]] |> round(2)`, robust RMSEA = `r fitstats_struc01[["rmsea.robust"]] |> round(2)`, 95% CI = [`r fitstats_struc01[["rmsea.ci.lower.scaled"]] |> round(2)`, `r fitstats_struc01[["rmsea.ci.upper.scaled"]] |> round(2)`]).
+Figure 4 presents a visual overview of the final model.
+
+<br>
+
+```{r}
+#| label: Figure4
+#| fig-width: 6.5
+#| fig-height: 7
+#| dpi: 600
+#| out-width: 6in
+#| fig-cap: | 
+#|   **Figure 4.** Overview of the final measurement model of WM performance.
+Ellipses represent latent variables, rectangles represent manifest variables, and circles represent unstandardized residual variances.
+Unidirectional lines represent standardized factor loadings and bidirectional lines represent covariances.
+All four manifest WM measures loaded on a latent WM capacity factor, reflecting the fact that people have to hold information active in WM on all tasks.
+We fixed the loading of WM capacity on the Binding Task to 1, reflecting the idea that the ability to create and maintain bindings is the main limiting factor in WM capacity [@gruszka_2017; @oberauer_2009; @wilhelm_2013].
+WM updating was modeled as a latent factor capturing the residual variance in the updating task after accounting for variance related to WM capacity.
+WM = working memory; Ospan = Operation Span; Rspan = Rotation Span.
+knitr::include_graphics("fig/fig4.png")
+```
+
+<br>
+
+### Associations between adversity and WM
+
+The main results of the associations between the adversity measures and WM are summarized in Figure 5.
+None of the adversity measures were significantly associated with WM capacity after adjusting for multiple testing (all *p*s ≥ `r sem_reg_coef |> filter(lhs == "wmc_l", pvalue_adj == min(pvalue_adj)) |> pull(pvalue_adj) |> formatC(digits = 3, width = 3, format = 'f') |> str_remove(pattern = "^0")`).
+We also did not find evidence for practical equivalence for associations between any of the adversity measures and WM capacity (all *p*s ≥ `r equivalence_tests |> filter(lhs == "wmc_l", eq_pvalue == min(eq_pvalue)) |> pull(eq_pvalue) |> formatC(digits = 3, width = 3, format = 'f') |> str_remove(pattern = "^0")`).
+Similarly, none of the adversity measures were significantly associated with WM updating after adjusting for multiple testing (all *p*s ≥ `r sem_reg_coef |> filter(lhs == "upd_l", rhs == "inr_m") |> pull(pvalue_adj) |> formatC(digits = 3, width = 3, format = 'f') |> str_remove(pattern = "^0")`).
+We also did not find evidence for practical equivalence to zero for associations between any of the adversity measures and WM updating (all *p*s ≥ `r equivalence_tests |> filter(lhs == "upd_l") |> filter(eq_pvalue == min(eq_pvalue)) |> pull(eq_pvalue) |> formatC(digits = 3, width = 3, format = 'f') |> str_remove(pattern = "^0")`).
+
+<br>
+
+```{r}
+#| label: Figure5
+#| fig-width: 6.5
+#| fig-height: 4
+#| dpi: 600
+#| out-width: 6in
+#| fig-cap: | 
+#|   **Figure 5.** Results of the structural part of the SEM model testing the association between threat, deprivation, and unpredictability on latent estimates of WM capacity and WM updating.
+The gray area shows the area of practical equivalence.
+Solid points indicate effects outside the area of practical equivalence, which was true for all effects.
+Standard errors represent the 95% confidence intervals.
+CV = coefficient of variation; INR = income-to-needs ratio; M = mean; WM = working memory.
+fig5
+```
+<br>
+
+## Posthoc exploratory analyses
+
+We conducted two posthoc exploratory (non-preregistered) analyses, described in more detail in the supplemental materials.
+First, to contextualize our findings based on latent WM estimates, we estimated associations between adversity and performance on the separate WM tasks using four linear regressions.
+Threat had small, significant negative associations with performance on the Rotation Span Task ($\beta$ = `r lm_reg_coef_expl |> filter(term == "threat_comp", task == "rspan") |> pull(estimate) |> formatC(digits = 2, width = 2, format = 'f')`, *p* = `r lm_reg_coef_expl |> filter(term == "threat_comp", task == "rspan") |> pull(p.value) |> formatC(digits = 3, width = 3, format = 'f') |> str_remove("^0")`), Operation Span Task ($\beta$ = `r lm_reg_coef_expl |> filter(term == "threat_comp", task == "Ospan") |> pull(estimate) |> formatC(digits = 2, width = 2, format = 'f')`, *p* = `r lm_reg_coef_expl |> filter(term == "threat_comp", task == "Ospan") |> pull(p.value) |> formatC(digits = 3, width = 3, format = 'f') |> str_remove("^0")`), and Binding Task ($\beta$ = `r lm_reg_coef_expl |> filter(term == "threat_comp", task == "binding") |> pull(estimate) |> formatC(digits = 2, width = 2, format = 'f')`, *p* = `r lm_reg_coef_expl |> filter(term == "threat_comp", task == "binding") |> pull(p.value) |> formatC(digits = 3, width = 3, format = 'f') |> str_remove("^0")`).
+None of the types of adversity were significantly associated with performance on the Updating Task (all *p*s > `r lm_reg_coef_expl |> filter(task == "updating") |> pull(pvalue_adj) |> min() |> formatC(digits = 3, width = 3, format = 'f') |> str_remove("^0")`), and only the association with unpredictability in the income-to-needs was practically equivalent to zero (*p* = `r equivalence_tests_lm_expl |> filter(task == "updating", term == "inr_cv") |> pull(eq_pvalue) |> formatC(digits = 3, width = 3, format = 'f') |> str_remove("^0")`).
+
+Second, the inconclusive nature of our confirmatory results could indicate that the true effect sizes were smaller than the effect size of interest that we used for our power analysis ($\beta$ = 0.1; i.e, that we lacked sufficient power).
+To explore this, we conducted an alternative test for the absence of an association between adversity and WM by constraining regression paths between adversity and WM factors to zero in the SEM.
+Constraining all paths to latent WM capacity to zero significantly reduced model fit, although the change in AIC was below the cut-off as proposed by @burnham_2002, $\Delta$ AIC = `r (max(anova_expl02$AIC) - min(anova_expl02$AIC)) |> formatC(digits = 2, width = 2, format = 'f')`, $\Delta$ $\chi$(`r anova_expl02 |> filter(term == "fit_struc_expl02") |> pull(Df.diff)`) = `r anova_expl02 |> filter(term == "fit_struc_expl02") |> pull(Chisq.diff) |> formatC(digits = 2, width = 2, format = 'f')`, *p* = `r anova_expl02 |> filter(term == "fit_struc_expl02") |> pull(p.value) |> formatC(digits = 3, width = 3, format = 'f') |> str_remove("^0")`, Robust CFI = `r fitstats_expl02[["cfi.robust"]] |> round(2)`, robust RMSEA = `r fitstats_expl02[["rmsea.robust"]] |> round(2)`, 95% CI = [`r fitstats_expl02[["rmsea.ci.lower.scaled"]] |> round(2)`, `r fitstats_expl02[["rmsea.ci.upper.scaled"]] |> round(2)`].
+Constraining all paths to latent WM updating did not significantly reduce model fit, $\Delta$ AIC = `r (max(anova_expl03$AIC) - min(anova_expl03$AIC)) |> formatC(digits = 2, width = 2, format = 'f')`, $\Delta$ $\chi$(`r anova_expl03 |> filter(term == "fit_struc_expl03") |> pull(Df.diff)`) = `r anova_expl03 |> filter(term == "fit_struc_expl03") |> pull(Chisq.diff) |> formatC(digits = 2, width = 2, format = 'f')`, *p* = `r anova_expl03 |> filter(term == "fit_struc_expl03") |> pull(p.value) |> formatC(digits = 3, width = 3, format = 'f') |> str_remove("^0")`, Robust CFI = `r fitstats_expl03[["cfi.robust"]] |> round(2)`, robust RMSEA = `r fitstats_expl03[["rmsea.robust"]] |> round(2)`, 95% CI = [`r fitstats_expl03[["rmsea.ci.lower.scaled"]] |> round(2)`, `r fitstats_expl03[["rmsea.ci.upper.scaled"]] |> round(2)`].
+
+## Deviation from the Stage 1 protocol
+
+In the Stage 1 protocol, we planned to first access the dependent variables to construct the SEM, and then access the independent variables. 
+Due to an unintended error, the first author already accessed the datasets containing the measures that would be used to compute the independent variables before finalizing the SEM.
+However, beyond reading them into the R environment, these data were not yet inspected, manipulated, or summarized. 
+We contacted the PCI recommender upon finding out about this deviation, and agreed to describe this deviation as done here.
+For the sake of transparency, we timestamped the scripts for processing the independent variables at the moment of this unintended data access (<https://github.com/StefanVermeent/liss_wm_profiles_2023/blob/d143e551018ba27313643a15bed57f329974272d/scripts/2_pipeline/1_ivs.R>).
+They contain the code to read in the data, but no code yet for any type of data cleaning or variable computation.
+
+# Discussion
+
+We investigated associations between adversity (threat, material deprivation, and unpredictability) and WM capacity, a person's ability to hold information available for later processing, as well as WM updating, a person's ability to mentally replace old with new information.
+We distinguished between WM capacity and updating on a latent level using four different tasks, three of which are primarily construed as WM capacity tasks, and one that is primarily construed as a WM updating task.
+The WM capacity factor loaded on performance of all four tasks, in line with previous findings [@frischkorn_2022; @gruszka_2017; @oberauer_2009; @wilhelm_2013].
+An additional WM updating factor accounted for the portion of variance in the Updating Task that was not explained by WM capacity.
+
+We did not find any consistent associations between adversity and WM capacity nor updating in our preregistered analyses. 
+On the one hand, none of the associations significantly differed from zero. 
+On the other hand, none of the associations fell within the pre-specified region of practical equivalence to zero (i.e., a between-person difference in performance that is practically equivalent to zero).
+The conclusions from these confirmatory (preregistered) analyses differed in several respects from posthoc exploratory (non-preregistered) analyses focusing on associations between adversity and performance on the individual tasks.
+In these latter analyses, higher levels of exposure to neighborhood threat had small, significant negative associations with the Binding, Operation Span, and Rotation Span Tasks, which are all typically considered WM capacity tasks [@conway_2005; @wilhelm_2013].
+
+The confirmatory results were not consistent with hypotheses generated from a deficit framework.
+A large literature has documented negative associations between exposure to early-life adversity---especially deprivation---and WM capacity, which persists into adulthood [@farah_2006; @goodman_2019; @sheridan_2022; @sheridan_2014; @young_2018; but see @nweze_2021].
+Similarly, studies with young adults have found that a higher frequency of recent as well as lifetime stressful major life events (i.e., negative events with a clear onset and offset, unlike chronic adversity) is also negatively associated with WM capacity [@klein_2001; @shields_2017; @shields_2019].
+Exploratory analyses suggested there may be small associations between adversity and WM capacity, but that our tests of these associations were underpowered.
+This would mean that associations between adversity exposure in adulthood and WM capacity are smaller than we expected based on the literature outlined above, and would require a larger sample size to reliably detect.
+
+The results were also not consistent with hypotheses generated from adaptation frameworks.
+Recently, a small set of studies documented intact and even enhanced WM updating performance in adolescents and adults who reported more exposure to childhood adversity [@young_2018; @young_2022].
+These associations have been interpreted as reflecting developmental adaptations to adversity: in more threatening and unpredictable environments, it may be beneficial to be able to rapidly update the items held in WM [@frankenhuis_2013; @ellis_2017; @ellis_2022; @frankenhuis_2020].
+In contrast, we did not find consistent associations between adversity exposure and WM updating.
+These findings are inconclusive, as we also did not find evidence for practical equivalence in our preregistered analysis.
+However, additional exploratory analyses suggested that the association between adversity exposure in adulthood and WM updating was negligible.
+
+The Updating Task shared a large proportion of variance with the WM capacity measures, which aligns with prior psychometric work focused on the structure of WM [@lewandowsky_2010; @oberauer_2000; @wilhelm_2013].
+This highlights an important methodological issue for the field of adversity research, especially researchers working from adaptation frameworks, who hypothesize distinct effects of adversity on different components of WM (in contrast to deficit-oriented researchers, who expect adversity to have a negative effect on all components of WM).
+Specifically, adaptation-oriented researchers have hypothesized that certain types of adversity may enhance WM updating through developmental adaptation, while impairing WM capacity [@ellis_2022; @young_2018; @young_2022].
+So far, this hypothesis has---to our knowledge---only been tested based on raw performance on single WM updating tasks.
+However, if true, performance on single WM updating tasks may substantially underestimate positive associations between adversity and WM updating, as raw performance may be influenced by both deficit and adaptation processes (the former influencing WM capacity, inadvertently measured in WM updating tasks).
+Leveraging these psychometric insights will be pivotal to better understanding associations between adversity and WM for future studies.
+
+Aside from psychometric considerations, a second potential reason for the discrepancy between our findings and those from previous studies is that our investigation focused on adverse experiences in adulthood.
+In contrast, most previous studies have focused on the effects of either childhood adversity or stressful life events.
+It is possible that, relative to childhood adversity, the association between adversity in adulthood and WM varies as a function of other factors.
+For example, the association between adversity in adulthood and WM might be stronger for people who also experienced adversity during childhood, either due to early developmental calibration to chronic stress and/or due to greater lifetime exposure to stress [@hostinar_2013; @shields_2017].
+
+## Strengths, limitations, and future directions
+
+This study had several strengths. 
+First, the sample was drawn from the Dutch LISS panel, which provides a large, representative sample of the Dutch population.
+Second, we drew on the longitudinal nature of the LISS panel to estimate three key dimensions of adversity exposure (threat, deprivation, and unpredictability), using several indicators for each.
+Third, we included four WM tasks, and used SEM to separate variance related to WM capacity from variance related to WM updating.
+This allowed us to more precisely estimate capacity and updating as two key components of WM.
+
+This study also had limitations.
+First, WM updating was measured as the residual variance of a single task after accounting for WM capacity.
+This means that the latent WM updating measure was not a pure measure of WM updating, but also included measurement error.
+This decision was mainly guided by the limited number of tasks that could be included due to time constraints.
+To obtain a more reliable measure of WM updating, it would be better to include several different WM updating tasks, just like we used several different WM capacity tasks.
+Second, as this was an online study, we had only limited control over the environment in which people completed the study. 
+The models accounted for self-reported noise and distractions, and we excluded participants who interacted with other browser tabs during the WM tasks.
+Yet, there may have been other, unmeasured factors that could lower the reliability of our study relative to lab-based studies.
+Third, our results appeared to be underpowered, despite including 759 participants, which suggests that the associations between adversity and WM in adulthood are smaller than expected based on previous literature.
+Finally, our study did not include genetic measures. 
+It is well-established that genetic variation accounts for a substantial portion of the individual differences in executive functions [@friedman_2008]. 
+However, for genetics to have confounded our study, it would need to have caused both individual differences in cognition and in adversity exposures—producing non-causal associations between adversity and cognition.
+Testing this fuller picture would require using genetically informative designs.
+
+Future research could build on the current study in four ways.
+First, modeling WM ability on a latent level using multiple tasks could be applied more broadly in the field of adversity research, as studies rarely directly account for the overlap in key cognitive processes across WM tasks WM tasks.
+This is especially important for adaptation-based research focusing on WM updating ability, as WM capacity plays a substantial role in performance on updating tasks.
+Second, future work is needed to better understand the role of developmental timing: is adversity experienced earlier or later in life associated differently with WM across the lifespan?
+Third, more research is needed to better understand the relationship between more objective (e.g., income-to-needs ratio) and subjective (e.g., perceived scarcity) indicators of adversity, as well as their respective association with cognitive functioning [@smith_2021].
+In our study, mean INR and mean perceived scarcity correlated moderately, suggesting that they capture similar but separable aspects of material deprivation, which could show different associations with cognition.
+Fourth, the field needs to account for functional heterogeneity within adversity-exposed populations [@masten_2001]. 
+In a recent study, the majority of U.S. adolescents with low socioeconomic resources performed on par with their privileged peers [@shariq_2024]. 
+The deficit pattern observed in the population as a whole was driven by a much smaller, cognitively less resilient, subgroup. 
+A valuable direction is to combine such a ‘person-centered’ approach with structural equation modeling to estimate specific WM abilities among different subgroups within adversity-exposed populations.
+
+## Conclusion
+
+Over the last decade, adversity research has been shifting towards a more balanced view, focusing not just on cognitive deficits but also on potential adaptations.
+This has spurred a growing number of studies investigating more precise links between specific types of adversity and different cognitive abilities.
+Adaptation perspectives in particular have emphasized the need to be more precise about how specific types of adversity are associated with specific cognitive abilities.
+However, this increased need for precision in the measurement of cognitive abilities requires more advanced psychometric approaches.
+For this, adversity researchers can draw, more than they currently do, on decades of psychometric research focused on WM and other cognitive abilities.
+Doing so will ultimately lead to a better understanding of the unique abilities that develop in contexts of adversity, as well as more precise intervention targets.
+
+\pagebreak
+
+# Acknowledgements
+
+We would like to thank Kathrin Sadus for advising us on the implementation of the working memory tasks.
 

### 2024-06-25 manuscript/staging/stage_results.R

@@ -1,0 +1,261 @@
+
+# 1. Libraries ---------------------------------------------------------------
+
+library(tidyverse)
+library(lavaan)
+library(haven)
+
+source('scripts/0_custom_functions/corr_plot.R')
+
+# 2. Load data ------------------------------------------------------------
+
+clean_data <- read_csv("data/liss_data/clean_data.csv")
+descr <- read_sav(file = "data/liss_data/wm/L_Cognitieve_Vaardigheden_v2_1.0p.sav") |> as_tibble()
+
+load("analysis_objects/confirmatory_results.RData")
+load("analysis_objects/exploratory_results.RData")
+
+
+# 3. Descriptives table -------------------------------------------------------
+
+table1 <- descr |>
+  filter(nomem_encr %in% unique(clean_data$nomem_encr)) |>
+  select(p_leeftijd, geslacht) |>
+  mutate(geslacht = ifelse(geslacht == 1, "Male", "Female")) |>
+  summarise(
+    age_m = mean(p_leeftijd, na.rm = T),
+    age_sd = sd(p_leeftijd, na.rm = T),
+    sex = sum(geslacht == 'Female')/n()*100
+  ) |>
+  transmute(
+    `Mean age (SD)` = paste0(round(age_m, 1), " (", round(age_sd, 1), ")"),
+    `Sex (% Female)` = round(sex, 1) |> as.character()) |>
+  pivot_longer(everything(), names_to = "cat", values_to = "stat") |>
+  bind_rows(
+    descr |>
+      filter(nomem_encr %in% unique(clean_data$nomem_encr)) |>
+      select(oplmet) |>
+      mutate(
+        oplmet = case_when(
+          oplmet == 1 ~ "primary school",
+          oplmet == 2 ~ "vmbo (intermediate secondary education)",
+          oplmet == 3 ~ "havo/vwo (higher secondary education)",
+          oplmet == 4 ~ "mbo (intermediate vocational education)",
+          oplmet == 5 ~ "hbo (higher vocational education)",
+          oplmet == 6 ~ "wo (university)",
+          oplmet == 7 ~ "other",
+          TRUE ~ "missing"
+        ),
[…]

### 2024-09-06 scripts/2_pipeline/1_ivs.R
code of new non-preregistered analyses

@@ -189,7 +189,7 @@ assertthat::assert_that(all(crime_waves_parsed$q_101 %in% c(1,2, NA)))
 
 write_csv(crime_waves_parsed, "data/liss_data/crime/crime_waves_parsed.csv")
 
-#TODO: Sanity checks
+
 
 
 ## 1.5 Working Memory data ----

### 2024-09-06 scripts/2_pipeline/4_exploratory.R

@@ -3,6 +3,8 @@
 
 library(tidyverse)
 library(lavaan)
+library(bain)
+library(flextable)
 
 # 2. Load data ------------------------------------------------------------
 
@@ -198,7 +200,40 @@ anova_expl03 <- anova(fit_struc01, fit_struc_expl03) |>
 fitstats_expl03 <- fitMeasures(fit_struc_expl03)[c("cfi.robust", "rmsea.robust", "rmsea.ci.lower.scaled", "rmsea.ci.upper.scaled")]
 
 
+## 6. Robustness check using Bayes Factors
+
+set.seed(35721)
+
+bay_wmc_inr_cv <- bain(fit_struc01, hypothesis = 'wmc_l~inr_cv < 0.1      & wmc_l~inr_cv > -0.1', fraction = 1, standardize = TRUE)
+bay_wmc_inr_m <- bain(fit_struc01, hypothesis  = 'wmc_l~inr_m < 0.1       & wmc_l~inr_m > -0.1', fraction = 1, standardize = TRUE)
+bay_wmc_psc_cv <- bain(fit_struc01, hypothesis = 'wmc_l~p_scar_cv < 0.1   & wmc_l~p_scar_cv > -0.1', fraction = 1, standardize = TRUE)
+bay_wmc_psc_m <- bain(fit_struc01, hypothesis  = 'wmc_l~p_scar_m < 0.1    & wmc_l~p_scar_m > -0.1', fraction = 1, standardize = TRUE)
+bay_wmc_thr <- bain(fit_struc01, hypothesis    = 'wmc_l~threat_comp < 0.1 & wmc_l~threat_comp > -0.1', fraction = 1, standardize = TRUE)
+
+bay_upd_inr_cv <- bain(fit_struc01, hypothesis = 'upd_l~inr_cv < 0.1      & upd_l~inr_cv > -0.1', fraction = 1, standardize = TRUE)
+bay_upd_inr_m <- bain(fit_struc01, hypothesis  = 'upd_l~inr_m < 0.1       & upd_l~inr_m > -0.1', fraction = 1, standardize = TRUE)
+bay_upd_psc_cv <- bain(fit_struc01, hypothesis = 'upd_l~p_scar_cv < 0.1   & upd_l~p_scar_cv > -0.1', fraction = 1, standardize = TRUE)
[…]

### 2024-09-06 supplement/supplement.qmd
updated supplemental materials with new non-preregistered analyses

@@
 
-To contextualize our confirmatory (preregistered) findings, we conducted twoexploratory (non-preregistered) analyses.
-Below, we present the results of analyses exploring the associations between adversity and performance on the separate WM tasks using linear regression.
-The other exploratory analyses, which were based on alternative SEM specifications, are presented in the supplemental materials.
+To contextualize our confirmatory (preregistered) findings, we conducted three exploratory (non-preregistered) analyses.
+First, we explored associations between adversity and performance on the separate WM tasks using linear regression.
+Second, we constrained regression paths in the SEM to zero, as an alternative to the equivalence tests.
+Third, we computed Bayes Factors for the equivalence tests.
+
+## 3.1 Linear regression analyses
 
@@
 <br>
+
+## 3.2 Bayes Factors for equivalence tests
+
+We calculated Bayes factors for the preregistered equivalence tests using the *bain* package [@hoijtink_2019].
+We evaluated evidence in favor of the hypothesis that the regression coefficients fell within the equivalence bounds ($\beta$ ≤ -0.1 and $\beta$ ≥ 0.1)
+
+Third, as a robustness check, we calculated Bayes factors for the preregistered equivalence tests using the *bain* package [@hoijtink_2019], in which we evaluated evidence in favor of the hypothesis that the effects fell within the equivalence bounds, relative to the hypothesis that the effects fell outside the equivalence bounds.
+The results are summarized in Table S3.
+For all but one association, the model comparisons showed at least strong evidence in favor of the hypothesis that the effects fell within the equivalence bounds (BF~10~ ranging between `r BF_expl04 |> filter(Hypothesis != '-0.1 ≤ (WM capacity ~ Threat) ≥ 0.1') |> pull(BF.c) |> min()` and `r BF_expl04 |> filter(Hypothesis != '-0.1 ≤ (WM capacity ~ Threat) ≥ 0.1') |> pull(BF.c) |> max()`.
+The only exception was the association between threat and WM capacity, for which we found moderate evidence for the hypothesis that the effect fell within the equivalence bounds (BF~10~ = `r BF_expl04 |> filter(Hypothesis == '-0.1 ≤ (WM capacity ~ Threat) ≥ 0.1') |> pull(BF.c)`.
+Thus, these results were inconsistent with the preregistered frequentist equivalent tests.
+
+```{r}
+table_s3 <- BF_expl04 |> 
+  select(Hypothesis, BF.c) |>
+  flextable::flextable() |>
+  compose(i = 1, j = 2, as_paragraph("BF", as_sub('10')), part = "header") %>%
+  border_remove() %>% 
+  add_header_row(
+    values = " ",
+    colwidths = 2
+  ) |> 
+  flextable::compose(
+    i = 1, j = 1, 
+    as_paragraph(as_b("Table S3. "), "Bayes Factors for practical equivalence tests."),
+    part = "header"
+  ) |> 
+  add_footer_row(
+    values = " ",
+    colwidths = 2
+  ) |> 
+  flextable::compose(
+    i = 1, j = 1, 
+    as_paragraph("BF = Bayes factor; CV = coefficient of variance, INR = income-to-needs ratio, M = mean, Perc. Scarcity = perceived scarcity, WM = working memory"), 
+    part = "footer"
+  ) |> 
+  border(i = 1, border.bottom = fp_border_default(), part = "header") %>% 
+  border(i = 2, border.bottom = fp_border_default(), part = "header") %>% 
+  border(i = 10, border.bottom = fp_border_default(), part = "body") %>% 
+  autofit()
+
+table_s3
+```
 

### 2024-09-06 manuscript/manuscript.qmd
create a manuscript with tracked changes
supporting files for stage 2 revised manuscript

@@
 
-#### **Working memory performance in adverse environments: Enhanced, impaired, or intact?**
+#### **Inconclusive evidence for associations between adverse experiences in adulthood and working memory performance**
 
@@
 
-# Working memory performance in adverse environments: Enhanced, impaired, or intact?
+# Inconclusive evidence for associations between adverse experiences in adulthood and working memory performance
 
@@
 For early-life adversity, this negative association is already present during childhood, and persists into adulthood [@bos_2009; @evans_2009; @hackman_2010; @farah_2006; @goodman_2019; @noble_2007; but see @nweze_2021].
-Studies with college students have found a link between both recent and lifetime experiences of stressful major life events (discrete negative events that have a clear onset and offset, unlike chronic adversity) with lowered WM capacity [@klein_2001; @shields_2017; @shields_2019].
+Studies with college students have found a link between both recent and lifetime experiences of stressful major life events (discrete negative events that have a clear onset and offset, unlike chronic adversity) with lower WM capacity [@klein_2001; @shields_2017; @shields_2019].
 
@@
 Crucially, as deficit and adaptation processes can operate in concert [@frankenhuis_2020], we could find support (or lack thereof) for both frameworks in the same model. 
-We distinguished between three between-person data patterns: (1) lowered performance, (2) enhanced performance, and (3) intact performance.
-We defined lowered performance as a statistically significant negative association between a type of adversity and WM capacity or updating (irrespective of effect size).
-We defined enhanced performance as a statistically significant positive association between a type of adversity and WM capacity or updating (irrespective of effect size).
-We defined intact performance as an association between a type of adversity and WM capacity or updating that has a standardized effect smaller than 0.1 *and* larger than -0.1---even if the effect is statistically different from zero---which we tested using Two One-Sided T-Tests (TOST) equivalence testing [see the 'Primary analyses' section\; @lakens_2018].
+We distinguished between three between-person data patterns: (1) lower performance, (2) higher performance, and (3) practically equivalent performance.
+We defined lower performance as a statistically significant negative association between a type of adversity and WM capacity or updating (irrespective of effect size).
+We defined higher performance as a statistically significant positive association between a type of adversity and WM capacity or updating (irrespective of effect size).
+We defined practically equivalent performance as an association between a type of adversity and WM capacity or updating that has a standardized effect smaller than 0.1 *and* larger than -0.1---even if the effect is statistically different from zero---which we tested using Two One-Sided T-Tests (TOST) equivalence testing [see the 'Primary analyses' section\; @lakens_2018].
 
@@
 First, if adaptive processes enhance WM updating and there are no impairment processes operating, we can expect a positive association between adversity and WM updating. 
-Second, if, adaptive processes operate in concert with general impairment processes, we can expect intact WM updating in combination with lowered WM capacity. 
-If neither impairment nor adaptative processes are operating, we can expect both WM updating and capacity to be intact.
+Second, if, adaptive processes operate in concert with general impairment processes, we can expect practically equivalent WM updating performance in combination with lower WM capacity. 
+If neither impairment nor adaptative processes are operating, we can expect both WM updating and capacity to be practically equivalent.
 
@@
 
-In addition, we implemented the Neighborhood Violence Scale [NVS\; @frankenhuis_2018; @frankenhuis_2020] in the newly collected data.
-The NVS includes seven items measuring perceived exposure to neighborhood violence (e.g., "Crime is common in the neighborhood where I live"; "Where I live, it is important to be able to defend yourself against physical harm").
+In addition, we implemented the Neighborhood Violence Scale [@frankenhuis_2018; @frankenhuis_2020] in the newly collected data.
+The Neighborhood Violence Scale includes seven items measuring perceived exposure to neighborhood violence (e.g., "Crime is common in the neighborhood where I live"; "Where I live, it is important to be able to defend yourself against physical harm").
 Participants answered these questions on a scale of 1 ("Completely disagree") to 7 ("Completely agree").
@@
 Because correlations were below .60 (see Table 2), we then used PCA to extract only the first principal component score (R^2 = .20).
-The threat component was most strongly determined by the NVS (`r pca_threat_fit$loadings[3] |> formatC(digits = 2, width = 2, format = 'f')`), followed by the perceived neighborhood crime scale from the LISS archive (`r pca_threat_fit$loadings[1] |> formatC(digits = 2, width = 2, format = 'f')`) and crime victimization (`r pca_threat_fit$loadings[2] |> formatC(digits = 2, width = 2, format = 'f')`).
+The threat component was most strongly determined by the Neighborhood Violence Scale (`r pca_threat_fit$loadings[3] |> formatC(digits = 2, width = 2, format = 'f')`), followed by the perceived neighborhood crime scale from the LISS archive (`r pca_threat_fit$loadings[1] |> formatC(digits = 2, width = 2, format = 'f')`) and crime victimization (`r pca_threat_fit$loadings[2] |> formatC(digits = 2, width = 2, format = 'f')`).
 
@@
 If monthly household income is missing (or zero) for an entire year for a participant, we used, if available, the yearly net household income they reported in the LISS archive ([https://doi.org/10.57990/1gr4-bf42](https://doi.org/10.57990/1gr4-bf42)), dividing it by 12 to obtain a monthly estimate.
-First, we divided the average income per year by the *poverty threshold*, as determined by Statistics Netherlands (CBS) [@brakel_2023\; CBS, personal communication, December 15, 2023].
+First, we divided the average income per year by the *poverty threshold*, as determined by Statistics Netherlands [@brakel_2023\; CBS, personal communication, December 15, 2023].
 As thresholds are only provided for households with up to three children, we applied the threshold of a household with three children to households with more than three children.
@@
 We computed unpredictability over time in perceived scarcity using the coefficient of variation, which is the within-person standard deviation across years divided by the mean [@key_2017; @liu_2022; @ugarte_2023; @walasek_2024; @young_2020].
-The mean and standard deviation in income have been found to be strongly negatively correlated, indicating that people with lower incomes tend to experience less variability in income [@li_2018; @young_2023].
+The mean and standard deviation in income have been found to be strongly negatively correlated, indicating that people with lower incomes tend to experience less variability in income [@li_2018; @young_2024].
 For that reason, the standard deviation alone has been called into question as a measure of adversity, as the same fluctuation in income can have a greater relative impact for people close to the poverty line than for people with high incomes.
@@
 
+The Stage 1 protocol of this Registered Report can be found at [https://osf.io/dp7wc](https://osf.io/dp7wc).
+
 ### Data access
@@
 
-To statistically test whether small effects were practically equivalent to zero---suggesting intact performance---we used Two One-Sided T-tests (TOST) equivalence testing [@lakens_2018], using -0.1 and 0.1 as equivalence bounds.
-TOST equivalence testing allows us to conclude intact performance based on a significant effect, rather than erroneously interpreting a non-significant effect as evidence for the absence of an effect.
+To statistically test whether small effects were practically equivalent to zero we used Two One-Sided T-tests (TOST) equivalence testing [@lakens_2018], using -0.1 and 0.1 as equivalence bounds.
+TOST equivalence testing allows us to conclude practically equivalent performance based on a significant effect, rather than erroneously interpreting a non-significant effect as evidence for the absence of an effect.
 We considered any effect that fell within this region to reflect practical equivalence, that is, a between-person difference in performance that is practically equivalent to zero. 
@@
 
-## Posthoc exploratory analyses
+## Posthoc non-preregistered analyses
 
-We conducted two posthoc exploratory (non-preregistered) analyses, described in more detail in the supplemental materials.
+We conducted three posthoc non-preregistered analyses, described in more detail in the supplemental materials.
 First, to contextualize our findings based on latent WM estimates, we estimated associations between adversity and performance on the separate WM tasks using four linear regressions.
@@
 Constraining all paths to latent WM updating did not significantly reduce model fit, $\Delta$ AIC = `r (max(anova_expl03$AIC) - min(anova_expl03$AIC)) |> formatC(digits = 2, width = 2, format = 'f')`, $\Delta$ $\chi$(`r anova_expl03 |> filter(term == "fit_struc_expl03") |> pull(Df.diff)`) = `r anova_expl03 |> filter(term == "fit_struc_expl03") |> pull(Chisq.diff) |> formatC(digits = 2, width = 2, format = 'f')`, *p* = `r anova_expl03 |> filter(term == "fit_struc_expl03") |> pull(p.value) |> formatC(digits = 3, width = 3, format = 'f') |> str_remove("^0")`, Robust CFI = `r fitstats_expl03[["cfi.robust"]] |> round(2)`, robust RMSEA = `r fitstats_expl03[["rmsea.robust"]] |> round(2)`, 95% CI = [`r fitstats_expl03[["rmsea.ci.lower.scaled"]] |> round(2)`, `r fitstats_expl03[["rmsea.ci.upper.scaled"]] |> round(2)`].
+Thus, these results were somewhat inconsistent with the preregistered frequentist equivalent tests.
+
+Third, as a non-preregistered robustness check, we calculated Bayes factors for the preregistered equivalence tests using the *bain* package [@hoijtink_2019], in which we evaluated whether the observed data are more likely under the hypothesis that the effects fall within the equivalence bounds, relative to the hypothesis that the effects fall outside of the equivalence bounds.
+The results are summarized in Table S3.
+For all but one association, the model comparisons showed at least strong evidence in favor of the data being more likely under the hypothesis that the effects fell within the equivalence bounds (BF~10~ ranging between `r BF_expl04 |> filter(Hypothesis != '-0.1 ≤ (WM capacity ~ Threat) ≥ 0.1') |> pull(BF.c) |> min()` and `r BF_expl04 |> filter(Hypothesis != '-0.1 ≤ (WM capacity ~ Threat) ≥ 0.1') |> pull(BF.c) |> max()`.
+The only exception was the association between threat and WM capacity, for which we found moderate evidence in favor of the data being more likely under the hypothesis that the effect fell within the equivalence bounds (BF~10~ = `r BF_expl04 |> filter(Hypothesis == '-0.1 ≤ (WM capacity ~ Threat) ≥ 0.1') |> pull(BF.c)`.
+Thus, these results were inconsistent with the preregistered frequentist equivalent tests, which did not find evidence for practical equivalence.
 
@@
 An additional WM updating factor accounted for the portion of variance in the Updating Task that was not explained by WM capacity.
-
 We did not find any consistent associations between adversity and WM capacity nor updating in our preregistered analyses. 
@@
 On the other hand, none of the associations fell within the pre-specified region of practical equivalence to zero (i.e., a between-person difference in performance that is practically equivalent to zero).
-The conclusions from these confirmatory (preregistered) analyses differed in several respects from posthoc exploratory (non-preregistered) analyses focusing on associations between adversity and performance on the individual tasks.
-In these latter analyses, higher levels of exposure to neighborhood threat had small, significant negative associations with the Binding, Operation Span, and Rotation Span Tasks, which are all typically considered WM capacity tasks [@conway_2005; @wilhelm_2013].
 
@@
 Similarly, studies with young adults have found that a higher frequency of recent as well as lifetime stressful major life events (i.e., negative events with a clear onset and offset, unlike chronic adversity) is also negatively associated with WM capacity [@klein_2001; @shields_2017; @shields_2019].
-Exploratory analyses suggested there may be small associations between adversity and WM capacity, but that our tests of these associations were underpowered.
-This would mean that associations between adversity exposure in adulthood and WM capacity are smaller than we expected based on the literature outlined above, and would require a larger sample size to reliably detect.
-
 The results were also not consistent with hypotheses generated from adaptation frameworks.
-Recently, a small set of studies documented intact and even enhanced WM updating performance in adolescents and adults who reported more exposure to childhood adversity [@young_2018; @young_2022].
+Recently, a small set of studies documented intact and even higher WM updating performance in adolescents and adults who reported more exposure to childhood adversity [@young_2018; @young_2022].
 These associations have been interpreted as reflecting developmental adaptations to adversity: in more threatening and unpredictable environments, it may be beneficial to be able to rapidly update the items held in WM [@frankenhuis_2013; @ellis_2017; @ellis_2022; @frankenhuis_2020].
@@
 These findings are inconclusive, as we also did not find evidence for practical equivalence in our preregistered analysis.
-However, additional exploratory analyses suggested that the association between adversity exposure in adulthood and WM updating was negligible.
+
+A set of non-preregistered robustness checks were comparatively more consistent with practically equivalent performance, although they did not fully rule out the existence of small associations between adversity exposure and working memory performance.
+First, A Bayesian reanalysis of the preregistered equivalence tests (using the same equivalence bounds) provided strong evidence in favor of the hypothesis that working memory performance was practically equivalent, in contrast to the preregistered analyses.
+Second, constraining the regression paths in the SEM to zero somewhat reduced model fit for WM capacity, but not for WM updating.
+This suggests that there may have been systematic associations with WM capacity that were smaller than the equivalence bounds used in the (Bayesian) equivalence tests.
+If true, the associations would be smaller than we expected based on the literature outlined above, and would require a larger sample size to reliably detect.
+These analyses were not part of the registered analysis protocol, and therefore should be interpreted with sufficient caution pending replication.
 
@@
 For this, adversity researchers can draw, more than they currently do, on decades of psychometric research focused on WM and other cognitive abilities.
-Doing so will ultimately lead to a better understanding of the unique abilities that develop in contexts of adversity, as well as more precise intervention targets.
+Here, our psychometric investigation of WM yielded inconclusive associations with adverse experiences in adulthood.
+Building on this work will ultimately lead to a better understanding of the unique abilities that develop in contexts of adversity, as well as more precise intervention targets.
 
@@
 First, if adaptive processes enhance WM updating and there are no impairment processes operating, we can expect a positive association between adversity and WM updating. 
-Second, if, adaptive processes operate in concert with general impairment processes, we can expect intact WM updating in combination with lowered WM capacity.
+Second, if, adaptive processes operate in concert with general impairment processes, we can expect intact WM updating in combination with lower WM capacity.
 If neither impairment nor adaptative processes are operating, we can expect both WM updating and capacity to be intact.",
@@
     "Contrary to predictions of deficit perspectives, we might find that all associations between adversity and WM capacity are either practically equivalent or positive.
-This would suggest that WM capacity is either unaffected or even enhanced by adversity.
+This would suggest that WM capacity is either unaffected or even higher in people with more adversity exposure.
 
@@
 We might also find that material deprivation, but not unpredictability or neighborhood threat, is positively associated with WM updating.
-This would suggest that an enhanced updating ability has an adaptive benefit for individuals experiencing material deprivation."
+This would suggest that higher updating ability has an adaptive benefit for individuals experiencing material deprivation."
     ),

### 2024-09-06 manuscript/staging/stage_results.R

@@ -65,8 +65,8 @@ table1 <- descr |>
       ) |>
       pivot_longer(everything(), names_to = "cat", values_to = "stat")
   ) |>
-  add_row(.before = 3, cat = "Highest completed education") |>
-  add_row(.before = 12, cat = "Number of waves") |>
+  add_row(.before = 3, cat = "Highest completed education (%)") |>
+  add_row(.before = 12, cat = "Mean number of waves (SD)") |>
   flextable() |>
   set_header_labels(cat = "Category", stat = "Statistic") |>
   padding(i = c(4:11, 13:15), j = 1, padding.left = 10)
@@ -96,7 +96,7 @@ table2 <- clean_data |>
       "Current situation (CV)",
       "Perceived scarcity (CV)",
       "Neighborhood safety",
-      "NVS",
+      "Neighborhood Violence Scale",
       "Crime victimization",
       "Threat"
     ),

### 2024-10-10 manuscript/manuscript.qmd
changes to conclusion for Stage 2 revision 2

@@
 
-Over the last decade, adversity research has been shifting towards a more balanced view, focusing not just on cognitive deficits but also on potential adaptations.
+Our psychometric investigation yielded inconclusive evidence for associations between adverse experiences in adulthood and WM capacity and updating ability: Differences in abilities were not significantly different from zero, yet also not negligibly small.
+This study is part of a recent shift in adversity research towards a more balanced view, focusing not just on cognitive deficits but also on potential adaptations.
 This has spurred a growing number of studies investigating more precise links between specific types of adversity and different cognitive abilities.
@@
 For this, adversity researchers can draw, more than they currently do, on decades of psychometric research focused on WM and other cognitive abilities.
-Here, our psychometric investigation of WM yielded inconclusive associations with adverse experiences in adulthood.
-Building on this work will ultimately lead to a better understanding of the unique abilities that develop in contexts of adversity, as well as more precise intervention targets.
+Our findings suggest that this may lead to a more complicated picture compared to traditional investigations into raw performance.
+However, this will ultimately lead to a better understanding of the unique abilities that develop in contexts of adversity, as well as more precise intervention targets.
 
